### 2018-05-14 morgue_supp_mat.Rmd
create morgue for old ideas

---
title: "Morgue: Supplementary material"
output: html_notebook
---

GAM estimates of selection gradients differ from GAMM

```{r COMPARE GAM WITH GAMM}

## CONTROL PLANTS ----

## Gall Height
gam_red.control_size <- gam(gall_survival ~
                             s(c.Gall_Height_mm) + 
                             s(c.gall_individuals, k=9) +
                             s(c.Density_per_100_shoots),
                           data = control_df, family = binomial(link = "logit"))
summary(gam_red.control_size)

gam.plot(gam_red.control_size, pages=1)
gam.plot(gamm_red.control$gam, pages=1)

## Gall individuals
control_df_gall.level <- control_df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise(gall_survival_mean = mean(gall_survival),
            gall_individuals = mean(gall_individuals),                 # note that there is no variance, it is the actual value
            Gall_Height_mm_mean = mean(Gall_Height_mm),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>% # remember that this is pseudo-replicated at the gall-level
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals = gall_individuals - mean(gall_individuals),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.control_indiv <- gam(gall_survival_mean ~
                              s(c.Gall_Height_mm_mean) + 
                              s(c.gall_individuals, k=9) +
                              s(c.Density_per_100_shoots),
                             data = control_df_gall.level, family = binomial(link = "logit")) # we didn't include weights because this would be correlated with the number of individuals per gall
summary(gam_red.control_indiv)

gam.plot(gam_red.control_indiv, pages=1)
gam.plot(gamm_red.control$gam, pages=1)

## Gall density
control_df_plant.level <- control_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%    # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

#test.control_df_plant.level <- control_df %>%
#  group_by(Treatment.focus, Plant_Position) %>%
#  summarise(gall_survival_mean = mean(gall_survival),
#            gall_sample_size = n(),
#            gall_individuals_mean = mean(gall_individuals),                 
#            Gall_Height_mm_mean = mean(Gall_Height_mm),
#            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%                # note that there is no variance, it is the actual value
#  ungroup() %>%
#  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
#         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
#         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

gam_red.control_density <- gam(gall_survival_mean ~
                                s(c.Gall_Height_mm_mean) + 
                                s(c.gall_individuals_mean, k=9) +
                                s(c.Density_per_100_shoots, k=3),  # adjusting k value to avoid overfitting of Density   
                             data = control_df_plant.level,
                             family = binomial(link = "logit"), 
                             weights = gall_sample_size) 
summary(gam_red.control_density)

gam.plot(gam_red.control_density, pages=1)
gam.plot(gamm_red.control$gam, pages=1)

## TREATMENT PLANTS ----

## Gall Height
gam_red.treatment_size <- gam(gall_survival ~
                         s(c.Gall_Height_mm) + 
                         s(c.gall_individuals, k=9) +
                         s(c.Density_per_100_shoots),
                       data = treatment_df, family = binomial(link = "logit"))
summary(gam_red.treatment_size)

gam.plot(gam_red.treatment_size, pages=1)
gam.plot(gamm_red.treatment$gam, pages=1)

## Gall individuals
treatment_df_gall.level <- treatment_df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise(gall_survival_mean = mean(gall_survival),
            gall_individuals = mean(gall_individuals),              # note that there is no variance, it is the actual value
            Gall_Height_mm_mean = mean(Gall_Height_mm),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>% # remember this is pseudo-replicated at the gall-level
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals = gall_individuals - mean(gall_individuals),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.treatment_indiv <- gam(gall_survival_mean ~
                              s(c.Gall_Height_mm_mean) + 
                              s(c.gall_individuals, k=9) +
                              s(c.Density_per_100_shoots),
                             data = treatment_df_gall.level, family = binomial(link = "logit"))
summary(gam_red.treatment_indiv)

gam.plot(gam_red.treatment_indiv, pages=1)
gam.plot(gamm_red.treatment$gam, pages=1)

## Gall density
treatment_df_plant.level <- treatment_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%  # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

gam_red.treatment_density <- gam(gall_survival_mean ~
                                s(c.Gall_Height_mm_mean) + 
                                s(c.gall_individuals_mean, k=9) +
                                s(c.Density_per_100_shoots),
                             data = treatment_df_plant.level, family = binomial(link = "logit"), 
                             weights = gall_sample_size) 
summary(gam_red.treatment_density)

gam.plot(gam_red.treatment_density, pages=1)
gam.plot(gamm_red.treatment$gam, pages=1)

#### CALCULATE SELECTION GRADIENTS FROM GAM ----

## GALL SIZE ----
sg.gam.size_control <- gradient.calc(mod = gam_red.control_size, 
                                 phenotype = "c.Gall_Height_mm", 
                                 covariates = c("c.gall_individuals","c.Density_per_100_shoots"))
sg.gam.size_treatment <- gradient.calc(mod = gam_red.treatment_size, 
                                   phenotype = "c.Gall_Height_mm", 
                                   covariates = c("c.gall_individuals","c.Density_per_100_shoots"))

## GALL INDIVIDUALS ----
sg.gam.individuals_control <- gradient.calc(mod = gam_red.control_indiv, 
                                        phenotype = "c.gall_individuals", 
                                        covariates = c("c.Gall_Height_mm_mean","c.Density_per_100_shoots"))
sg.gam.individuals_treatment <- gradient.calc(mod = gam_red.treatment_indiv, 
                                          phenotype = "c.gall_individuals", 
                                          covariates = c("c.Gall_Height_mm_mean","c.Density_per_100_shoots"))

## GALL DENSITY ----
sg.gam.density_control <- gradient.calc(mod = gam_red.control_density, 
                                    phenotype = "c.Density_per_100_shoots", 
                                    covariates = c("c.gall_individuals_mean","c.Gall_Height_mm_mean"))
sg.gam.density_treatment <- gradient.calc(mod = gam_red.treatment_density, 
                                      phenotype = "c.Density_per_100_shoots", 
                                      covariates = c("c.gall_individuals_mean","c.Gall_Height_mm_mean"))

## PRINT TABLE ----
rbind(
  mutate(sg.gam.size_control$ests, Food_Web_Treatment = "Complex"),
  mutate(sg.gam.size_treatment$ests, Food_Web_Treatment = "Simple"),
  mutate(sg.gam.individuals_control$ests, Food_Web_Treatment = "Complex"),
  mutate(sg.gam.individuals_treatment$ests, Food_Web_Treatment = "Simple"),
  mutate(sg.gam.density_control$ests, Food_Web_Treatment = "Complex"),
  mutate(sg.gam.density_treatment$ests, Food_Web_Treatment = "Simple")
)

round(rbind(
  sg.gam.size_control$ests,
  sg.gam.size_treatment$ests,
  sg.gam.individuals_control$ests,
  sg.gam.individuals_treatment$ests,
  sg.gam.density_control$ests,
  sg.gam.density_treatment$ests
),3)

## BOOTSTRAPPED INTERVALS ----
gam.bootstrap_df <- bind_rows(
  data.frame(Bootstrap_Estimates = sg.gam.size_control$boot[ ,1], Phenotype = "c.Gall_Height_mm", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.size_control$boot[ ,2], Phenotype = "c.Gall_Height_mm", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.size_treatment$boot[ ,1], Phenotype = "c.Gall_Height_mm", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.size_treatment$boot[ ,2], Phenotype = "c.Gall_Height_mm", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_control$boot[ ,1], Phenotype = "c.gall_individuals", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_control$boot[ ,2], Phenotype = "c.gall_individuals", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_treatment$boot[ ,1], Phenotype = "c.gall_individuals", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.individuals_treatment$boot[ ,2], Phenotype = "c.gall_individuals", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.density_control$boot[ ,1], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.density_control$boot[ ,2], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.gam.density_treatment$boot[ ,1], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.gam.density_treatment$boot[ ,2], Phenotype = "c.Density_per_100_shoots", Selection_Type = "Gamma", Food_Web_Treatment = "Simple")
)

gam.bootstrap_df %>%
  ggplot(., aes(x=Phenotype, y=Bootstrap_Estimates, fill=Food_Web_Treatment)) +
  geom_boxplot() +
  geom_hline(yintercept=0, linetype="dotted") +
  facet_wrap(~Selection_Type) +
  coord_flip()

```