### 2018-06-22 elsevier_test/elsevier_test.tex
updating methods section

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
%\makeatletter
%\def\ps@pprintTitle{%
% \let\@oddhead\@empty
% \let\@evenhead\@empty
% \def\@oddfoot{\it \hfill\today}%
% \let\@evenfoot\@oddfoot}
%\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Food-web complexity flattens the fitness landscape of an insect herbivore},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\usepackage[nomarkers]{endfloat}

\begin{document}
\begin{frontmatter}

  \title{Food-web complexity flattens the fitness landscape of an insect
herbivore}
    \author[Department of Evolutionary Biology and Environmental Studies, University
of Zurich, Zurich, Switzerland]{Matthew A. Barbour\corref{c1}}
   \ead{matthew.barbour@ieu.uzh.ch} 
   \cortext[c1]{Corresponding Author}
    \author[Another University]{Jordi Bascompte}
   \ead{jordi.bascompte@ieu.uzh.ch}

      \address[Some Institute of Technology]{Department, Street, City, State, Zip}
    \address[Another University]{Department, Street, City, State, Zip}

  \begin{abstract}
  true
  \end{abstract}

 \end{frontmatter}

\emph{Text based on elsarticle sample manuscript, see
\url{http://www.elsevier.com/author-schemas/latex-instructions\#elsarticle}}

\section{Introduction}\label{introduction}

Biological diversity -- from genes, to phenotypes, to species -- has
fascinated evolutionary biologists for decades.

Much of this biological diversity has been shaped by natural selection
via trophic interactions, such as resource competition (cite charater
displacement), mutualistic exchanges of resource (cite? ), and predation
(cite recent Heath and Stireman paper).

NEED to recognize that evolutionary biologists have begun to explore how
community context affects evolutionary change (work by Sharon Strauss,
Casey terHorst, Lutz Becks, etc.). These results have begun to show
interesting patterns whereby the composition of species in a community
can alter the direction and strength of natural selection imposed on
species embedded within these communities (cite).

Put another way, these results have begun to show biological diversity,
in terms of differences between species, can drive evolutionary change.

While there is clear evidence for pairwise trophic interactions to drive
evolution by natural selection, its unclear how biological diversity
itself imposes natural selection and drives evolutionary change.

Exploring the effects of biological diversity requires an explicit
examination of the network structure of trophic interactions between
species in a community. Theoretical models have begun to examine how the
network structure of species interactions drives evolutionary change
(Nuismer paper; Guimeras paper; Ecology Letters paper from a spanish
guy\ldots{}); however, we are currently lacking experimental tests.

In contrast, ecologists have begun to embrace the complexity of the
natural world, and seeking to identify the complex networks of
interactions that underlie community structure and ecosystem function.
However, these studies have not examined how evolutionary processes
feedback to shape the structure and evolution of these interaction
networks.

Food-web complexity may influence selection gradients in at least two
ways. First, if more diverse predator communities are more efficient as
suppressing prey densities (e.g.~biodiversity-ecosystem fucntion; Ives
2005 Ecology Letters), then this will result in lower mean fitness. A
reduction in mean fitness, all else equal, will intensify natural
selection and thus could increase the rate of evolutionary change.
Alternatively, if predators are functionally distinct, more diverse
communities can reduce the strength of selection. This is because each
predator has a different functional relationship between prey phenotype
and the probability of an interaction.

Here, we conducted a field experiment to test the effect of food-web
complexity on the fitness landscape of a species embedded within this
food web. To do this, we used a common garden experiment with a host
plant (\emph{Salix hooeriana}), an abundant herbivore (\emph{Iteomyia
salicisverruca}), and the diverse community of insect parasitoids that
attack it. Prior work in this system has shown that there is directional
selection for larger galls, likely because larger galls provide more of
a refuge from parasitoid attack. However, there is also evidence that
different parasitoid species impose differential selection on gall
phenotypes.

\section{Materials \& Methods}\label{materials-methods}

We conducted our study within a four-year old common garden of
\emph{Salix hookeriana} located at Humboldt Bay National Wildlife Refuge
(HBNWR) (40°40'53``N, 124°12'4''W) near Loleta, California, USA. This
common garden consists of 26 different willow genotypes that were
collected from a single population of willows growing around Humboldt
Bay. Clonal replicates (stem cuttings) of each genotype were planted in
a completely randomized design in two hectares of a former cattle
pasture at HBNWR. Willows in our garden begin flowering in February and
reach their peak growth in early August. During this study, willows had
reached 5 - 9m in height. Further details on the genotyping and planting
of the common garden are available in (Barbour et al. (2015)).

We setup our food-web manipulation soon after galls began developing on
\emph{S. hookeriana} in early June of 2013. We manipulated gall exposure
to parasitoids across 128 plants from 8 different plant genotypes that
span the range of trait variation observed in this willow population
(Barbour et al. (2015)). On treatment plants (8 replicates per
genotype), we enclosed 14 galled leaves with organza bags (MANUFACTORER
DETAILS) to exclude three parasitoid species that attack during larva
development (hereafter larval parasitoids). This treatment did not
exclude the egg parasitoid \emph{Platygaster} sp. which attacks prior to
gall initiation (note that in Cecidomyiid midges, larva initiate gall
development CITE). On control plants, we used flagging tape to mark 14
galled leaves per plant, allowing the full suite of parasitoids to
attack \emph{Iteomyia}. Marking galls with flagging tape ensured that we
compared control and treatment galls with similar phenology when we
collected galls later in the season. In late August, we collected marked
and bagged galls from each plant. We placed galls into 30 mL vials and
kept them in the lab for 4 months at room temperature. We then opened
galls under a dissecting scope and determined whether larva survived to
pupation (our measure of fitness) or were parasitized.

We collected data on three different phenotypes for each larva that have
been shown to be important in our work (Barbour et al. (2016)) and in
other work with Cecidomyiid midges (Weis, Price, and Lynch (1983),
Heath, Abbot, and Stireman (2018)). First, we measured the size of each
gall chamber to the nearest 0.01 mm at its maximum diameter
(perpendicular to the direction of plant tissue growth). Our previous
work has shown that a larger gall diameter provides a refuge for larva
from parasitoid attack (Barbour et al. (2016)). Second, we counted the
number of chambers in each gall, which is indicative of the clutch size
of adult female midges (Weis, Price, and Lynch (1983)). All larva
collected from the same multi-chambered gall were scored with the same
clutch size. Third, we estimated gall density as the number of chambers
per 100 shoots per plant. We did this by summing the number of chambers
across all galls found on five randomly sampled branches per tree. To
account for potential differences in the number of shoots per branch for
each plant genotype (CITE other willow work), we then counted the number
of shoots on the fifth branch to estimate the number of larva per 100
shoots for each plant. All larva collected from the same plant were
scored with the same gall density phenotype. This phenotype is
indicative of the preference of female midges for particular plant
traits (hereafter `female preference'). The measurement of insect
densities on a plant is a commonly used index for measuring oviposition
preference (Gripenberg et al. (2010)), although caution must be taken in
inferring `preference' (Singer (1986)). This is because oviposition is a
complex process and the number of individuals observed on plants in the
field can be influenced by processes other than preference. While we
recognize this, since our data comes from a randomized placement of
genotypes and that we can still detect the presence of an individual
even if it dies before development (empty chamber lacking an exit hole
and without any evidence of larva or parasitoid). We cannot control for
egg predation, although this likely plays a comparatively minor role in
determining gall survival (Hawkins, Cornell, and Hochberg (1997)).

To identify the appropriate level of model complexity for testing the
effects of food-web complexity on the fitness landscapes, we compared
models using Aikaike Information Criteria. The maximal complexity we
explored was a generalized additive mixed model that fit cubic splines
to each trait as well as linear interactions between a maximum of two
traits as well as an interaction with food-web treatment. We then
examined simpler models where

This analysis enabled us to explore the appropriate complexity of the
model to include for our analyses (i.e.~include non-linear and
correlational selection gradients and whether they varied with food-web
treatment).

We used generalized additive mixed models (GAMMs, cite Bolker et al.
2008) to test the effects of food-web complexity on the shape of fitness
landscape. Larva survival (0 or 1) was our response variable and measure
of fitness. We specified our food-web treatment, each gall trait, and
all possible statistical interactions, as fixed effects to fully explore
the effects of food-web complexity on the fitness landscape. This
analysis implicitly assumes that selection is linear, which we felt was
a necessary trade-off for exploring the shape of the fitness landscape.
We specified plant genotype, plant individual nested within genotype,
and multi-chambered gall nested within plant individual, as random
effects.

To quantify selection gradients, we fit separate statistical models to
data from each food-web treatment. We used the method of Frederic J
Janzen and Hal S Stearn (1998) to calculate selection gradients and used
parametric bootstrapping to calculate their 95\% confidence intervals
(Bolker et al. (2009)).

To account for the correlated structure of our gall phenotypes
(oviposition preference at plant-level; clutch size at gall-level; gall
diameter at chamber-level), we specified gall ID nested within plant ID
nested within plant genotype as random intercepts in our statistical
models.

From these GAMMs, we estimate selection gradients by assuming the mean
value of our random effects (i.e.~setting them to zero). This was
appropriate for our analysis, since we were interested in estimating the
fitness landscape, which is function of population mean fitness and mean
trait values.

\section{Results}\label{results}

We found that more phenotypic traits were under selection in the simple
vs.~complex food web. In both complex and simple food webs, gall
diameter was under strong directional selection, with larger galls
resulting in higher larval survival (complex Beta = ; simple Beta =
)(Fig. 2A). In complex food webs, there was no evidence of selection on
clutch size (\(\beta_{clutch}=\)) or female preference
(\(\beta_{preference}=\))(orange lines in Fig. 2B,C). In simple food
webs, however, clutch size and female preference were under strong
directional selection, with smaller clutch sizes and weaker preferences
resulting in higher larval survival (blue lines in Fig. 2B,C). These
different selection pressures resulted in different adaptive landscapes
in complex vs.~simple food webs, with evidence for more rugged
landscapes in the simple rather than the complex food web (Fig. 3).
Depending on the trait combinations used to create the landscape, we
found that the ruggedness of the adaptive landscape ranged from 10\%
higher (Fig. 3A) to 274\% higher (Fig. 3C) in the simple vs.~complex
food web. Our model comparison suggested that it was unnecessary to test
for the effects of non-linear or correlational selection gradients (ref.
supp. mat.).

\section{Discussion}\label{discussion}

There are various bibliography styles available. You can select the
style of your choice in the preamble of this document. These styles are
Elsevier styles based on standard styles like Harvard and Vancouver.
Please use BibTeX~to generate your bibliography and include DOIs
whenever available.

Here are two sample references: ({\textbf{???}}; {\textbf{???}}).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Barbour2016}{}
Barbour, Matthew A, Miguel A Fortuna, Jordi Bascompte, Joshua R
Nicholson, Riitta Julkunen-Tiitto, Erik S Jules, and Gregory M
Crutsinger. 2016. ``Genetic Specificity of a Plant--insect Food Web:
Implications for Linking Genetic Variation to Network Complexity.''
\emph{Proceedings of the National Academy of Sciences} 113 (8). National
Acad Sciences: 2128--33.

\hypertarget{ref-Barbour2015}{}
Barbour, Matthew A, Mariano A Rodriguez-Cabal, Elizabeth T Wu, Riitta
Julkunen-Tiitto, Carol E Ritland, Allyson E Miscampbell, Erik S Jules,
and Gregory M Crutsinger. 2015. ``Multiple Plant Traits Shape the
Genetic Basis of Herbivore Community Assembly.'' \emph{Funct. Ecol.} 29
(8): 995--1006.

\hypertarget{ref-Bolker2009}{}
Bolker, Benjamin M, Mollie E Brooks, Connie J Clark, Shane W Geange,
John R Poulsen, M Henry H Stevens, and Jada-Simone S White. 2009.
``Generalized Linear Mixed Models: A Practical Guide for Ecology and
Evolution.'' \emph{Trends Ecol. Evol.} 24 (3): 127--35.

\hypertarget{ref-Janzen1998}{}
Frederic J Janzen and Hal S Stearn. 1998. ``Logistic Regression for
Empirical Studies of Multivariate Selection.'' \emph{Evolution} 52 (6):
1564--71.

\hypertarget{ref-Gripenberg2010}{}
Gripenberg, Sofia, Peter J Mayhew, Mark Parnell, and Tomas Roslin. 2010.
``A Meta-Analysis of Preference-Performance Relationships in
Phytophagous Insects.'' \emph{Ecol. Lett.} 13 (3): 383--93.

\hypertarget{ref-Hawkins1997}{}
Hawkins, Bradford A, Howard V Cornell, and Michael E Hochberg. 1997.
``Predators, Parasitoids, and Pathogens as Mortality Agents in
Phytophagous Insect Populations.'' \emph{Ecology} 78 (7). Ecological
Society of America: 2145--52.

\hypertarget{ref-Heath2018}{}
Heath, Jeremy J, Patrick Abbot, and John O Stireman 3rd. 2018.
``Adaptive Divergence in a Defense Symbiosis Driven from the Top down.''
\emph{Am. Nat.} 192 (1): E21--E36.

\hypertarget{ref-Singer1986}{}
Singer, Michael C. 1986. ``The Definition and Measurement of Oviposition
Preference in Plant-Feeding Insects.'' In \emph{Insect-Plant
Interactions}, edited by James R Miller and Thomas A Miller, 65--94. New
York, NY: Springer New York.

\hypertarget{ref-Weis1983}{}
Weis, Arthur E, Peter W Price, and Michael Lynch. 1983. ``Selective
Pressures on Clutch Size in the Gall Maker Asteromyia Carbonifera.''
\emph{Ecology} 64 (4). Ecological Society of America: 688--95.

\end{document}