### 2018-07-06 GRC_2018.Rmd
testing reveal.js presentation

---
title: "From ecosystems to genes: an experimental test of how network structure affects the fitness landscape"
output: 
  revealjs::revealjs_presentation:
    theme: white
    transition: fade
---

# R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```