### 2018-10-08 To_Do.Rmd
update of current analyses

---
title: "To_Do"
author: "Matt Barbour"
date: "8/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Top Priority

> Decide on which target journal to submit to.

> Output regression stats for the absence of an effect of food-web on larval survival for chamber diameter and female preference.

> Update model_4 to give separate effects of food-web complexity on larval survival.

> Translate required data and figures to main manuscript.

> Write, write, write!!!

## Lower Priority

> if I want to include host plant effects, I need to exclude chamber diameter as a phenotype under selection by the plant. including it would further confound selection on Chamber diameter since none-none larva are found in smaller galls.

> Filter to get G-matrices at the end of each simulation, then compare differences in how the selective environment alters the evolvability of those populations.

> Put statistical analyses in the supplement to show that there was no differences in phenotypes among the food-web treatments.

> Modify color of figures for illustration of simple and complex food webs and make uncompressed.