### 2018-11-22 manuscript_complexity_selection.tex
updates through sending first draft to Jordi

\documentclass[11pt,]{article}
\usepackage{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Matthew A. Barbour\^{}\{1,2,\textbackslash{}ast\}; Christopher J. Greyson-Gaito\^{}\{1,3\}; Arezoo Sootodeh\^{}\{3\}; Brendan Locke\^{}\{4\}; Jordi Bascompte\^{}\{2\}},
            pdftitle={Phenotypic evolution is more constrained in simple food webs},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Phenotypic evolution is more constrained in simple food webs}
    \author{Matthew A. Barbour\(^{1,2,\ast}\) \\ Christopher J. Greyson-Gaito\(^{1,3}\) \\ Arezoo Sootodeh\(^{3}\) \\ Brendan Locke\(^{4}\) \\ Jordi Bascompte\(^{2}\)}
    \date{}

%%% From AmNat_MS_template.tex
%\documentclass[11pt]{article}
%\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
%\usepackage[authoryear,sectionbib,sort]{natbib}
\linespread{1.7}
%\usepackage[utf8]{inputenc}
\usepackage{lineno}

%%%%%%% Sections not currently working properly
%\usepackage{titlesec}
%\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
%\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
%\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
%\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]\renewcommand{\refname}{Literature Cited}

%%% My added packages
%\usepackage{amsmath} % for matrices
%\usepackage{graphicx} % for organizing figures
%\usepackage{lmodern} % for tilde
%\usepackage[T1]{fontenc} % for tilde

%%%%%%%%% Previous - likely useful though
%\usepackage{setspace}\doublespacing
%\usepackage{float}
%\let\origfigure\figure
%\let\endorigfigure\endfigure
%\renewenvironment{figure}[1][2] {
%    \expandafter\origfigure\expandafter[H]
%} {
%    \endorigfigure
%}

\begin{document}

\maketitle

\noindent 1. University of British Columbia, Department of Zoology,
Vancouver, BC V6T 1Z4, Canada;

\noindent 2. University of Zurich, Department of Evolutionary Biology
and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich,
Switzerland;

\noindent 3. University of Guelph, Department of Integrative Biology,
Guelph, ON N1G 2W1, Canada;

\noindent 4. Humboldt State University, Department of Biological
Sciences, Arcata, CA 95521, USA.

\(^\ast\) Corresponding author; e-mail:
\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}}

\bigskip

\emph{Manuscript elements}: Figure 1, figure 2, figure 3, figure 4. All
figures should be printed in color.

\bigskip

\emph{Keywords}: fitness landscape, host-parasitoid, natural selection,
G-matrix, community context, ecological networks.

\bigskip

\emph{Manuscript type}: Article.

\bigskip

\footnotesize Prepared using an \emph{Am. Nat.} inspired \LaTeX{}
template for Rmarkdown. \normalsize

\linenumbers{} \modulolinenumbers[3]

\newpage

\section{Abstract}\label{abstract}

Adaptive landscapes describe the selective constraints acting on
phenotypic evolution, but often treat the community context of evolving
populations as a ``black box''. In contrast, ecological networks provide
an explicit representation of the community context, yet often ignore
the role of natural selection. Global change is simplifying the
structure of ecological networks; however, we are currently in a poor
position to predict how these simplified communities will affect the
evolutionary potential of populations. Here, we integrate ecological
networks and adaptive landscapes to examine how changes in food-web
complexity shape evolutionary constraints. We conducted a field
experiment that manipulated the diversity of insect parasitoids
(food-web complexity) that were able to impose selection on an insect
herbivore. We then measured herbivore survival as a function of three
key phenotypic traits. We found that more traits were under selection in
simple vs.~complex food webs. The adaptive landscape was more neutral in
complex food webs because different parasitoid species impose different
selection pressures, minimizing relative fitness differences among
phenotypes. Our results suggest that phenotypic evolution becomes more
constrained in simplified food webs. This indicates that the
simplification of ecological communities may constrain the adaptive
potential of populations to future environmental change.

\newpage

\section{Introduction}\label{introduction}

The adaptive landscape provides a powerful framework for understanding
how natural selection has shaped the evolution of biodiversity ---from
genes, to phenotypes, to species (Wright 1931; Simpson 1944; Arnold,
Pfrender, and Jones 2001). More than a metaphor, the adaptive landscape
links quantitative genetic and phenotypic variation to evolution by
natural selection (Lande 1979; Stevan J. Arnold and Wade 1984a; Stevan
J. Arnold and Wade 1984b). Ecological interactions often play a key role
in shaping adaptive landscapes, as evidenced by the role of antagonistic
(Schluter 2000; Abrams 2000) and mutualistic (Bronstein, Alarcón, and
Geber 2006) interactions in driving evolutionary change. Although there
is clear evidence that pairwise interactions can shape the adaptive
landscape, we also know that most species interact with multiple species
in a community context (Darwin 1859). Understanding how the adaptive
landscape is shaped by community context represents a major frontier for
evolutionary ecologists (McPeek 2017; terHorst et al. 2018). Resolution
on this front is urgently needed though, given the rapid impacts of
climate change on ecological communities (Scheffers et al. 2016).

Ecological networks, such as a food web describing who eats whom,
provide an explicit representation of the community context. These
networks describe the interdependency of populations within an
ecological community, providing an effective framework for predicting
how the strength and organization of ecological interactions (network
structure) affect population and community dynamics (Bascompte and
Jordano 2014; McCann 2012). Nevertheless, these networks often lack
quantitative information on how ecological interactions influence
selection on phenotypic traits. Thus, it remains unclear how network
structure shapes adaptive landscapes of interacting populations. This
represents a major barrier in our current ability to predict how changes
in community context will affect phenotypic evolution.

Here, we integrate ecological networks and adaptive landscapes to
understand how community context constrains evolutionary change (Arnold
1992). Different aspects of evolutionary constraints can be inferred by
quantifying the slope and curvature of the adaptive landscape. For
example, the slope is determined by directional selection gradients
acting on each phenotypic trait and influences the trajectory of
evolutionary change (Lande 1979; Arnold 1992). Evolutionary trajectories
become more constrained with an increase in the number of traits under
selection, as this diminishes the number of optimal solutions (Arnold
2003). The curvature of the adaptive landscape can also constrain
evolution through its indirect effect on genetic constraints (Arnold
1992; Hansen and Houle 2008). Genetic constraints are largely governed
by a population's \textbf{G}-matrix ---the additive genetic variances
and covariances between traits (Hansen and Houle 2008). In general,
genetic constraints will increase with the number of traits under
directional or stabilizing selection, as this will decrease the additive
genetic variance in those traits (Hansen and Houle 2008). Genetic
constraints may also increase with the number of trait combinations
under correlational selection, as this type of selection decreases the
evolutionary independence of traits (Hansen and Houle 2008). If we want
to predict how community context constrains evolutionary change, we must
understand how ecological networks shape the adaptive landscape.

To examine the effect of network structure on adaptive landscapes, we
conducted a field experiment that manipulated the complexity of trophic
interactions (hereafter food-web complexity) associated with an abundant
insect herbivore (\emph{Iteomyia salicisverruca}; Family
Cecidomyiidae)(fig. \ref{fig:Conceptual}). The larvae of this herbivore
induce tooth-shaped galls when they feed on the developing leaves of
willow trees (\emph{Salix} sp., Russo 2006). These galls protect larva
from attack by generalist predators (e.g.~ants, spiders), but they
suffer high mortaltiy from egg and larval parasitoids (Barbour et al.
2016). We manipulated food-web complexity by either excluding the guild
of larval parasitoids (simple food web) or allowing both egg and larval
parasitoids to impose selection on gall midge traits (complex food web;
fig. \ref{fig:Conceptual}). Food-web complexity may influence the
adaptive landscape in a number of ways. For example, if each parasitoid
imposes directional selection on a different trait, then more complex
food webs may constrain evolution by increasing the number of traits
under selection. Alternatively, if parasitoids impose selection on
different values of a trait, then their selective effects would cancel
each other out in more complex food webs. To examine these different
possibilities (among others), we applied modern statistical methods to
quantify the effect of food-web complexity on the slope and curvature of
the gall midge's adaptive landscape. Taken together, our study gives
insight to how local extinctions, and concomitant decreases in food-web
complexity, may constrain the evolution of interacting populations.

\section{Methods}\label{methods}

\subsection{Study Site}\label{study-site}

We conducted our study within a four-year old common garden experiment
of coastal willow (\emph{Salix hookeriana}) located at Humboldt Bay
National Wildlife Refuge (HBNWR) (40 40'53``N, 124 12'4''W) near Loleta,
California, USA. This common garden consists of 26 different willow
genotypes that were collected from a single population of willows
growing around Humboldt Bay. Stem cuttings of each genotype (25
replicates per genotypes) were planted in a completely randomized design
in two hectares of a former cattle pasture at HBNWR. Willows at our
study site begin flowering in February and reach their peak growth in
early August. During this study, willows had reached 5 - 9m in height.
Further details on the genotyping and planting of the common garden are
available in Barbour et al. (2015).

\subsection{Manipulating Food-web
Complexity}\label{manipulating-food-web-complexity}

We setup our food-web manipulation across 128 plants soon after galls
began developing on willows in early June of 2013. These 128 plants came
from eight different plant genotypes that spanned the range of trait
variation observed in this willow population (Barbour et al. 2015). For
the complex food-web treatment (eight replicates per genotype), we used
flagging tape to mark 14 galled leaves per plant (\textasciitilde{}30
larvae), allowing the full suite of egg and larval parasitoids to impose
selection. Marking galls with flagging tape ensured that we compared
control and treatment galls with similar phenology when we collected
galls later in the season. For the simple food-web treatment, we
enclosed 14 galled leaves with 10x15cm organza bags (ULINE, Pleasant
Prairie, WI, USA) to exclude three parasitoid species that attack during
larval development. This treatment did not exclude the egg parasitoid
\emph{Platygaster} sp., which attacks prior to gall initiation (larva
initiate gall development in Cecidomyiid midges: Gagné 1989). In late
August, we collected marked and bagged galls from each plant, placed
them into 30 mL vials and kept them in the lab for 4 months at room
temperature. We then opened galls under a dissecting scope and
determined whether larvae survived to pupation (our measure of fitness)
or were parasitized. Since we were interested in selection imposed by
parasitoids, we excluded unknown sources of mortality. For the food-web
treatment that excluded larval parasitoids (simple food web), we further
restricted our data by removing any incidental instances of parasitism
by a larval parasitoid. This represented less than 3\% of the
observations in this food-web treatment and allowed us to focus our
inferences of selection on those imposed by the egg parasitoid. Our
final dataset contains survival estimates for 1285 larvae from 613 galls
and 111 plants.

\subsection{Measuring Phenotypic
Traits}\label{measuring-phenotypic-traits}

We collected data on three different traits that we expected to
influence larval survival based on our previous work (Barbour et al.
2016) and others work with gall midges (Weis, Price, and Lynch 1983;
Heath, Abbot, and Stireman 2018). First, we measured gall diameter as
the size of each gall chamber to the nearest 0.01 mm at its maximum
diameter (perpendicular to the direction of plant tissue growth). Our
previous work has shown that larger galls are associated with higher
survival (Barbour et al. 2016). Second, we measured clutch size by
counting the number of chambers in each gall (Weis, Price, and Lynch
1983; Heath, Abbot, and Stireman 2018). All larvae collected from the
same multi-chambered gall were scored with the same clutch size. Third,
we measured oviposition (egg laying) preference as the density of larvae
observed on a plant in an independent survey. We did this randomly
sampling five branches per tree and counting the number of individual
gall chambers (number of larvae). We then converted these counts to a
measure of larval density per 100 shoots by counting the number of
shoots on the last branch we sampled. All larvae collected from the same
plant were scored with the same oviposition preference. Measuring larval
densities on plants in the field is a common method for measuring
oviposition preference (Gripenberg et al. 2010); however, caution must
be taken in inferring `preference' as larval densities can be influenced
by processes other than preference (Singer 1986). Fortunately, a couple
of features of our study system suggest that larval densities may be a
good proxy for oviposition preference. For example, since our data comes
from a randomized placement of plant genotypes in a common garden, there
is no consistent bias in which plant genotypes females are exposed to
while searching for oviposition sites. Also, egg predation is a minor
source of mortality for galling insects in general (Hawkins, Cornell,
and Hochberg 1997); therefore, we do not expect any prior egg predation
to bias our estimates of observed larval densities.

\subsection{Quantifying the Adaptive
Landscape}\label{quantifying-the-adaptive-landscape}

Our analyses consisted of three parts. First, we used generalized linear
mixed models (GLMM) to quantify selection surfaces \textemdash linear
and nonlinear relationships between absolute fitness (\(W\)) and
phenotypic traits (\(z_i\)) of individuals \textemdash in each food-web
treatment. Second, we translated selection surfaces into the scale of
relative fitness (\(w\)) in order to calculate selection gradients.
Third, we used our estimates of selection gradients to characterize the
slope and curvature of the adaptive landscape. Note that inferring
adaptive landscapes from selection surfaces assumes that trait
distributions are multivariate normal (Lande and Arnold 1983). To
approximate this assumption, we log-transformed clutch size and
square-root transformed oviposition preference. We then scaled all
phenotypic traits (mean=0 and SD=1) in order to calculate standardized
selection gradients that were comparable across traits and with other
studies of natural selection.

\textbf{Selection surface}: Since larval survival was our measure of
absolute fitness, we used a GLMM that assumed a binomial error
distribution (and logit-link function). To approximate the selection
surface, we modelled larval survival as a function of food-web treatment
as well as linear (\(\alpha_{z_i}\)), quadratic (\(\alpha_{z_i:z_i}\)),
and linear interactions (\(\alpha_{z_i:z_j}\)) between each trait. We
also allowed these trait-fitness relationships (\(\alpha\)) to vary
between food-web treatments. Note that to obtain valid estimates of
linear trait-fitness relationships, we removed nonlinear terms prior to
estimating linear relationships (Lande and Arnold 1983). Other
approaches have been advocated for approximating selection surfaces
(Schluter 1988); however, our approach enables us to calculate selection
gradients, and thus is more appropriate for approximating the adaptive
landscape (Arnold 2003). To account for the nonindependence of clutch
size (measured at gall level) and oviposition preference (measured at
plant level) as well as any independent effects of willow genotype on
larval survival, we modelled gall ID nested within plant ID nested
within genotype ID as random effects. Although statistical models with
random effects are not common in analyses of natural selection, we think
this is a mistake since modelling random effects can mitigate biased
estimates of selection due to environmental covariances between traits
and fitness (Rausher 1992). Since our end goal was to characterize the
relationship between mean trait values and mean fitness (adaptive
landscape), we assumed the mean value of our random effects
(i.e.~setting them to zero) when calculating trait-fitness
relationships. We then used parametric bootstrapping (1,000 replicates)
to estimate the effect of food-web treatment on larval survival as well
as trait-fitness relationships in each food-web treatment. To determine
whether trait-fitness relationships differed between food-web
treatments, we calculated the difference in bootstrapped replicates
between treatments.

\textbf{Selection gradients}: We used the method of Janzen and Stern
(1998) to translate trait-fitness relationships (\(\alpha\)) into the
scale of relative fitness in order to calculate directional
(\(\beta_{z_i}\)), quadratic (\(\gamma_{z_i:z_i}\)), and correlational
(\(\gamma_{z_i:z_j}\)) selection gradients. Briefly, this method
calculates the average gradient of selection surfaces by multiplying the
average of \(W(z)[1-W(z)]\) by each regression coefficient (e.g.
\(\alpha_{z_i}\), \(\alpha_{z_i:z_i}\), or \(\alpha_{z_i:z_j}\)). We
then divided this average gradient by mean fitness (\(\bar W\)) to put
it on the scale of relative fitness, and thus interpretable as a
selection gradient. We estimated selection gradients separately for each
food-web treatment. Note that we doubled all quadratic terms prior to
calculating selection gradients to put them on the same scale as
estimates of directional and correlational selection (Stinchcombe et al.
2008).

\textbf{Adaptive landscape}: We took advantage of existing theory that
translates estimates of selection gradients to the adaptive landscape
(Phillips and Arnold 1989; Arnold 2003). The shape of the adaptive
landscape can be described by both its slope and curvature, each of
which describe different aspects of constraints on phenotypic evolution.
Specifically, the slope corresponds to the column vector of directional
selection gradients, which in our study corresponds to:

\[\text{Slope} = \beta = \begin{pmatrix} \beta_{\text{Diam}} \\ \beta_{\text{Clutch}} \\ \beta_{\text{Pref}} \end{pmatrix} \]
This slope describes the selective constraints acting on the trajectory
of phenotypic evolution, which is governed by the `Lande equation',
\(\Delta \bar z=\text{G}\beta\) (Lande 1979). Quantitative predictions
of evolutionary trajectories requires knowledge of a population's
\textbf{G}-matrix; however, we can gain qualitative insight to
evolutionary constraints by counting the number of phenotypic traits
under directional selection (selective constraints, Arnold 1992).

The curvature (\textbf{C}) of the adaptive landscape is a matrix that is
determined by directional, quadratic, and correlational selection
gradients, such that
\boldmath\(\text{C}=\gamma - \beta \beta^\text{T}\), or more explicitly:

\[\textbf{C} = \begin{pmatrix} \gamma_{\text{Diam:Diam}}&& \\ \gamma_{\text{Clutch:Diam}}&\gamma_{\text{Clutch:Clutch}}& \\ \gamma_{\text{Pref:Diam}} & \gamma_{\text{Pref:Clutch}} &\gamma_{\text{Pref:Pref}} \end{pmatrix} - \begin{pmatrix} \beta_{\text{Diam}}\beta_{\text{Diam}}&& \\ \beta_{\text{Clutch}}\beta_{\text{Diam}}&\beta_{\text{Clutch}}\beta_{\text{Clutch}}& \\ \beta_{\text{Pref}}\beta_{\text{Diam}} & \beta_{\text{Pref}}\beta_{\text{Clutch}} &\beta_{\text{Pref}}\beta_{\text{Pref}} \end{pmatrix}\]
\[\textbf{C} = \begin{pmatrix} C_{\text{Diam:Diam}}&& \\ C_{\text{Clutch:Diam}} & C_{\text{Clutch:Clutch}} & \\ C_{\text{Pref:Diam}} & C_{\text{Pref:Clutch}} & C_{\text{Pref:Pref}} \end{pmatrix}\]

Note that we ommitted the upper triangle of each matrix for clarity
since it is simply the reflection of the lower triangle. This curvature
matrix describes the indirect effects of selection on genetic
constraints within a generation
(\(\Delta \text{G}=\text{G}(\gamma -\beta \beta^{\text{T}})\text{G}\)).
For example, the sign of diagonal terms of the curvature matrix dictate
whether selection will increase (+), decrease (-), or cause no change
(0) in the additive genetic variance of a trait. Similarly, any nonzero
covariance terms (off-diagonal) are indicative of selection for trait
integration (less trait independence). Therefore, we can infer the
indirect effects of selection on genetic constraints by counting the
number of negative signs along the diagonal (decrease in additive
genetic variance) and the number of nonzero terms along the off-diagonal
(trait integration) of the curvature matrix.

\subsection{Adjusting for biased measurements of
selection}\label{adjusting-for-biased-measurements-of-selection}

Rather than imposing selection, parasitoids may themselves influence the
expression of herbivore traits which could bias measurements of
selection. In our system, it was plausible that parasitoids may
influence chamber diameter by altering larval feeding behavior or
killing larvae before they complete their development. To estimate this
potential bias, we subset our data to only include galls where there was
variation in larval survival within the same gall (i.e.~1 \textgreater{}
mean survival \textgreater{} 0). If we assume that larvae within each
gall should have similar chamber diameters because they come from the
same clutch and experience the same local environment (an assumption our
data supports: gall ID explains 54\% of the variance in chamber
diameter), then the relationship between chamber diameter and larval
survival in this data subset represents the effect of parasitism on
trait expression (i.e.~bias). We used a GLMM with the same structure as
described above except that we only modelled a linear relationship
between chamber diameter and larval survival (\(\alpha_{\text{Diam}}\)).
We detected a positive bias in both food-web treatments (complex
\(\alpha_{\text{Diam}}\)= 0.36 {[}0.05, 0.67{]}; simple
\(\alpha_{\text{Diam}}\)= 0.42 {[}0.01, 0.82{]}), indicating that
unadjusted relationships would overestimate the strength of selection on
chamber diameter. To account for this bias, we substracted our mean
estimates of bias from our estimates with the full dataset prior to
calculating chamber diameter's trait-fitness relationship and
directional selection gradient.

\subsection{Measuring selection on the extended phenotype of egg
parasitoids}\label{measuring-selection-on-the-extended-phenotype-of-egg-parasitoids}

Once parasitized, the gall phenotype becomes the extended phenotype of
the egg parasitoid. This extended phenotype may influence the egg
parasitoid's survival in the face of larval parasitoids, and thus
experiences selection. Our food-web manipulation allows us to measure
selection imposed by larval parasitoids on the extended phenotype of egg
parasitoids. Using the same models as described above, we substituted
egg parasitism as our response variable to quantify trait-fitness
relationships and selection gradients acting on the egg parasitoid. Note
that we cannot test the effect of food-web complexity on the egg
parasitoid's adaptive landscape ---we can only estimate the selection
imposed by larval parasitoids. This comparison is still useful though in
determining the extent to which the community context may have indirect
evolutionary effects by altering selection on multiple interacting
populations.

All analyses and visualizations were conducted in R (R Core Team 2018).
Unless otherwise noted, we report mean estimates of trait-fitness
relationships and selection gradients with 95\% confidence intervals in
brackets. Note that for visualizing the adaptive landscape we restrict
trait axes to \(\pm 1\) SD of the mean trait value. This emphasizes the
fact that we can only reliably estimate the shape of the adaptive
landscape near the mean phenotype of the population (Arnold, Pfrender,
and Jones 2001). We also plot mean larval survival on a natural log
scale to accurately reflect the shape of the adaptive landscape (Arnold
2003).

\section{Results}\label{results}

\subsection{Food-web complexity reduces selective
constraints}\label{food-web-complexity-reduces-selective-constraints}

\indent We found that more gall midge traits experienced directional
selection in the simple (3 of 3) vs.~complex food webs (1 of 3)(table
\ref{Table:Gradients}). For example, we observed directional selection
for smaller clutch sizes in the simple food web, but there was no
evidence of selection acting on this trait in the complex food web
(table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}C). This
absence of selection appeared to be a result of conflicting selection
pressures imposed by each guild of parasitoids (fig.
\ref{fig:UV_Landscape}C). Specifically, when we subset our data to focus
on differences between parasitoid guilds, we found that larval
parasitoids actually impose directional selection for larger clutch
sizes (\(\beta_{\text{Clutch}}\)= 0.13 {[}0.03, 0.23{]}). In the simple
food web, we also observed clear evidence of directional selection for
midges to avoid ovipositing on plants with high densities of
conspecifics (table \ref{Table:Gradients}; fig.
\ref{fig:UV_Landscape}B); however, this relationship was weaker in
complex food webs (table \ref{Table:Gradients}). This was likely a
result of larval parasitoids imposing greater mortality on egg
parasitoids at high gall midge densities (see Selection on the extended
phenotype of egg parasitoids section), and thus a less than additive
effect on gall midges. Chamber diameter experienced positive directional
selection in both food-web treatments, but selection was 65\% higher in
complex food webs (table \ref{Table:Gradients}; fig.
\ref{fig:UV_Landscape}). This was not due to any difference between egg
and larval parasitoids (fig. \ref{fig:UV_Landscape}), but was due to the
lower average survival in complex food webs (contrast -0.26 {[}-0.12,
-0.42{]}).

\subsection{Food-web complexity reduces genetic
constraints}\label{food-web-complexity-reduces-genetic-constraints}

\indent The curvature of the adaptive landscape indirectly affects
genetic constraints and is influenced by directional, quadratic, and
correlational selection gradients. Our food-web treatment did not alter
correlational selection for any combination of traits (table
\ref{Table:Gradients}). Similarly, our food-web treatment did not
influence quadratic selection on either chamber diameter or clutch size
(table \ref{Table:Gradients}; fig. \ref{fig:UV_Landscape}A,C). In
contrast, our food-web treatment did alter quadratic selection acting on
oviposition preference (table \ref{Table:Gradients}). The negative
relationship between oviposition preference and larval survival dampened
at high densities in complex, but not simple food webs (fig.
\ref{fig:UV_Landscape}). This dampened relationship was partly due to a
trend for nonlinear selection imposed by larval parasitoids
(\(\gamma_{\text{Pref:Pref}}\)= 0.18 {[}-0.01, 0.42{]}), but was also
magnified by the lower average survival in complex food webs.

To estimate the net effect of food-web complexity on the curvature
(\(C\)) of the adaptive landscape, we only retained nonzero selection
gradients (i.e.~95\% CI did not overlap zero) prior to calculating the
curvature (\(\gamma - \beta \beta^\text{T}\)) in each food-web
treatment. We found that the curvatures of the adaptive landscape
exhibited the following structures:

\[\textbf{C} = \begin{pmatrix} C_{\text{Diam:Diam}}&& \\ C_{\text{Clutch:Diam}} & C_{\text{Clutch:Clutch}} & \\ C_{\text{Pref:Diam}} & C_{\text{Pref:Clutch}} & C_{\text{Pref:Pref}} \end{pmatrix}\]

\[\textbf{C}_{\text{Complex}} = \begin{pmatrix} 
-0.12 &  &  \\  
0 & 0 &  \\  
0 & 0 & 0.33 \end{pmatrix}\]

\[\textbf{C}_{\text{Simple}} = \begin{pmatrix} 
-0.04 &  &  \\  
0.02 & -0.01 &  \\  
0.03 & -0.01 & -0.03 \end{pmatrix}\]

The structure of these matrices indicate that there are more selective
constraints imposed on the gall midge in the simple (6 of 6) vs.~complex
(1 of 6) food web. Simple food webs acted to decrease genetic variance
for all three phenotypic traits (negative diagonal terms), whereas only
one trait (chamber diameter) experienced a decrease in additive genetic
variance in the complex food web. For genetic covariances, the simple
food web favored integration among all three phenotypic traits (nonzero
off-diagonal terms), and thus constraints along all three axes of
covariance (fig. \ref{fig:MV_Landscape}). In contrast, there was no
strong evidence of trait integration in the complex food web (fig.
\ref{fig:MV_Landscape}).

\subsection{Selection on the extended phenotype of egg
parasitoids}\label{selection-on-the-extended-phenotype-of-egg-parasitoids}

Gall midge preference influenced the probability of observing egg
parasitoids in complex vs.~simple food webs (contrast
\(\alpha_{\text{Pref}}\)= 1 {[}0.03, 2.22{]} ). Specifically, larval
parasitoids decreased the probability of observing egg parasitoids at
high gall midge densities (fig. \ref{fig:EggPtoid_Selection}). Note
however that this trait-fitness relationship did not translate into a
significant difference in selection (contrast \(\beta_{\text{Pref}}\)=
0.24 {[}-0.09, 0.62{]} ). This was likely because there was a tendency
(albeit not statistically signficant) for a lower probability of
egg-parasitoid survival in complex vs.~simple food webs (contrast=0.1
{[} -0.08, 0.31{]}), which masks this trait-fitness relationship. We
also observed two other interesting trends, although neither
significantly differed between treatments. Specifically, there was a
tendency for the impact of larval parasitoids to increase nonlinearly at
higher gall midge densities (fig. \ref{fig:EggPtoid_Selection}; complex
\(\alpha_{\text{Pref:Pref}}\)= -0.49 {[}-1.08, -0.1{]}; but contrast
\(\alpha_{\text{Pref:Pref}}\)= 0.46 {[}-0.03, 1.1{]}). There was also a
tendency for egg parasitoids to obtain a refuge in large gall chambers
at high gall midge densities (complex \(\alpha_{\text{Pref:Diam}}\)=
0.25 {[}0.02, 0.5{]}; but contrast \(\alpha_{\text{Pref:Diam}}\)= -0.25
{[}-0.57, 0.04{]}).

\section{Discussion}\label{discussion}

\begin{enumerate}

 \item Summary of key results
    \begin{enumerate}
    \item Gall midges experience more selective and constraints in simple food webs.
    \end{enumerate}

 \item Natural history and comparison to other studies of natural selection in galling insects.
    \begin{enumerate}
    \item \textbf{Important Studies for Comparison}: @Abrahamson1997; @Weis1983; @Heath2018; Start, Weis, and Gilbert In Press Am. Nat; @Start2016; @Start2018urbanization; Peter Price's work; Timothy Craig's work; Cheryl Briggs work.
    \item \textbf{Topics to discuss}: Persistent pattern in selection for larger galls; Contrasting effects of egg and larval parasitoids on clutch size; Nonlinear selection at high gall densities is likely reflective of the poorer searching ability of larval vs. egg parasitoids (and thus it saturates more quickly).  
    \end{enumerate}

 \item Selective constraints:
    \begin{enumerate}
    \item \textbf{Key Result}: More traits under directional selection in simple (3 of 3) vs. complex (1 of 3) food webs. Need to be wary of interpreting no selection on oviposition preference in complex food webs though. Even if there is selection on oviposition preference, this does not change our qualitative result.
    \item \textbf{Implications}: The trajectory of phenotypic evolution is less constrained in complex vs. simple food webs. In other words, there are more potential adaptive peaks in complex food webs. 
    \end{enumerate}

 \item Indirect effects of selection on genetic constraints: 
    \begin{enumerate}
    \item \textbf{Key Result}: More traits experience a decrease in additive genetic variance and more trait combinations experience selection for trait integration in simple food webs. 
    \item \textbf{Implications}: The structure of the G-matrix is under less influence of selection in complex vs. simple food webs. Thus the trajectory of phenotypic evolution may be dictated more by genetic rather than selective constraints. This is in line with empirical evidence (from complex natural systems) that evolution tends to proceed along genetic lines of least resistance [@Schluter1996].
    \end{enumerate}

 \item Indirect evolutionary effects:
    \begin{enumerate}
    \item \textbf{Key Result}: Evidence of selection on the extended phenotype of egg parasitoids.
    \item \textbf{Implications}: Hints at a dynamic adaptive landscape and potential indirect evolutionary effects in complex food webs.
    \end{enumerate}

 \item Caveats: 
    \begin{enumerate} 
    \item \textbf{Key Caveat}: Experimental artifact of lower larval survival in complex food-web treatment.
    \item \textbf{Implications}: Artificially increased the magnitude of selection in complex food webs. Note however, that this does not qualitatively change any of our results. In fact, we are actually underestimating the constraints imposed by our simple food-web treatment. This is because the egg parasitoids would be released from their intraguild predator, and thus could increase in abundance over future generations and impose stronger selection on gall midges. 
    \end{enumerate}

\item Conclusions

\end{enumerate}

\section{Acknowledgements}\label{acknowledgements}

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Abrams2000}{}
Abrams, Peter A. 2000. ``The Evolution of Predator-Prey Interactions:
Theory and Evidence.'' \emph{Annual Review of Ecology and Systematics}
31 (1): 79--105.
doi:\href{https://doi.org/10.1146/annurev.ecolsys.31.1.79}{10.1146/annurev.ecolsys.31.1.79}.

\hypertarget{ref-Arnold1992}{}
Arnold, Stevan J. 1992. ``Constraints on Phenotypic Evolution.''
\emph{The American Naturalist} 140. University of Chicago Press:
S85--S107.

\hypertarget{ref-Arnold2003}{}
Arnold, Stevan J. 2003. ``Performance Surfaces and Adaptive
Landscapes1.'' \emph{Integrative and Comparative Biology} 43 (3):
367--75.
doi:\href{https://doi.org/10.1093/icb/43.3.367}{10.1093/icb/43.3.367}.

\hypertarget{ref-Arnold1984applications}{}
Arnold, Stevan J., and Michael J. Wade. 1984a. ``On the Measurement of
Natural and Sexual Selection: Applications.'' \emph{Evolution} 38 (4):
720--34. \url{http://www.jstor.org/stable/2408384}.

\hypertarget{ref-Arnold1984theory}{}
---------. 1984b. ``On the Measurement of Natural and Sexual Selection:
Theory.'' \emph{Evolution} 38 (4): 709--19.
\url{http://www.jstor.org/stable/2408383}.

\hypertarget{ref-Arnold2001}{}
Arnold, Stevan J., Michael E. Pfrender, and Adam G. Jones. 2001. ``The
Adaptive Landscape as a Conceptual Bridge Between Micro- and
Macroevolution.'' \emph{Genetica} 112 (1): 9--32.
doi:\href{https://doi.org/10.1023/A:1013373907708}{10.1023/A:1013373907708}.

\hypertarget{ref-Barbour2016}{}
Barbour, Matthew A., Miguel A. Fortuna, Jordi Bascompte, Joshua R.
Nicholson, Riitta Julkunen-Tiitto, Erik S. Jules, and Gregory M.
Crutsinger. 2016. ``Genetic Specificity of a Plantinsect Food Web:
Implications for Linking Genetic Variation to Network Complexity.''
\emph{Proceedings of the National Academy of Sciences} 113 (8). National
Academy of Sciences: 2128--33.
doi:\href{https://doi.org/10.1073/pnas.1513633113}{10.1073/pnas.1513633113}.

\hypertarget{ref-Barbour2015}{}
Barbour, Matthew A., Mariano A. Rodriguez-Cabal, Elizabeth T. Wu, Riitta
Julkunen-Tiitto, Carol E. Ritland, Allyson E. Miscampbell, Erik S.
Jules, and Gregory M. Crutsinger. 2015. ``Multiple Plant Traits Shape
the Genetic Basis of Herbivore Community Assembly.'' \emph{Functional
Ecology} 29 (8): 995--1006.
doi:\href{https://doi.org/10.1111/1365-2435.12409}{10.1111/1365-2435.12409}.

\hypertarget{ref-Bascompte2014}{}
Bascompte, Jordi, and Pedri Jordano. 2014. \emph{Mutualistic Networks}.
Princeton University Press.
\url{http://www.jstor.org/stable/j.ctt5hhnpq}.

\hypertarget{ref-Bronstein2006}{}
Bronstein, Judith L., Ruben Alarcón, and Monica Geber. 2006. ``The
Evolution of Plant--insect Mutualisms.'' \emph{New Phytologist} 172 (3):
412--28.
doi:\href{https://doi.org/10.1111/j.1469-8137.2006.01864.x}{10.1111/j.1469-8137.2006.01864.x}.

\hypertarget{ref-Darwin1859}{}
Darwin, Charles. 1859. \emph{On the Origin of Species by Means of
Natural Selection}. London: Murray.

\hypertarget{ref-Gagne1989}{}
Gagné, R J. 1989. ``The Plant-Feeding Gall Midges of North America.''
agris.fao.org.

\hypertarget{ref-Gripenberg2010}{}
Gripenberg, Sofia, Peter J. Mayhew, Mark Parnell, and Tomas Roslin.
2010. ``A Meta-Analysis of Preference--performance Relationships in
Phytophagous Insects.'' \emph{Ecology Letters} 13 (3): 383--93.
doi:\href{https://doi.org/10.1111/j.1461-0248.2009.01433.x}{10.1111/j.1461-0248.2009.01433.x}.

\hypertarget{ref-Hansen2008}{}
Hansen, T F, and D Houle. 2008. ``Measuring and Comparing Evolvability
and Constraint in Multivariate Characters.'' \emph{J. Evol. Biol.} 21
(5): 1201--19.

\hypertarget{ref-Hawkins1997}{}
Hawkins, Bradford A., Howard V. Cornell, and Michael E. Hochberg. 1997.
``Predators, Parasitoids, and Pathogens as Mortality Agents in
Phytophagous Insect Populations.'' \emph{Ecology} 78 (7): 2145--52.
doi:\href{https://doi.org/10.1890/0012-9658(1997)078\%5B2145:PPAPAM\%5D2.0.CO;2}{10.1890/0012-9658(1997)078{[}2145:PPAPAM{]}2.0.CO;2}.

\hypertarget{ref-Heath2018}{}
Heath, Jeremy J., Patrick Abbot, and John O. Stireman. 2018. ``Adaptive
Divergence in a Defense Symbiosis Driven from the Top down.'' \emph{The
American Naturalist} 192 (1): E21--E36.
doi:\href{https://doi.org/10.1086/697446}{10.1086/697446}.

\hypertarget{ref-Janzen1998}{}
Janzen, Fredric J., and Hal S. Stern. 1998. ``Logistic Regression for
Empirical Studies of Multivariate Selection.'' \emph{Evolution} 52 (6):
1564--71.
doi:\href{https://doi.org/10.1111/j.1558-5646.1998.tb02237.x}{10.1111/j.1558-5646.1998.tb02237.x}.

\hypertarget{ref-Lande1979}{}
Lande, Russell. 1979. ``Quantitative Genetic Analysis of Multivariate
Evolution, Applied to Brain:body Size Allometry.'' \emph{Evolution} 33
(1Part2): 402--16.
doi:\href{https://doi.org/10.1111/j.1558-5646.1979.tb04694.x}{10.1111/j.1558-5646.1979.tb04694.x}.

\hypertarget{ref-Lande1983}{}
Lande, Russell, and Stevan J. Arnold. 1983. ``The Measurement of
Selection on Correlated Characters.'' \emph{Evolution} 37 (6): 1210--26.
\url{http://www.jstor.org/stable/2408842}.

\hypertarget{ref-McCann2012}{}
McCann, Kevin S. 2012. \emph{Food Webs}. Princeton University Press.
\url{http://www.jstor.org/stable/j.ctt7rr0s}.

\hypertarget{ref-McPeek2017}{}
McPeek, Mark A. 2017. \emph{Evolutionary Community Ecology}. Princeton
University Press. \url{http://www.jstor.org/stable/j.ctt1p0vjqn}.

\hypertarget{ref-Phillips1989}{}
Phillips, Patrick C., and Stevan J. Arnold. 1989. ``Visualizing
Multivariate Selection.'' \emph{Evolution} 43 (6): 1209--22.
doi:\href{https://doi.org/10.1111/j.1558-5646.1989.tb02569.x}{10.1111/j.1558-5646.1989.tb02569.x}.

\hypertarget{ref-R2018}{}
R Core Team. 2018. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\hypertarget{ref-Rausher1992}{}
Rausher, Mark D. 1992. ``THE Measurement of Selection on Quantitative
Traits: BIASES Due to Environmental Covariances Between Traits and
Fitness.'' \emph{Evolution} 46 (3): 616--26.
doi:\href{https://doi.org/10.1111/j.1558-5646.1992.tb02070.x}{10.1111/j.1558-5646.1992.tb02070.x}.

\hypertarget{ref-Russo2006}{}
Russo, Ron. 2006. \emph{Field Guide to Plant Galls of California and
Other Western States}. 1st ed. University of California Press.
\url{http://www.jstor.org/stable/10.1525/j.ctt1pnc33}.

\hypertarget{ref-Scheffers2016}{}
Scheffers, Brett R., Luc De Meester, Tom C. L. Bridge, Ary A. Hoffmann,
John M. Pandolfi, Richard T. Corlett, Stuart H. M. Butchart, et al.
2016. ``The Broad Footprint of Climate Change from Genes to Biomes to
People.'' \emph{Science} 354 (6313). American Association for the
Advancement of Science.
doi:\href{https://doi.org/10.1126/science.aaf7671}{10.1126/science.aaf7671}.

\hypertarget{ref-Schluter1988}{}
Schluter, Dolph. 1988. ``ESTIMATING the Form of Natural Selection on a
Quantitative Trait.'' \emph{Evolution} 42 (5): 849--61.
doi:\href{https://doi.org/10.1111/j.1558-5646.1988.tb02507.x}{10.1111/j.1558-5646.1988.tb02507.x}.

\hypertarget{ref-Schluter2000}{}
---------. 2000. ``Ecological Character Displacement in Adaptive
Radiation.'' \emph{The American Naturalist} 156 (S4): S4--S16.
\url{http://www.jstor.org/stable/10.1086/303412}.

\hypertarget{ref-Simpson1944}{}
Simpson, George Gaylord. 1944. \emph{Tempo and Mode in Evolution}.
Columbia University Press.

\hypertarget{ref-Singer1986}{}
Singer, Michael C. 1986. ``The Definition and Measurement of Oviposition
Preference in Plant-Feeding Insects.'' In \emph{Insect-Plant
Interactions}, edited by James R. Miller and Thomas A. Miller, 65--94.
New York, NY: Springer New York.
doi:\href{https://doi.org/10.1007/978-1-4612-4910-8_3}{10.1007/978-1-4612-4910-8\_3}.

\hypertarget{ref-Stinchcombe2008}{}
Stinchcombe, John R, Aneil F Agrawal, Paul A Hohenlohe, Stevan J Arnold,
and Mark W Blows. 2008. ``Estimating Nonlinear Selection Gradients Using
Quadratic Regression Coefficients: Double or Nothing?'' \emph{Evolution}
62 (9): 2435--40.

\hypertarget{ref-terHorst2018}{}
terHorst, Casey P, Peter C Zee, Katy D Heath, Thomas E Miller, Abigail I
Pastore, Swati Patel, Sebastian J Schreiber, Michael J Wade, and Matthew
R Walsh. 2018. ``Evolution in a Community Context: Trait Responses to
Multiple Species Interactions*.'' \emph{Am. Nat.}, January. University
of Chicago PressChicago, IL.

\hypertarget{ref-Weis1983}{}
Weis, Arthur E., Peter W. Price, and Michael Lynch. 1983. ``Selective
Pressures on Clutch Size in the Gall Maker Asteromyia Carbonifera.''
\emph{Ecology} 64 (4): 688--95.
\url{http://www.jstor.org/stable/1937190}.

\hypertarget{ref-Wright1931}{}
Wright, Sewall. 1931. ``EVOLUTION in Mendelian Populations.''
\emph{Genetics} 16 (2). Genetics: 97--159.
\url{http://www.genetics.org/content/16/2/97}.

\newpage

\section{Tables}\label{tables}

\begin{table}[h]
\caption{Standardized selection gradients acting on gall midges in complex vs. simple food webs.}
\label{Table:Gradients}
\centering
\begin{tabular}{lccc}
\\ 
\hline
\textbf{Selection gradient} & \textbf{Complex} & \textbf{Simple} & \textbf{Contrast}  \\ 
\hline
$\beta_{\text{Diam}}$ & 
\textbf{
0.34 [
0.22,
0.47] }& 
\textbf{
0.21 [
0.13,
0.3] }&

-0.13 [
-0.26,
0.01] \\

$\beta_{\text{Clutch}}$ & 
0.06 [
-0.04,
0.17] & 
\textbf{
-0.09 [
-0.18,
-0.02] }& 
\textbf{
-0.15 [
-0.29,
-0.02] }\\

$\beta_{\text{Pref}}$ &
-0.12 [
-0.29,
0.05] & 
\textbf{
-0.16 [
-0.25,
-0.07] }&

-0.04 [
-0.22,
0.14] \\

$\gamma_{\text{Diam,Diam}}$ &
0.14 [
-0.06,
0.32] &

0.1 [
-0.03,
0.25] &

-0.04 [
-0.27,
0.2] \\

$\gamma_{\text{Clutch,Clutch}}$ & 
-0.05 [
-0.25,
0.16] &

-0.12 [
-0.27,
0.03] &

-0.06 [
-0.33,
0.19] \\

$\gamma_{\text{Pref,Pref}}$ & 
\textbf{
0.33 [
0.06,
0.64] }&

0.01 [
-0.14,
0.17] & 
\textbf{
-0.32 [
-0.66,
0] }\\

$\gamma_{\text{Diam,Clutch}}$ & 
-0.05 [
-0.17,
0.08] &

-0.07 [
-0.15,
0.02] &

-0.02 [
-0.16,
0.13] \\

$\gamma_{\text{Diam,Pref}}$ & 
-0.13 [
-0.28,
0] &

-0.02 [
-0.09,
0.06] &

0.12 [
-0.05,
0.29] \\

$\gamma_{\text{Clutch,Pref}}$ & 
0.03 [
-0.11,
0.18] &

0 [
-0.07,
0.07] &

-0.03 [
-0.19,
0.12] \\ 
\hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Values in brackets represent 95\% confidence intervals. Bold values indicate that the 95\% CI does not overlap zero. $\beta_{\text{Diam}}$ has been adjusted for bias.}
\end{table}

\newpage

\begin{table}[h]
\caption{Relationship between absolute fitness (larval survival) and phenotypic traits of gall midges in complex vs. simple food webs.}
\label{Table:Coefs}
\centering
\begin{tabular}{lccc}
\\ 
\hline
\textbf{Coefficient} & \textbf{Complex} & \textbf{Simple} & \textbf{Contrast}  \\ 
\hline
$\alpha_{\text{Diam}}$ & 
\textbf{
1.14 [
0.73,
1.57] }& 
\textbf{
1.1 [
0.68,
1.58] }&

-0.04 [
-0.57,
0.51] \\

$\alpha_{\text{Clutch}}$ & 
0.21 [
-0.13,
0.57] & 
\textbf{
-0.47 [
-0.94,
-0.09] }& 
\textbf{
-0.68 [
-1.27,
-0.13] }\\

$\alpha_{\text{Pref}}$ &
-0.41 [
-0.97,
0.15] & 
\textbf{
-0.84 [
-1.35,
-0.36] }&

-0.43 [
-1.16,
0.25] \\

$\alpha_{\text{Diam,Diam}}$ &
0.23 [
-0.1,
0.54] &

0.27 [
-0.07,
0.65] &

0.04 [
-0.43,
0.52] \\

$\alpha_{\text{Clutch,Clutch}}$ & 
-0.09 [
-0.42,
0.26] &

-0.31 [
-0.71,
0.08] &

-0.22 [
-0.77,
0.31] \\

$\alpha_{\text{Pref,Pref}}$ & 
\textbf{
0.56 [
0.1,
1.07] }&

0.03 [
-0.38,
0.46] &

-0.52 [
-1.18,
0.09] \\

$\alpha_{\text{Diam,Clutch}}$ & 
-0.16 [
-0.57,
0.25] &

-0.35 [
-0.78,
0.08] &

-0.19 [
-0.78,
0.44] \\

$\alpha_{\text{Diam,Pref}}$ & 
-0.44 [
-0.95,
0.01] &

-0.09 [
-0.5,
0.3] &

0.36 [
-0.27,
0.97] \\

$\alpha_{\text{Clutch,Pref}}$ & 
0.1 [
-0.37,
0.59] &

0 [
-0.37,
0.38] &

-0.1 [
-0.69,
0.49] \\ 
\hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Values in brackets represent 95\% confidence intervals. Bold values indicate that the 95\% CI does not overlap zero. $\alpha_{\text{Diam}}$ has been adjusted for bias.}
\end{table}

\newpage

\section{Figure Legends}\label{figure-legends}

\textbf{Figure \ref{fig:Conceptual}}: Experimental manipulation of
food-web complexity associated with a leaf-galling midge (\emph{Iteomyia
salicisverruca}) feeding on the willow \emph{Salix hookeriana}. In the
complex food-web treatment (A), we allowed the full suite of egg (blue
rectangle) and larval (yellow rectangle) parasitoids to impose
selection. In the simple food-web treatment (B), we used mesh bags to
exclude the guild of larval parasitoids, only allowing the egg
parasitoid (\emph{Platygaster} sp.) to impose selection. Black arrows
denote the flow of energy in this network of trophic interactions.
Larval parasitoids include the following species (from left to right):
\emph{Mesopolobus} sp. (Family: Pteromalidae); \emph{Tetrastichus} sp.
(Family: Eulophidae); and \emph{Torymus} sp. (Family: Torymidae).

\textbf{Figure \ref{fig:UV_Landscape}}: Adaptive landscape of gall midge
phenotypes in complex vs.~simple food webs. Each panel corresponds to a
different phenotypic trait: chamber diameter (A); oviposition preference
(B); and clutch size (C). Solid lines represent selection experienced in
complex (orange) and simple (blue) food webs. Transparent lines
represent bootstrapped replicates to show the uncertainty in selection.
For clarity, we only display 100 bootstraps even though inferences are
based on 1,000 replicates. Solid yellow lines represent selection
imposed by larval parasitoids in the complex food web. Note that mean
larval survival is plotted on a natural log scale to accurately reflect
the adaptive landscape.

\textbf{Figure \ref{fig:MV_Landscape}}: Two dimensional view of adaptive
landscapes of gall midge phenotypes in complex vs.~simple food webs.
Each panel corresponds to a different combination of phenotypic traits:
clutch size and chamber diameter (A); clutch size and oviposition
preference (B); oviposition preference and chamber diameter (C).
Selection favors trait integration for each combination of traits in
simple food webs, whereas there is no strong evidence for trait
integration in complex food webs. Note that mean larval survival is
plotted on a natural log scale to accurately reflect the adaptive
landscape.

\textbf{Figure \ref{fig:EggPtoid_Selection}}: Selection imposed by
larval parasitoids on the extended phenotype of egg parasitoids
(\emph{Platygaster} sp.). The solid line represents the average
difference in the probability of observing the egg parasitoid in complex
vs.~simple food webs as a function of gall midge oviposition preference.
Transparent lines represent bootstrapped replicates to show the
uncertainty in selection. For clarity, we only display 100 bootstraps
even though inferences are based on 1,000 replicates. The decrease in
the probability of observing egg parasitoids at high gall-midge
densities indicates the increasing impact of larval parasitoids.

\begin{figure}
\centering
\includegraphics{analyses/complex_simple_foodwebs.jpeg}
\caption{\label{fig:Conceptual}Experimental manipulation of food-web
complexity associated with a leaf-galling midge (\emph{Iteomyia
salicisverruca}) feeding on the willow \emph{Salix hookeriana}. In the
complex food-web treatment (A), we allowed the full suite of egg (blue
rectangle) and larval (yellow rectangle) parasitoids to impose
selection. In the simple food-web treatment (B), we used mesh bags to
exclude the guild of larval parasitoids, only allowing the egg
parasitoid (\emph{Platygaster} sp.) to impose selection. Black arrows
denote the flow of energy in this network of trophic interactions.
Larval parasitoids include the following species (from left to right):
\emph{Mesopolobus} sp. (Family: Pteromalidae); \emph{Tetrastichus} sp.
(Family: Eulophidae); and \emph{Torymus} sp. (Family: Torymidae).}
\end{figure}

\begin{figure}
\centering
\includegraphics{analyses/UV_landscapes.pdf}
\caption{\label{fig:UV_Landscape}}
\end{figure}

\begin{figure}
\centering
\includegraphics{analyses/MV_landscapes.pdf}
\caption{\label{fig:MV_Landscape}Two dimensional view of adaptive
landscapes of gall midge phenotypes in complex vs.~simple food webs.
Each panel corresponds to a different combination of phenotypic traits:
clutch size and chamber diameter (A); clutch size and oviposition
preference (B); oviposition preference and chamber diameter (C).
Selection favors trait integration for each combination of traits in
simple food webs, whereas there is no strong evidence for trait
integration in complex food webs. Note that mean larval survival is
plotted on a natural log scale to accurately reflect the adaptive
landscape.}
\end{figure}

\begin{figure}
\centering
\includegraphics{analyses/selection_on_Platygaster.pdf}
\caption{\label{fig:EggPtoid_Selection}Selection imposed by larval
parasitoids on the extended phenotype of egg parasitoids
(\emph{Platygaster} sp.). The solid line represents the average
difference in the probability of observing the egg parasitoid in complex
vs.~simple food webs as a function of gall midge oviposition preference.
Transparent lines represent bootstrapped replicates to show the
uncertainty in selection. For clarity, we only display 100 bootstraps
even though inferences are based on 1,000 replicates. The decrease in
the probability of observing egg parasitoids at high gall-midge
densities indicate that larval parasitoids impose nonlinear selection on
the extended phenotype of egg parasitoids.}
\end{figure}

\end{document}