### 2019-01-16 finalize_manuscript_dataset.Rmd
wrote script to finalize dataset for manuscript.

---
title: "Finalize dataset for *Phenotypic evolution is more constrained in simpler food webs*"
author: "Matthew A. Barbour"
date: "1/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)  # for managing data
```

Focus cleaned dataset on galls marked for the experiment (phenology == "early") and that were collected while they were still on the tree (rather than bags that fell to the ground). Subsetting galls that remained on trees removes potential bias between treatments, because we were not able to recover galls that had fallen from the tree in the Control treatment (they only had flagging tape placed nearby rather than being enclosed within a bag).

```{r Filter dataset}
gall_selection.df <- read_csv("data/raw_data_management/gall_selection_data.csv") %>%
  filter(phenology == "early", Location == "tree") %>%
  select(Treatment = Treatment.focus, 
         # unique identifying information at the plant- and gall-level
         Genotype, # plant genotype ID
         Plant_Position, # unique plant ID (also corresponds to position within the common garden experiment)
         Gall_Number, # unique multi-chambered gall ID
         Gall_Letter, # unique chamber ID within gall number
         # rename traits and add measurement units to variable
         chamber.diameter_mm = Gall_Height_mm, # chamber-level
         oviposition.preference_individuals.per.100.shoots = Density_per_100_shoots, # plant-level
         clutch.size_individuals = gall_individuals, # gall-level
         pupa, # successful pupation of larval Iteomyia salicisverruca
         egg.ptoid = platy, # reared egg parasitoid, Platygaster sp. from gall chamber
         larval.ptoid = ectos, # reared larval parasitoid (most of which were ectoparasitoids) from gall chamber
         early.larval.death = none.none) # early larval death = no larva or exit hole associated with gall chamber

write_csv(gall_selection.df, "data/gall_selection_manuscript_dataset.csv")
```