### 2019-05-21 analyses/reproduce_analyses.Rmd
update analyses folder.

---
title: "Consumer extinctions constrain phenotypic evolution in the resulting food web"
subtitle: "Supplementary information"
author: "Code author: Matthew A. Barbour"
date: "`r Sys.Date()`"
output: 
  tufte::tufte_html:
    toc: true
---

\

This document presents a detailed overview of the analyses presented in the main text. It highlights the essential R code used to conduct these analyses, but does not display code related to table and plot aesthetics. The underlying code for this document is provided in **reproduce_analyses.Rmd** and is available at https://github.com/mabarbour/complexity_selection.

\

```{r setup, include=FALSE}
# Suppress messages, warnings, and errors for document aesthetics.
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE, echo=FALSE, dev="png")
```

```{r required libraries, results='hide'}
# Load required libraries.
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(latex2exp)  # using Latex notation for axes labels
library(stringr)
library(MVN)        # assessing multivariate normality assumption
```

```{r custom functions}
## Write custom functions to avoid repetitive code.

# Transform logit to probability
inverse_logit <- function(x) exp(x)/(1+exp(x))

# Get confidence intervals from bootstrapped samples
conf.high <- function(x) quantile(x, probs = 0.975)
conf.low <- function(x) quantile(x, probs = 0.025)

# Get data for plotting bootstrapped estimates in Figures 2, 3, and 4 of main text
bootstrap_fitness <- function(logistic_model, newdata, bootstraps, intervals=c(0.025,0.975)){

  # Absolute fitness
  get_absolute_fitness <- function(.) predict(., newdata=newdata, type="response", re.form=~0)

  if(is.null(bootstraps) == FALSE){

    get_bootstraps_absolute_fitness <-bootMer(logistic_model, FUN = get_absolute_fitness, nsim=bootstraps, parallel="multicore", ncpus=32, seed=34)

   get_intervals_absolute_fitness <- apply(get_bootstraps_absolute_fitness$t, 2, function(x) x[order(x)][c(round(bootstraps*intervals[1],0),round(bootstraps*intervals[2],0))])

  absolute_fitness_df <- data.frame(newdata, 
                    average = get_absolute_fitness(logistic_model), 
                    lower = get_intervals_absolute_fitness[1, ],
                    upper = get_intervals_absolute_fitness[2, ],
                    t(get_bootstraps_absolute_fitness$t))
  } else {
    absolute_fitness_df <- data.frame(newdata, 
                    average = get_absolute_fitness(logistic_model))
  }

  # Relative fitness
  get_relative_fitness <- function(.) get_absolute_fitness(.)/mean(get_absolute_fitness(.))

  if(is.null(bootstraps) == FALSE){

    get_bootstraps_relative_fitness <-bootMer(logistic_model, FUN = get_relative_fitness, nsim=bootstraps, parallel="multicore", ncpus=32, seed=34)

  get_intervals_relative_fitness <- apply(get_bootstraps_relative_fitness$t, 2, function(x) x[order(x)][c(round(bootstraps*intervals[1],0),round(bootstraps*intervals[2],0))])

  relative_fitness_df <- data.frame(newdata, 
                    average = get_relative_fitness(logistic_model), 
                    lower = get_intervals_relative_fitness[1, ],
                    upper = get_intervals_relative_fitness[2, ],
                    t(get_bootstraps_relative_fitness$t))
  } else {
      relative_fitness_df <- data.frame(newdata, 
                    average = get_relative_fitness(logistic_model))
  }

  # Organize data
  return(list(absolute_fitness = absolute_fitness_df, relative_fitness = relative_fitness_df))
}
```

```{r color palette}
# Color-blind friendly palette: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/; 
# order: orange, light blue, green, yellow, dark blue, red, pink
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000", "#999999")
original_color <- cbPalette[6] 
extinction_color <- cbPalette[5] 
treatment_colors <- c(original_color, extinction_color)
```

## Summary of dataset

```{r read and tidy data}
## Read and tidy data for analyses.

full_df <- read_csv("../data/gall_selection_manuscript_dataset.csv") %>%
  mutate(gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)), # there were a couple of times where two pupa were found in a single chamber. I set these to 1 to permit an analysis with a binomial model
         egg.ptoid = as.numeric(ifelse(egg.ptoid > 0, 1, 0)), # there were a couple occassions where two egg parasitoids were found in a single chamber. I set these to 1 to permit an analysis with a binomial model.
         unknown_mortality = ifelse(rowSums(select(., pupa:early.larval.death)) < 1, 1, 0))

gall_selection.df <- full_df %>%
  # convert appropriate variables to characters instead of integers
  mutate(Plant_Position = as.character(Plant_Position),
         Gall_Number = as.character(Gall_Number)) %>%
  unite(Gall_ID, Gall_Number, Gall_Letter, remove = FALSE) %>%

  # eliminate unknown sources of mortality
  filter(egg.ptoid > 0 | larval.ptoid > 0 | pupa > 0) %>% 
  # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  filter(Treatment == "Ectoparasitoid exclusion" & larval.ptoid < 1 | Treatment == "Control") %>% 
  mutate(Foodweb = ifelse(Treatment=="Control", "Original", "Extinction"), # rename Treatment to match presentation in manuscript
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg.ptoid = as.numeric(ifelse(egg.ptoid > 0, 1, 0)), # there were a couple occassions where two egg parasitoids were found in a single chamber. I set these to 1 to permit an analysis with a binomial model.

         # transform and scale traits to approximate multivariate normal assumption
         sc.Diam = as.numeric(scale(chamber.diameter_mm)), 
         sc.log.Clutch = as.numeric(scale(log(clutch.size_individuals))), 
         sc.sqrt.Pref = as.numeric(scale(sqrt(oviposition.preference_individuals.per.100.shoots)))) 
```

Below is a summary of the raw trait data for the gall midge, *Iteomyia salicisverruca*. Gall chamber diameter (**Diam**) is measured in mm, clutch size (**Clutch**) is measured as the number of chambers per gall, and oviposition preference (**Pref**) is measured as the number of chambers per 100 shoots on the plant. Note that each chamber contains one individual larva.

```{r raw trait summary}
raw_means <- gall_selection.df %>%
  transmute(Diam = chamber.diameter_mm, 
            Clutch = clutch.size_individuals, 
            Pref = oviposition.preference_individuals.per.100.shoots) %>%
  summarise_all(funs(mean))

raw_SDs <- gall_selection.df %>%
  transmute(Diam = chamber.diameter_mm, 
            Clutch = clutch.size_individuals, 
            Pref = oviposition.preference_individuals.per.100.shoots) %>%
  summarise_all(funs(sd))

data.frame(Trait = c("Diam", "Clutch", "Pref"),
           Mean = round(c(raw_means$Diam, raw_means$Clutch, raw_means$Pref), 1),
           SD = round(c(raw_SDs$Diam, raw_SDs$Clutch, raw_SDs$Pref), 1)) %>%
  knitr::kable(., caption = "Mean and SD of phenotypic traits.")
```

Clutch size and oviposition preference were not normally distributed, so we log and square-root transformed these traits, respectively.

```{r transformed trait summary}
trans_means <- gall_selection.df %>%
  transmute(log.Clutch = log(clutch.size_individuals), 
            sqrt.Pref = sqrt(oviposition.preference_individuals.per.100.shoots)) %>%
  summarise_all(funs(mean))

trans_SDs <- gall_selection.df %>%
  transmute(log.Clutch = log(clutch.size_individuals), 
            sqrt.Pref = sqrt(oviposition.preference_individuals.per.100.shoots)) %>%
  summarise_all(funs(sd))

data.frame(Trait = c("log.Clutch", "sqrt.Pref"),
           Mean = round(c(trans_means$log.Clutch, trans_means$sqrt.Pref), 1),
           SD = round(c(trans_SDs$log.Clutch, trans_SDs$sqrt.Pref), 1)) %>%
  knitr::kable(., caption = "Mean and SD of transformed phenotypic traits.")
```

For our subsequent analyses, we scaled each phenotypic trait (mean = 0, SD = 1) so that estimates of selection surfaces and gradients were comparable between traits and with traits in other studies. Below, we show a random subset of the data used in our subsequent analyses.

```{r illustrate data used in analyses}
set.seed(34)
sample_n(gall_selection.df, size = 10) %>%
  select(Foodweb, Genotype, Plant_Position, Gall_Number, Gall_ID, gall_survival, sc.Diam, sc.log.Clutch, sc.sqrt.Pref) %>%
  mutate(sc.Diam = round(sc.Diam, 2), sc.log.Clutch = round(sc.log.Clutch, 2), sc.sqrt.Pref = round(sc.sqrt.Pref, 2)) %>%
  knitr::kable(., caption = "Random sample of 10 data points used for our analyses.")
```

- **Foodweb** = food-web treatment (Original or Extinction)
- **Genotype** = plant genotype ID (n = `r length(unique(gall_selection.df$Genotype))`)
- **Plant_Position** = plant ID (n = `r length(unique(gall_selection.df$Plant_Position))`)
- **Gall_Number** = multi-chambered gall ID (n = `r length(unique(gall_selection.df$Gall_Number))`)
- **Gall_ID** = individual chamber ID (n = `r length(unique(gall_selection.df$Gall_ID))`)
- **gall_survival** = measure of larval fitness (0 = dead; 1 = survived to pupation)
- **sc.Diam** = scaled chamber diameter
- **sc.log.Clutch** = scaled logarithm of clutch size
- **sc.sqrt.Pref** = scaled square-root of oviposition preference

\

## Multivariate normality of phenotypic traits

We used graphical checks to evaluate whether our transformations resulted in traits with a multivariate normal distribution. The histograms below show that our transformations resulted in approximately normal distributions for each phenotypic trait.

```{r Univariate_histograms,  fig.cap="Histograms of each phenotypic trait after transformation. The red line illustrates a normal distribution."}
mvn_univariate_hist <- mvn(select(gall_selection.df, sc.Diam, sc.log.Clutch, sc.sqrt.Pref), univariatePlot = "histogram")
```

Note also that in the multivariate quantile-quantile (Q-Q) plot, most points fall along the expected line, suggesting that our transformations provide a reasonable approximation of a multivariate normal distribution.

```{r Multivariate_QQ, fig.cap="Multivariate quantile-quantile (Q-Q) plot to assess deviations from multivariate normality (black line)."}
mvn_multivariate_qq <- mvn(select(gall_selection.df, sc.Diam, sc.log.Clutch, sc.sqrt.Pref), multivariatePlot = "qq")  
```

\

## Selection surfaces for the gall midge: food-web treatments

We wrote the generalized linear mixed-effects model (GLMM) in a way that independently estimates the effect of food-web treatment, each trait, and all two-way and three-way statistical interactions, on larval survival. The resulting estimates and confidence intervals are useful for determining whether selection surfaces differ from zero, but not whether they differ between food-web treatments. In the main text, we discuss why we do not put much stock in differences in the magnitude of selection between food-web treatments (last paragraph of **Quantifying the Adaptive Landscape**), and therefore do not present these calculations.

```{r Food-web model, cache=TRUE, echo=TRUE}
foodweb_model <- glmer(
  gall_survival ~ 
    -1 + Foodweb + 
    Foodweb:(sc.Diam + sc.log.Clutch + sc.sqrt.Pref) +
    Foodweb:(I(sc.Diam^2) + I(sc.log.Clutch^2) + I(sc.sqrt.Pref^2)) +
    Foodweb:(sc.Diam:sc.log.Clutch + sc.Diam:sc.sqrt.Pref + sc.log.Clutch:sc.sqrt.Pref) +
    (1|Genotype/Plant_Position/Gall_Number),
  data = gall_selection.df,
  family = binomial(link = logit), 
  control=glmerControl(optimizer = "bobyqa"))
```

We then used parametric bootstrapping (n = 1000) to calculate 95% confidence intervals of selection surfaces.

```{r Bootstrap food-web model, cache=TRUE, echo=TRUE}
n_boots_analysis <- 1000

boot_foodweb_model <- bootMer(
  x = foodweb_model, 
  FUN = fixef, 
  nsim = n_boots_analysis, 
  parallel="multicore", ncpus=32, seed=34)
```

In order to reliably estimate linear relationships, we removed all higher-order terms from the model [(Stinchcombe et al. 2008)](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1558-5646.2008.00449.x),

```{r Linear food-web model, cache=TRUE, echo=TRUE}
linear_foodweb_model <- glmer(
  gall_survival ~ 
    -1 + Foodweb + 
    Foodweb:(sc.Diam + sc.log.Clutch + sc.sqrt.Pref) +
    (1|Genotype/Plant_Position/Gall_Number),
  data = gall_selection.df,
  family = binomial(link = logit), 
  control=glmerControl(optimizer = "bobyqa"))
```

and refit using parametric bootstrapping.

```{r Bootstrap linear food-web model, cache=TRUE, echo=TRUE}
boot_linear_foodweb_model <- bootMer(
  x = linear_foodweb_model, 
  FUN = fixef, 
  nsim = n_boots_analysis, 
  parallel="multicore", ncpus=32, seed=34)
```

Chamber diameter, but not the other phenotypes, could be influenced by parasitism itself rather than being under natural selection. To estimate this potential bias, we subset our data to only include multi-chambered galls where there was variability in larval survival. We then fit a reduced model to estimate the bias in selection surfaces of chamber diameter in each food-web treatment.

```{r Bias in food-web model, cache=TRUE, echo=TRUE}

# Subset data
biased_foodweb_df <- gall_selection.df %>%
  group_by(Foodweb, Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  ungroup()

# Fit linear model with only chamber diameter 
biased_foodweb_model <- glmer(
  gall_survival ~ -1 + Foodweb + 
    Foodweb:sc.Diam + 
    (1|Genotype/Plant_Position/Gall_Number),
  data = biased_foodweb_df,
  family = binomial(link = logit), control=glmerControl(optimizer = "bobyqa"))

# Accuracy of confidence intervals is not a priority here, so we just use the asymptotic estimates rather than parametric bootstrapping.
biased_foodweb_confint <- tidy(biased_foodweb_model, conf.int=TRUE) %>% filter(group=="fixed")
```

```{r biased foodweb table}
# Print table
biased_foodweb_confint %>%
  #select(term, estimate, conf.low, conf.high) %>%
  filter(term %in% c("FoodwebOriginal:sc.Diam","FoodwebExtinction:sc.Diam")) %>%
  transmute(Term = term, Estimate = round(estimate, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
  knitr::kable(., caption = "Estimates of bias in selection surface of chamber diameter in each food-web treatment.")
```

We then gathered bootstrapped estimates from both models, only retaining nonlinear terms from the full model,

```{r Gather food-web alphas, echo=TRUE}
foodweb_bind <- bind_cols(
  as_tibble(boot_linear_foodweb_model$t),
  # only retain nonlinear coefficients from full model
  select(as_tibble(boot_foodweb_model$t), `FoodwebExtinction:I(sc.Diam^2)`:`FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref`) 
)
```

and subtracted the bias in chamber diameter to better approximate its selection surface.

```{r Adjust food-web alphas for chamber diameter, echo=TRUE}
foodweb_adj_Diam <- foodweb_bind %>%
  mutate(`FoodwebOriginal:sc.Diam` = `FoodwebOriginal:sc.Diam` - fixef(biased_foodweb_model)["FoodwebOriginal:sc.Diam"],
         `FoodwebExtinction:sc.Diam` = `FoodwebExtinction:sc.Diam` - fixef(biased_foodweb_model)["FoodwebExtinction:sc.Diam"])
```

We then summarize these bootstrapped estimates and 95% confidence intervals as a table.

```{r Tidy food-web alphas}

# Calculate differences between food-web treatments, then estimate means and confidence intervals
foodweb_alphas <- foodweb_adj_Diam %>%
  mutate(`FoodwebExtinction:(Intercept)`=inverse_logit(FoodwebExtinction), `FoodwebOriginal:(Intercept)`=inverse_logit(FoodwebOriginal)) %>% # logit transform to make intercept interpretable as a probability
  select(-FoodwebExtinction, -FoodwebOriginal) %>% # translated into (Intercept) terms
  mutate(`FoodwebDiff:(Intercept)` = `FoodwebExtinction:(Intercept)` - `FoodwebOriginal:(Intercept)`,
         `FoodwebDiff:sc.Diam` = `FoodwebExtinction:sc.Diam` - `FoodwebOriginal:sc.Diam`,
         `FoodwebDiff:sc.log.Clutch` = `FoodwebExtinction:sc.log.Clutch` - `FoodwebOriginal:sc.log.Clutch`,
         `FoodwebDiff:sc.sqrt.Pref` = `FoodwebExtinction:sc.sqrt.Pref` - `FoodwebOriginal:sc.sqrt.Pref`,
         `FoodwebDiff:I(sc.Diam^2)` = `FoodwebExtinction:I(sc.Diam^2)` - `FoodwebOriginal:I(sc.Diam^2)`,
         `FoodwebDiff:I(sc.log.Clutch^2)` = `FoodwebExtinction:I(sc.log.Clutch^2)` - `FoodwebOriginal:I(sc.log.Clutch^2)`,
         `FoodwebDiff:I(sc.sqrt.Pref^2)` = `FoodwebExtinction:I(sc.sqrt.Pref^2)` - `FoodwebOriginal:I(sc.sqrt.Pref^2)`,
         `FoodwebDiff:sc.Diam:sc.log.Clutch` = `FoodwebExtinction:sc.Diam:sc.log.Clutch` - `FoodwebOriginal:sc.Diam:sc.log.Clutch`,
         `FoodwebDiff:sc.Diam:sc.sqrt.Pref` = `FoodwebExtinction:sc.Diam:sc.sqrt.Pref` - `FoodwebOriginal:sc.Diam:sc.sqrt.Pref`,
         `FoodwebDiff:sc.log.Clutch:sc.sqrt.Pref` = `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref`) %>%
  summarise_all(funs(mean, conf.high, conf.low))

# Tidy up estimates
tidy_foodweb_alphas <- foodweb_alphas %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction", "Diff"))) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         coefficient_type = ifelse(term=="(Intercept)","Mean fitness",
                                   ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                          ifelse(grepl(":", .$term)==TRUE, "Correlational","Linear"))))
```

```{r Table of food web coefficients}
tidy_foodweb_alphas %>%
  filter(type != "Diff") %>%
  #select(Term = term, type, mean, conf.low, conf.high) %>% # coefficient_type, P_cutoff
  transmute(Term = term, Treatment = type, Estimate = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
  arrange(desc(Term)) %>%
 knitr::kable(., caption = "Table of estimates for larval survival and selection surfaces in each food-web treatment.")
```

```{r Table of differences in food web coefficients, include=FALSE}
tidy_foodweb_alphas %>%
  filter(type == "Diff") %>%
  #select(Term = term, mean, conf.low, conf.high) %>% # coefficient_type, P_cutoff
  transmute(Term = term, Treatment = type, Estimate = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
 knitr::kable(., caption = "Table of differences in mean fitness and selection surfaces for gall midges between food-web treatments.")
```

The table is good for details, but it is easier to see the results as a figure.

```{r Plot food-web coefficients, fig.cap="Plot of larval survival and selection surfaces in each food-web treatment."}
## Note that we do not focus on differences between treatments.

# Plot helpers
dodge_width <- position_dodge(width=0.5)
y_limits_foodweb_coefs <- c(min(filter(tidy_foodweb_alphas, type!="Diff")$conf.low), 
                            max(filter(tidy_foodweb_alphas, type!="Diff")$conf.high))
P_asterisk_foodweb_coefs <- 1.5
P_size <- 8
legend_foodweb_labels <- c("Original","Extinction")
point.size <- 3

# Plot mean fitness
plot_foodweb_mean_fitness <- tidy_foodweb_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_foodweb_alphas, type=="Diff", term=="(Intercept)"), aes(y=0.8, x=1.5, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  xlab("Food web") +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels)

# Plot linear trait-fitness relationships
plot_foodweb_linear_coefs <- tidy_foodweb_alphas %>%
  filter(type!="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_foodweb_alphas, type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")), aes(y=P_asterisk_foodweb_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Coefficient", limits=y_limits_foodweb_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam]","alpha[Clutch]","alpha[Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels)

# Plot quadratic selection surfaces
plot_foodweb_quadratic_coefs <- tidy_foodweb_alphas %>%
  filter(type!="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_foodweb_alphas, type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")), aes(y=P_asterisk_foodweb_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Coefficient", limits=y_limits_foodweb_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam:Diam]","alpha[Clutch:Clutch]","alpha[Pref:Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels)

# Plot correlational trait-fitness relationaships
plot_foodweb_correlational_coefs <- tidy_foodweb_alphas %>%
  filter(type!="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_foodweb_alphas, type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")), aes(y=P_asterisk_foodweb_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Coefficient", limits=y_limits_foodweb_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam:Clutch]","alpha[Diam:Pref]","alpha[Clutch:Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels)

# Format plots for manuscript
plot_foodweb_coefs <- plot_grid(
  plot_foodweb_mean_fitness + theme(legend.position = "none"),
  plot_foodweb_linear_coefs + theme(legend.position = "none"),
  plot_foodweb_quadratic_coefs + theme(legend.position = "none"),
  plot_foodweb_correlational_coefs + theme(legend.position = "none"), 
  nrow=2, align="hv", labels = "")

figure_foodweb_coefs <- plot_foodweb_coefs 
figure_foodweb_coefs
```

\

## Selection gradients for the gall midge: food-web treatments

Using the methods proposed by [Janzen and Stern (1998)](https://onlinelibrary.wiley.com/doi/10.1111/j.1558-5646.1998.tb02237.x), we used our estimates of mean fitness and selection surfaces to calculate selection gradients.

```{r Food-web selection gradients, echo=TRUE}

# Estimate mean fitness and mean "brackets" for each food-web treatment (see Janzen and Stern 1998 equation 4 for details about "brackets")
foodweb_original_predict <- predict(foodweb_model, newdata=filter(gall_selection.df, Foodweb=="Original"), type="response")
foodweb_original_mean_brackets <- mean(foodweb_original_predict * (1 - foodweb_original_predict))
foodweb_original_mean_fitness <- mean(foodweb_original_predict)

foodweb_extinction_predict <- predict(foodweb_model, newdata=filter(gall_selection.df, Foodweb=="Extinction"), type="response")
foodweb_extinction_mean_brackets <- mean(foodweb_extinction_predict * (1 - foodweb_extinction_predict))
foodweb_extinction_mean_fitness <- mean(foodweb_extinction_predict)

foodweb_original_gradient <- function(x) foodweb_original_mean_brackets * x / foodweb_original_mean_fitness
foodweb_extinction_gradient <- function(x) foodweb_extinction_mean_brackets * x / foodweb_extinction_mean_fitness

foodweb_original_raw_gradients <- select(foodweb_adj_Diam, starts_with("FoodwebOriginal:")) %>%
  transmute_all(funs(foodweb_original_gradient))

foodweb_extinction_raw_gradients <- select(foodweb_adj_Diam, starts_with("FoodwebExtinction:")) %>%
  transmute_all(funs(foodweb_extinction_gradient))
```

```{r Tidy up selection gradients}
# Calculate differences between food-web treatments, then estimate means and confidence intervals
foodweb_grads <- bind_cols(foodweb_original_raw_gradients, foodweb_extinction_raw_gradients) %>%
  mutate(`FoodwebDiff:sc.Diam` = `FoodwebExtinction:sc.Diam` - `FoodwebOriginal:sc.Diam`,
         `FoodwebDiff:sc.log.Clutch` = `FoodwebExtinction:sc.log.Clutch` - `FoodwebOriginal:sc.log.Clutch`,
         `FoodwebDiff:sc.sqrt.Pref` = `FoodwebExtinction:sc.sqrt.Pref` - `FoodwebOriginal:sc.sqrt.Pref`,
         `FoodwebDiff:I(sc.Diam^2)` = `FoodwebExtinction:I(sc.Diam^2)` - `FoodwebOriginal:I(sc.Diam^2)`,
         `FoodwebDiff:I(sc.log.Clutch^2)` = `FoodwebExtinction:I(sc.log.Clutch^2)` - `FoodwebOriginal:I(sc.log.Clutch^2)`,
         `FoodwebDiff:I(sc.sqrt.Pref^2)` = `FoodwebExtinction:I(sc.sqrt.Pref^2)` - `FoodwebOriginal:I(sc.sqrt.Pref^2)`,
         `FoodwebDiff:sc.Diam:sc.log.Clutch` = `FoodwebExtinction:sc.Diam:sc.log.Clutch` - `FoodwebOriginal:sc.Diam:sc.log.Clutch`,
         `FoodwebDiff:sc.Diam:sc.sqrt.Pref` = `FoodwebExtinction:sc.Diam:sc.sqrt.Pref` - `FoodwebOriginal:sc.Diam:sc.sqrt.Pref`,
         `FoodwebDiff:sc.log.Clutch:sc.sqrt.Pref` = `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref`) %>%
  summarise_all(funs(mean, conf.high, conf.low))

# Tidy up estimates
tidy_foodweb_grads <- foodweb_grads %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction", "Diff"))) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         gradient_type = ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                ifelse(grepl(":", .$term)==TRUE, "Correlational","Directional")),
         mean = ifelse(gradient_type=="Quadratic", 2*mean, mean),
         conf.low = ifelse(gradient_type=="Quadratic", 2*conf.low, conf.low),
         conf.high = ifelse(gradient_type=="Quadratic", 2*conf.high, conf.high))
```

We then reproduce the estimates of selection gradients and 95% confidence intervals reported in **Table 1** of the main text.

```{r Table of food web selection gradients}
tidy_foodweb_grads %>%
  filter(type != "Diff") %>%
  #select(term, type, mean, conf.low, conf.high) %>% # gradient_type, P_cutoff
  transmute(Term = term, Treatment = type, Estimate = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
  arrange(desc(Term)) %>%
knitr::kable(., caption = "Table of standardized selection gradients acting on gall midges in each food-web treatment.")
```

```{r Table of differences in selection gradients between food-web treatments, include=FALSE}
tidy_foodweb_grads %>%
  filter(type == "Diff") %>%
  #select(term, mean, conf.low, conf.high) %>% # gradient_type, P_cutoff
  transmute(Term = term, Treatment = type, Estimate = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
  knitr::kable(., caption = "Table of differences in selection gradients acting on gall traits between food-web treatment.")
```

```{r Curvature, include=FALSE}
# Calculate curvatures for original food web
foodweb_original_raw_curvs <- foodweb_original_raw_gradients %>%
  transmute(`FoodwebOriginal:C.Diam` = `FoodwebOriginal:I(sc.Diam^2)` - `FoodwebOriginal:sc.Diam`*`FoodwebOriginal:sc.Diam`,
         `FoodwebOriginal:C.Clutch` = `FoodwebOriginal:I(sc.log.Clutch^2)` - `FoodwebOriginal:sc.log.Clutch`*`FoodwebOriginal:sc.log.Clutch`,
         `FoodwebOriginal:C.Pref` = `FoodwebOriginal:I(sc.sqrt.Pref^2)` - `FoodwebOriginal:sc.sqrt.Pref`*`FoodwebOriginal:sc.sqrt.Pref`,
         `FoodwebOriginal:C.Diam:Clutch` = `FoodwebOriginal:sc.Diam:sc.log.Clutch` - `FoodwebOriginal:sc.Diam`*`FoodwebOriginal:sc.log.Clutch`,
         `FoodwebOriginal:C.Diam:Pref` = `FoodwebOriginal:sc.Diam:sc.sqrt.Pref` - `FoodwebOriginal:sc.Diam`*`FoodwebOriginal:sc.sqrt.Pref`,
         `FoodwebOriginal:C.Clutch:Pref` = `FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebOriginal:sc.log.Clutch`*`FoodwebOriginal:sc.sqrt.Pref`)

# Calculate curvatures for extinction food web
foodweb_extinction_raw_curvs <- foodweb_extinction_raw_gradients %>%
  transmute(`FoodwebExtinction:C.Diam` = `FoodwebExtinction:I(sc.Diam^2)` - `FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.Diam`,
         `FoodwebExtinction:C.Clutch` = `FoodwebExtinction:I(sc.log.Clutch^2)` - `FoodwebExtinction:sc.log.Clutch`*`FoodwebExtinction:sc.log.Clutch`,
         `FoodwebExtinction:C.Pref` = `FoodwebExtinction:I(sc.sqrt.Pref^2)` - `FoodwebExtinction:sc.sqrt.Pref`*`FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebExtinction:C.Diam:Clutch` = `FoodwebExtinction:sc.Diam:sc.log.Clutch` - `FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.log.Clutch`,
         `FoodwebExtinction:C.Diam:Pref` = `FoodwebExtinction:sc.Diam:sc.sqrt.Pref` - `FoodwebExtinction:sc.Diam`*`FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebExtinction:C.Clutch:Pref` = `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebExtinction:sc.log.Clutch`*`FoodwebExtinction:sc.sqrt.Pref`)

# Calculate differences between food-web treatments, then estimate means and confidence intervals
foodweb_curvs <- bind_cols(foodweb_original_raw_curvs, foodweb_extinction_raw_curvs) %>%
  mutate(`FoodwebDiff:C.Diam` = `FoodwebExtinction:C.Diam` - `FoodwebOriginal:C.Diam`,
         `FoodwebDiff:C.Clutch` = `FoodwebExtinction:C.Clutch` - `FoodwebOriginal:C.Clutch`,
         `FoodwebDiff:C.Pref` = `FoodwebExtinction:C.Pref` - `FoodwebOriginal:C.Pref`,
         `FoodwebDiff:C.Diam:Clutch` = `FoodwebExtinction:C.Diam:Clutch` - `FoodwebOriginal:C.Diam:Clutch`,
         `FoodwebDiff:C.Diam:Pref` = `FoodwebExtinction:C.Diam:Pref` - `FoodwebOriginal:C.Diam:Pref`,
         `FoodwebDiff:C.Clutch:Pref` = `FoodwebExtinction:C.Clutch:Pref` - `FoodwebOriginal:C.Clutch:Pref`) %>%
  summarise_all(funs(mean, conf.high, conf.low))

# Tidy up estimates
tidy_foodweb_curvs <- foodweb_curvs %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction", "Diff"))) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"))#,
         #gradient_type = ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
         #                       ifelse(grepl(":", .$term)==TRUE, "Correlational","Directional")),
         #mean = ifelse(gradient_type=="Quadratic", 2*mean, mean),
         #conf.low = ifelse(gradient_type=="Quadratic", 2*conf.low, conf.low),
         #conf.high = ifelse(gradient_type=="Quadratic", 2*conf.high, conf.high))
```

```{r Table of food web curvatures, include=FALSE}
tidy_foodweb_curvs %>%
  filter(type != "Diff") %>%
  select(term, type, mean, conf.low, conf.high, P_cutoff) %>% 
  mutate(mean = round(mean, 2), conf.low = round(conf.low, 2), conf.high = round(conf.high, 2)) %>%
knitr::kable(., caption = "Table of curvatures on gall traits in each food-web treatment.")
```

```{r Table of differences in curvatures between food-web treatments, include=FALSE}
tidy_foodweb_curvs %>%
  filter(type == "Diff") %>%
  select(term, mean, conf.low, conf.high, P_cutoff) %>%
  mutate(mean = round(mean, 2), conf.low = round(conf.low, 2), conf.high = round(conf.high, 2)) %>%
  knitr::kable(., caption = "Table of differences in curvatures on gall traits between food-web treatment.")
```

As before, we can visualize the table output as a figure.

```{r Plot food-web gradients, fig.cap="Plot of standardized selection gradients acting on gall midges in each food-web treatment."}

# Plot helpers
dodge_width <- position_dodge(width=0.5)
y_limits_foodweb_grads <- c(min(filter(tidy_foodweb_grads, type!="Diff")$conf.low), 
                            max(filter(tidy_foodweb_grads, type!="Diff")$conf.high))
P_asterisk_foodweb_grads <- 0.5
P_size <- 8
legend_foodweb_labels <- c("Original","Extinction")
legend_foodweb_title <- "Food web"
point.size <- 3

# Plot directional selection gradients
plot_foodweb_directional_grads <- tidy_foodweb_grads %>%
  filter(type!="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_foodweb_grads, type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")), aes(y=P_asterisk_foodweb_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_foodweb_grads) +
  scale_x_discrete(name="", labels=parse(text=c("beta[Diam]","beta[Clutch]","beta[Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels, name=legend_foodweb_title) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels, name=legend_foodweb_title)

# Plot quadratic selection gradients
plot_foodweb_quadratic_grads <- tidy_foodweb_grads %>%
  filter(type!="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_foodweb_grads, type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")), aes(y=P_asterisk_foodweb_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_foodweb_grads) +
  scale_x_discrete(name="", labels=parse(text=c("gamma[Diam:Diam]","gamma[Clutch:Clutch]","gamma[Pref:Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels, name=legend_foodweb_title) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels, name=legend_foodweb_title)

# Plot correlational selection gradients
plot_foodweb_correlational_grads <- tidy_foodweb_grads %>%
  filter(type!="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_foodweb_grads, type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")), aes(y=P_asterisk_foodweb_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_foodweb_grads) +
  scale_x_discrete(name="", labels=parse(text=c("gamma[Diam:Clutch]","gamma[Diam:Pref]","gamma[Clutch:Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels, name=legend_foodweb_title) +
  scale_color_manual(values=cbPalette[c(6,5)], labels=legend_foodweb_labels, name=legend_foodweb_title)

# Get legend
legend_foodweb_grads <- get_legend(plot_foodweb_directional_grads)

# Format plots for manuscript
plot_foodweb_grads <- plot_grid(
  plot_foodweb_directional_grads + theme(legend.position = "none", axis.title = element_blank()),
  plot_foodweb_quadratic_grads + theme(legend.position = "none", axis.title = element_blank()),
  plot_foodweb_correlational_grads + theme(legend.position = "none", axis.title = element_blank()), 
  nrow=2, align="hv", labels = "")
plot_legend_foodweb_grads <- ggdraw(plot_foodweb_grads) + draw_grob(legend_foodweb_grads, x = 0.7, y=-0.2)
y_title_foodweb_grads <- ggdraw() + draw_label("Selection gradient (SDs)", angle = 90)

figure_foodweb_grads <- plot_grid(y_title_foodweb_grads, plot_legend_foodweb_grads, ncol=2, rel_widths = c(0.05,1)) 
figure_foodweb_grads
```

\

## Selection surface for the gall midge: egg vs. larval parasitoids

Our extinction food-web treatment allows us to estimate the unique contribution of egg parasitoids to selection on gall midge traits. To estimate the unique contribution of larval parasitoids, we subset our data so that our original food-web treatment only contained attack by larval parasitoids (and gall survival). We then fit the same models as previously, including one to estimate bias.

```{r Egg vs. larval ptoid model, cache=TRUE, echo=TRUE}
# excludes cases of egg-parasitism from Original food web
egglarval_df <- gall_selection.df %>%
  filter(Foodweb == "Extinction" | Foodweb == "Original" & egg.ptoid < 1)

# fit model with same structure as foodweb_model
egglarval_model <- update(foodweb_model, data=egglarval_df)

# subset data to estimate bias
biased_egglarval_df <- egglarval_df %>%
  group_by(Foodweb, Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  ungroup()

# fit model with same structure as biased_foodweb_model
biased_egglarval_model <- update(biased_foodweb_model, data=biased_egglarval_df)
```

We then used parametric bootstrapping to estimate confidence intervals from the full model.

```{r Bootstrap egg vs. larval ptoid model, cache=TRUE, echo=TRUE}
boot_egglarval_model <- bootMer(
  x = egglarval_model, 
  FUN = fixef, 
  nsim = n_boots_analysis, 
  parallel="multicore", ncpus=32, seed=34)
```

We also fit a reduced model to reliably estimate linear selection surfaces,

```{r Linear egg vs. larval ptoid model, cache=TRUE, echo=TRUE}
linear_egglarval_model <- update(linear_foodweb_model, data = egglarval_df)
```

and used parametric bootstrapping to estimate those confidence intervals.

```{r Bootstrap linear egg vs. larval model, cache=TRUE, echo=TRUE}
boot_linear_egglarval_model <- bootMer(
  x = linear_egglarval_model, 
  FUN = fixef, 
  nsim = n_boots_analysis, 
  parallel="multicore", ncpus=32, seed=34)
```

We then combined estimates from the linear model with those nonlinear terms from the full model,

```{r Egg vs. larval ptoid alpha coefficients, echo=TRUE}
egglarval_bind <- bind_cols(
  as_tibble(boot_linear_egglarval_model$t),
  # only retain nonlinear coefficients from full model
  select(as_tibble(boot_egglarval_model$t), `FoodwebExtinction:I(sc.Diam^2)`:`FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref`) 
)

```

and adjusted chamber diameter coefficients to reduce bias.

```{r Adjust egg vs. larval alphas for chamber diameter, echo=TRUE}
egglarval_adj_Diam <- as_tibble(boot_egglarval_model$t) %>%
  mutate(`FoodwebOriginal:sc.Diam` = `FoodwebOriginal:sc.Diam` - fixef(biased_egglarval_model)["FoodwebOriginal:sc.Diam"],
         `FoodwebExtinction:sc.Diam` = `FoodwebExtinction:sc.Diam` - fixef(biased_egglarval_model)["FoodwebExtinction:sc.Diam"])
```

Below, we display the estimates and 95% confidence intervals of selection surfaces imposed by egg and larval parasitoids on gall midges as a table,

```{r Egg vs. larval ptoid alphas for table}

# Calculate differences between food-web treatments, then estimate means and confidence intervals
egglarval_alphas <- egglarval_adj_Diam %>%
  mutate(`FoodwebExtinction:(Intercept)`=inverse_logit(FoodwebExtinction), `FoodwebOriginal:(Intercept)`=inverse_logit(FoodwebOriginal)) %>% # logit transform to make intercept interpretable as a probability
  select(-FoodwebExtinction, -FoodwebOriginal) %>% # translated into (Intercept) terms
  mutate(`FoodwebDiff:(Intercept)` = `FoodwebExtinction:(Intercept)` - `FoodwebOriginal:(Intercept)`,
         `FoodwebDiff:sc.Diam` = `FoodwebExtinction:sc.Diam` - `FoodwebOriginal:sc.Diam`,
         `FoodwebDiff:sc.log.Clutch` = `FoodwebExtinction:sc.log.Clutch` - `FoodwebOriginal:sc.log.Clutch`,
         `FoodwebDiff:sc.sqrt.Pref` = `FoodwebExtinction:sc.sqrt.Pref` - `FoodwebOriginal:sc.sqrt.Pref`,
         `FoodwebDiff:I(sc.Diam^2)` = `FoodwebExtinction:I(sc.Diam^2)` - `FoodwebOriginal:I(sc.Diam^2)`,
         `FoodwebDiff:I(sc.log.Clutch^2)` = `FoodwebExtinction:I(sc.log.Clutch^2)` - `FoodwebOriginal:I(sc.log.Clutch^2)`,
         `FoodwebDiff:I(sc.sqrt.Pref^2)` = `FoodwebExtinction:I(sc.sqrt.Pref^2)` - `FoodwebOriginal:I(sc.sqrt.Pref^2)`,
         `FoodwebDiff:sc.Diam:sc.log.Clutch` = `FoodwebExtinction:sc.Diam:sc.log.Clutch` - `FoodwebOriginal:sc.Diam:sc.log.Clutch`,
         `FoodwebDiff:sc.Diam:sc.sqrt.Pref` = `FoodwebExtinction:sc.Diam:sc.sqrt.Pref` - `FoodwebOriginal:sc.Diam:sc.sqrt.Pref`,
         `FoodwebDiff:sc.log.Clutch:sc.sqrt.Pref` = `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref`) %>%
  summarise_all(funs(mean, conf.high, conf.low))

# Tidy up estimates
tidy_egglarval_alphas <- egglarval_alphas %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction", "Diff"))) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         coefficient_type = ifelse(term=="(Intercept)","Mean fitness",
                                   ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                          ifelse(grepl(":", .$term)==TRUE, "Correlational","Linear"))))
```

```{r Table of egg vs. larval coefficients}
tidy_egglarval_alphas %>%
  filter(type != "Diff") %>%
  #select(term, type, mean, conf.low, conf.high) %>% # coefficient_type, P_cutoff
  transmute(Term = term, 
            Type = ifelse(type == "Original", "Larval ptoid", "Egg ptoid"),
            Estimate = round(mean, 2), 
            `2.5%` = round(conf.low, 2), 
            `97.5%` = round(conf.high, 2)) %>% 
  arrange(desc(Term)) %>%
  knitr::kable(., caption = "Table of gall survival and standardized selection surfaces imposed by egg and larval parasitoids.")
```

```{r Table of differences in egg vs. larval coefficients, include=FALSE}
tidy_egglarval_alphas %>%
  filter(type == "Diff") %>%
  #select(term, mean, conf.low, conf.high) %>% # coefficient_type, P_cutoff
  transmute(Term = term, 
            Type = ifelse(type == "Original", "Larval ptoid", "Egg ptoid"),
            Estimate = round(mean, 2), 
            `2.5%` = round(conf.low, 2), 
            `97.5%` = round(conf.high, 2)) %>%   knitr::kable(., caption = "Table of differences in gall selection surfaces between egg and larval parasitoids.")
```

and in figure form.

```{r Plot egglarval coefficients, fig.cap="Plot of gall survival and standardized selection surfaces imposed by egg and larval parasitoids."}

# Plot helpers
dodge_width <- position_dodge(width=0.5)
y_limits_egglarval_coefs <- c(min(filter(tidy_egglarval_alphas, type!="Diff")$conf.low), 
                            max(filter(tidy_egglarval_alphas, type!="Diff")$conf.high))
P_asterisk_egglarval_coefs <- 1.3
P_size <- 8
legend_egglarval_labels <- c("Larval","Egg")
point.size <- 3

# Plot mean fitness
plot_egglarval_mean_fitness <- tidy_egglarval_alphas %>%
  filter(type!="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_egglarval_alphas, type=="Diff", term=="(Intercept)"), aes(y=0.9, x=1.5, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Larval survival", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="Parasitoid", breaks=c("Original","Extinction"), labels=c("Larval","Egg")) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels)

# Plot linear trait-fitness relationaships
plot_egglarval_linear_coefs <- tidy_egglarval_alphas %>%
  filter(type!="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_egglarval_alphas, type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")), aes(y=P_asterisk_egglarval_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Coefficient", limits=y_limits_egglarval_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam]","alpha[Clutch]","alpha[Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels)

# Plot quadratic selection surfaces
plot_egglarval_quadratic_coefs <- tidy_egglarval_alphas %>%
  filter(type!="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_egglarval_alphas, type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")), aes(y=P_asterisk_egglarval_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Coefficient", limits=y_limits_egglarval_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam:Diam]","alpha[Clutch:Clutch]","alpha[Pref:Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels)

# Plot correlational trait-fitness relationaships
plot_egglarval_correlational_coefs <- tidy_egglarval_alphas %>%
  filter(type!="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_egglarval_alphas, type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")), aes(y=P_asterisk_egglarval_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Coefficient", limits=y_limits_egglarval_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam:Clutch]","alpha[Diam:Pref]","alpha[Clutch:Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels)

# Format plots for manuscript
plot_egglarval_coefs <- plot_grid(
  plot_egglarval_mean_fitness + theme(legend.position = "none"),
  plot_egglarval_linear_coefs + theme(legend.position = "none"),
  plot_egglarval_quadratic_coefs + theme(legend.position = "none"),
  plot_egglarval_correlational_coefs + theme(legend.position = "none"), 
  nrow=2, align="hv", labels = "")

figure_egglarval_coefs <- plot_egglarval_coefs 
figure_egglarval_coefs
```

\

## Selection gradients for the gall midge: egg vs. larval parasitoids

We again use the method of [Janzen and Stern (1998)](https://onlinelibrary.wiley.com/doi/10.1111/j.1558-5646.1998.tb02237.x) to calculate selection gradients from our estimates of larval survival and selection surfaces,

```{r Egg vs. larval ptoid selection gradients, echo=TRUE}

# Estimate mean fitness and mean "brackets" for each food-web treatment (see Janzen and Stern 1998 equation 4 for details about "brackets")
egglarval_original_predict <- predict(egglarval_model, newdata=filter(egglarval_df, Foodweb=="Original"), type="response")
egglarval_original_mean_brackets <- mean(egglarval_original_predict * (1 - egglarval_original_predict))
egglarval_original_mean_fitness <- mean(egglarval_original_predict)

egglarval_extinction_predict <- predict(egglarval_model, newdata=filter(egglarval_df, Foodweb=="Extinction"), type="response")
egglarval_extinction_mean_brackets <- mean(egglarval_extinction_predict * (1 - egglarval_extinction_predict))
egglarval_extinction_mean_fitness <- mean(egglarval_extinction_predict)

egglarval_original_gradient <- function(x) egglarval_original_mean_brackets * x / egglarval_original_mean_fitness
egglarval_extinction_gradient <- function(x) egglarval_extinction_mean_brackets * x / egglarval_extinction_mean_fitness

egglarval_original_raw_gradients <- select(egglarval_adj_Diam, starts_with("FoodwebOriginal:")) %>%
  transmute_all(funs(egglarval_original_gradient))

egglarval_extinction_raw_gradients <- select(egglarval_adj_Diam, starts_with("FoodwebExtinction:")) %>%
  transmute_all(funs(egglarval_extinction_gradient))
```

and summarize our estimates and 95% confidence intervals as a table,

```{r Tidy egg vs. larval ptoid selection gradients}

# Calculate differences between food-web treatments, then estimate means and confidence intervals
egglarval_grads <- bind_cols(egglarval_original_raw_gradients, egglarval_extinction_raw_gradients) %>%
  mutate(`FoodwebDiff:sc.Diam` = `FoodwebExtinction:sc.Diam` - `FoodwebOriginal:sc.Diam`,
         `FoodwebDiff:sc.log.Clutch` = `FoodwebExtinction:sc.log.Clutch` - `FoodwebOriginal:sc.log.Clutch`,
         `FoodwebDiff:sc.sqrt.Pref` = `FoodwebExtinction:sc.sqrt.Pref` - `FoodwebOriginal:sc.sqrt.Pref`,
         `FoodwebDiff:I(sc.Diam^2)` = `FoodwebExtinction:I(sc.Diam^2)` - `FoodwebOriginal:I(sc.Diam^2)`,
         `FoodwebDiff:I(sc.log.Clutch^2)` = `FoodwebExtinction:I(sc.log.Clutch^2)` - `FoodwebOriginal:I(sc.log.Clutch^2)`,
         `FoodwebDiff:I(sc.sqrt.Pref^2)` = `FoodwebExtinction:I(sc.sqrt.Pref^2)` - `FoodwebOriginal:I(sc.sqrt.Pref^2)`,
         `FoodwebDiff:sc.Diam:sc.log.Clutch` = `FoodwebExtinction:sc.Diam:sc.log.Clutch` - `FoodwebOriginal:sc.Diam:sc.log.Clutch`,
         `FoodwebDiff:sc.Diam:sc.sqrt.Pref` = `FoodwebExtinction:sc.Diam:sc.sqrt.Pref` - `FoodwebOriginal:sc.Diam:sc.sqrt.Pref`,
         `FoodwebDiff:sc.log.Clutch:sc.sqrt.Pref` = `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref`) %>%
  summarise_all(funs(mean, conf.high, conf.low))

# Tidy up estimates
tidy_egglarval_grads <- egglarval_grads %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         gradient_type = ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                ifelse(grepl(":", .$term)==TRUE, "Correlational","Directional")),
         mean = ifelse(gradient_type=="Quadratic", 2*mean, mean),
         conf.low = ifelse(gradient_type=="Quadratic", 2*conf.low, conf.low),
         conf.high = ifelse(gradient_type=="Quadratic", 2*conf.high, conf.high))
```

```{r Egg vs. larval selection gradients}
tidy_egglarval_grads %>%
  filter(type != "Diff") %>%
  #select(term, type, mean, conf.low, conf.high) %>% # gradient_type, P_cutoff
  transmute(Term = term, 
            Parasitoid = ifelse(type == "Original", "Larval ptoid", "Egg ptoid"), 
            Estimate = round(mean, 2), 
            `2.5%` = round(conf.low, 2), 
            `97.5%` = round(conf.high, 2)) %>%
  arrange(desc(Term)) %>%
knitr::kable(., caption = "Table of selection gradients imposed by egg and larval parasitoids on gall midges.")
```

```{r Table of differences in selection gradients between egg and larval parasitoids, include=FALSE}
tidy_egglarval_grads %>%
  filter(type == "Diff") %>%
  #select(term, gradient_type, mean, conf.low, conf.high) %>% # gradient_type, P_cutoff
  transmute(Term = term, 
            Parasitoid = ifelse(type == "Original", "Larval ptoid", "Egg ptoid"), 
            Estimate = round(mean, 2), 
            `2.5%` = round(conf.low, 2), 
            `97.5%` = round(conf.high, 2)) %>%
  knitr::kable(., caption = "Table of differences in selection gradients imposed by egg and larval parasitoids.")
```

and as a figure.

```{r Plot egg vs. larval ptoid gradients, fig.cap="Plot of selection gradients imposed by egg and larval parasitoids on gall midges."}

## Note: I'm not plotting asterisks that denote significant differences (*P* < 0.05) between parasitoid guilds.

# Plot helpers
dodge_width <- position_dodge(width=0.5)
y_limits_egglarval_grads <- c(min(filter(tidy_egglarval_grads, type!="Diff")$conf.low), 
                            max(filter(tidy_egglarval_grads, type!="Diff")$conf.high))
P_asterisk_egglarval_grads <- 0.25
P_size <- 8
legend_egglarval_labels <- c("Larval","Egg")
legend_egglarval_title <- "Parasitoid"
point.size <- 3

# Plot directional selection gradients
plot_egglarval_directional_grads <- tidy_egglarval_grads %>%
  filter(type!="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction"))) %>%
  ggplot(., aes(x=term)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_egglarval_grads, type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")), aes(y=P_asterisk_egglarval_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_egglarval_grads) +
  scale_x_discrete(name="", labels=parse(text=c("beta[Diam]","beta[Clutch]","beta[Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels, name=legend_egglarval_title) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels, name=legend_egglarval_title)

# Plot quadratic selection gradients
plot_egglarval_quadratic_grads <- tidy_egglarval_grads %>%
  filter(type!="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction"))) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_egglarval_grads, type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")), aes(y=P_asterisk_egglarval_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_egglarval_grads) +
  scale_x_discrete(name="", labels=parse(text=c("gamma[Diam:Diam]","gamma[Clutch:Clutch]","gamma[Pref:Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels, name=legend_egglarval_title) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels, name=legend_egglarval_title)

# Plot correlational selection gradients
plot_egglarval_correlational_grads <- tidy_egglarval_grads %>%
  filter(type!="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction"))) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_egglarval_grads, type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")), aes(y=P_asterisk_egglarval_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_egglarval_grads) +
  scale_x_discrete(name="", labels=parse(text=c("gamma[Diam:Clutch]","gamma[Diam:Pref]","gamma[Clutch:Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels, name=legend_egglarval_title) +
  scale_color_manual(values=cbPalette[c(4,5)], labels=legend_egglarval_labels, name=legend_egglarval_title)

# Get legend
legend_egglarval_grads <- get_legend(plot_egglarval_directional_grads)

# Format plots for manuscript
plot_egglarval_grads <- plot_grid(
  plot_egglarval_directional_grads + theme(legend.position = "none", axis.title = element_blank()),
  plot_egglarval_quadratic_grads + theme(legend.position = "none", axis.title = element_blank()),
  plot_egglarval_correlational_grads + theme(legend.position = "none", axis.title = element_blank()), 
  nrow=2, align="hv", labels = "")
plot_legend_egglarval_grads <- ggdraw(plot_egglarval_grads) + draw_grob(legend_egglarval_grads, x = 0.7, y=-0.2)
y_title_egglarval_grads <- ggdraw() + draw_label("Selection gradient (SDs)", angle = 90)

figure_egglarval_grads <- plot_grid(y_title_egglarval_grads, plot_legend_egglarval_grads, ncol=2, rel_widths = c(0.05,1)) 
figure_egglarval_grads
```

## Reproduce Figure 2 in main text

We used the full model to generate predicted estimates of mean larval survival for changes in the mean trait value of the population. I restrict these predictions to +/- 1 SD because we can only reliably estimate the shape of the adaptive landscape near the mean phenotype of the population [(Arnold et al., 2001)](https://link.springer.com/article/10.1023/A:1013373907708). Other trait values are held constant at the mean phenotype (i.e. trait = 0). We then performed 100 bootstraps to illustrate the uncertainty in our estimates.

```{r Get Gall Diameter Fitness Landscape, cache=TRUE, echo=TRUE}
n_boots_plots <- 100

## Gall diameter
newdata_Diam <- bind_rows(
  expand.grid(Foodweb = "Original", sc.Diam = seq(-1,1,length.out=1000), sc.log.Clutch = 0, sc.sqrt.Pref = 0),
  expand.grid(Foodweb = "Extinction", sc.Diam = seq(-1,1,length.out=1000), sc.log.Clutch = 0, sc.sqrt.Pref = 0))

RF_Diam <- bootstrap_fitness(
  logistic_model = foodweb_model, 
  newdata = newdata_Diam,
  bootstraps=n_boots_plots)
```

```{r Get Clutch Size Fitness Landscape, cache=TRUE, echo=TRUE}
## Clutch size
newdata_Clutch <- bind_rows(
  expand.grid(Foodweb = "Original", sc.Diam = 0, sc.log.Clutch = seq(-1,1,length.out=1000), sc.sqrt.Pref = 0),
  expand.grid(Foodweb = "Extinction", sc.Diam = 0, sc.log.Clutch = seq(-1,1,length.out=1000), sc.sqrt.Pref = 0))

RF_Clutch <- bootstrap_fitness(
  logistic_model = foodweb_model,
  newdata = newdata_Clutch,
  bootstraps=n_boots_plots)
```

```{r Get Oviposition preference Fitness Landscape, cache=TRUE, echo=TRUE}
## Oviposition preference
newdata_Pref <- bind_rows(
  expand.grid(Foodweb = "Original", sc.Diam = 0, sc.log.Clutch = 0, sc.sqrt.Pref = seq(-1,1,length.out=1000)),
  expand.grid(Foodweb = "Extinction", sc.Diam = 0, sc.log.Clutch = 0, sc.sqrt.Pref = seq(-1,1,length.out=1000)))

RF_Pref <- bootstrap_fitness(
  logistic_model = foodweb_model, 
  newdata = newdata_Pref,
  bootstraps=n_boots_plots)
```

We then gathered these predictions to plot the adaptive landscape of gall midge traits in each food-web treatment. Note that we plot mean larval survival on a log-scale to reflect the mathematical definition of the adaptive landscape.

```{r Plot Univariate Landscapes}

## Create plots for each panel of Figure 2 in main text.

uni_breaks <- c(0.1,0.5,1.0)

# Gall size 
AF_uni_Diam <- RF_Diam$absolute_fitness %>% 
  select(-sc.log.Clutch, -sc.sqrt.Pref) %>%
  gather(ID, absolute_fitness, -sc.Diam, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate"),
         Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction"))) %>% 
  ggplot(., aes(x=sc.Diam, y=absolute_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Chamber diameter (SDs)") +
  ylab("Mean larval survival") +
  scale_color_manual(values = treatment_colors, name = "Food web") +
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_log10(breaks=uni_breaks, labels=uni_breaks) + annotation_logticks(sides = "l")+
  coord_cartesian(ylim=c(0.07,1))

# Oviposition preference
AF_uni_Pref <- RF_Pref$absolute_fitness %>% 
  select(-sc.Diam, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.sqrt.Pref, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate"),
         Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction"))) %>% 
  ggplot(., aes(x=sc.sqrt.Pref, y=absolute_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Oviposition preference (SDs)") +
  ylab("Mean larval survival") +
  scale_color_manual(values = treatment_colors, name = "Food web") +
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_log10(breaks=uni_breaks, labels=uni_breaks) + annotation_logticks(sides = "l") +
  coord_cartesian(ylim=c(0.07,1))

# Clutch size 
AF_uni_Clutch <- RF_Clutch$absolute_fitness %>% 
  select(-sc.Diam, -sc.sqrt.Pref) %>%
  gather(ID, absolute_fitness, -sc.log.Clutch, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate"),
         Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction"))) %>% 
  ggplot(., aes(x=sc.log.Clutch, y=absolute_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Clutch size (SDs)") +
  ylab("Mean larval survival") +
  scale_color_manual(values = treatment_colors, name = "Food web") +
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_log10(breaks=uni_breaks, labels=uni_breaks) + annotation_logticks(sides = "l") +
  coord_cartesian(ylim=c(0.07,1))
```

```{r Univariate-Fitness-Landscapes, fig.cap="Adaptive landscapes of gall midge phenotypes in the original food web and with the extinction of larval parasitoids."}

## Fit plots together to reproduce Figure 2.

# Get legend
AF_uni_legend <- get_legend(AF_uni_Diam + theme(legend.title.align = 0.5))

# Make plots
AF_uni_plots <- plot_grid(AF_uni_Diam + theme(legend.position = "none"), 
                       AF_uni_Pref + theme(legend.position = "none", 
                                           axis.title.y = element_blank()) + xlab("Preference (SDs)"),
                      AF_uni_Clutch + theme(legend.position = "none",
                                             axis.title.y = element_blank()), 
                       labels = "AUTO", nrow = 1, align='hv')

AF_gradients <- plot_grid(AF_uni_plots, AF_uni_legend, ncol = 2, rel_widths = c(0.8,0.2)) 
AF_gradients

save_plot(filename = "UV_landscapes.pdf", plot = AF_gradients, base_height = 5, base_width = 8.5)
```

```{r Plot Univariate RF Landscapes, fig.cap="Adaptive landscapes of gall phenotypes in terms of relative fitness.", include=FALSE}

## Replot gall midge adaptive landscapes in terms of relative fitness

RF_plot_foodweb_Clutch_df <- RF_Clutch$absolute_fitness %>% 
  select(-sc.Diam, -sc.sqrt.Pref) %>%
  gather(ID, absolute_fitness, -sc.log.Clutch, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate"),
         Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction"))) %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness)) %>%
  ungroup()

RF_uni_Clutch <- RF_plot_foodweb_Clutch_df %>%
  ggplot(., aes(x=sc.log.Clutch, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Clutch size (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  geom_hline(yintercept=1, linetype="dotted")

# Oviposition preference

RF_plot_foodweb_Pref_df <- RF_Pref$absolute_fitness %>% 
  select(-sc.Diam, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.sqrt.Pref, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate"),
         Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction"))) %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
  ungroup()

RF_uni_Pref <- RF_plot_foodweb_Pref_df %>%
  ggplot(., aes(x=sc.sqrt.Pref, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Oviposition preference (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  theme(legend.title.align = 0.5) +
  geom_hline(yintercept=1, linetype="dotted")

# Chamber diameter
RF_plot_foodweb_Diam_df <- RF_Diam$absolute_fitness %>% 
  select(-sc.sqrt.Pref, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.Diam, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate"),
         Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction"))) %>%
  group_by(Foodweb) %>%
  mutate(relative_fitness = absolute_fitness/mean(absolute_fitness))%>%
  ungroup()

RF_uni_Diam <- RF_plot_foodweb_Diam_df %>%
  ggplot(., aes(x=sc.Diam, y=relative_fitness, group=interaction(ID,Foodweb), color=Foodweb, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Chamber diameter (SDs)") +
  ylab("Relative fitness") +
  scale_color_manual(values = cbPalette[c(6,5)], name = "Food web") + 
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  coord_cartesian(ylim=c(0.2,2)) +
  theme(legend.title.align = 0.5) +
  geom_hline(yintercept=1, linetype="dotted")

# Get legend
RF_uni_legend <- get_legend(RF_uni_Diam)

# Make plots
RF_uni_plots <- plot_grid(RF_uni_Diam + theme(legend.position = "none", 
                                             axis.text.x = element_text(size=10),
                                             axis.title.x = element_text(size=11)),
                     RF_uni_Pref + theme(legend.position = "none",                          
                                             axis.title.y = element_blank(), 
                                             axis.text.x = element_text(size=10),
                                             axis.title.x = element_text(size=11)),
                     RF_uni_Clutch + theme(legend.position = "none",                          
                                             axis.title.y = element_blank(), 
                                             axis.text.x = element_text(size=10),
                                             axis.title.x = element_text(size=11)),
                     ncol = 3, align='hv')

RF_gradients <- plot_grid(RF_uni_plots, RF_uni_legend, ncol = 2, rel_widths = c(0.8,0.2)) 
RF_gradients
```

## Reproduce Figure 3 in main text

As before, we use the full model to generate predicted estimates of mean larval survival, but do so now for different phenotypic combinations.

```{r Get Clutch Size x Oviposition preference Landscape, cache=TRUE, echo=TRUE}
## Clutch size x Oviposition preference
newdata_Clutch.Pref <- bind_rows(
  expand.grid(Foodweb = "Original", sc.Diam = 0, sc.log.Clutch = seq(-1,1,length.out=1000), sc.sqrt.Pref = seq(-1,1,length.out=1000)),
  expand.grid(Foodweb = "Extinction", sc.Diam = 0, sc.log.Clutch = seq(-1,1,length.out=1000), sc.sqrt.Pref = seq(-1,1,length.out=1000)))

RF_Clutch.Pref <- bootstrap_fitness(
  logistic_model = foodweb_model, 
  newdata = newdata_Clutch.Pref,
  bootstraps=NULL)
```

```{r Get Gall Diameter x Clutch Size Landscape, cache=TRUE, echo=TRUE}
## Gall diameter x Clutch size
newdata_Diam.Clutch <- bind_rows(
  expand.grid(Foodweb = "Original", sc.Diam = seq(-1,1,length.out=1000), sc.log.Clutch = seq(-1,1,length.out=1000), sc.sqrt.Pref = 0),
  expand.grid(Foodweb = "Extinction", sc.Diam = seq(-1,1,length.out=1000), sc.log.Clutch = seq(-1,1,length.out=1000), sc.sqrt.Pref=0))

RF_Diam.Clutch <- bootstrap_fitness(
  logistic_model = foodweb_model, 
  newdata = newdata_Diam.Clutch,
  bootstraps=NULL)
```

```{r Get Gall Diameter x Oviposition preference Landscape, cache=TRUE, echo=TRUE}
## Gall diameter by oviposition preference
newdata_Diam.Pref <- bind_rows(
  expand.grid(Foodweb = "Original", sc.Diam = seq(-1,1,length.out=1000), sc.log.Clutch = 0, sc.sqrt.Pref = seq(-1,1,length.out=1000)),
  expand.grid(Foodweb = "Extinction", sc.Diam = seq(-1,1,length.out=1000), sc.log.Clutch = 0, sc.sqrt.Pref = seq(-1,1,length.out=1000)))

RF_Diam.Pref <- bootstrap_fitness(
  logistic_model = foodweb_model, 
  newdata = newdata_Diam.Pref,
  bootstraps=NULL)
```

And gather these predictions to plot two-dimensional views of the adaptive landscape for each combination of phenotypic traits in each food-web treatment.

```{r Get plots of multivariate adaptive landscape}

## Create plots for each panel.

# Plot helpers
AF_range <- bind_rows(RF_Clutch.Pref$absolute_fitness, RF_Diam.Clutch$absolute_fitness, RF_Diam.Pref$absolute_fitness) %>%
  summarise(min = min(average), max = max(average))

my_breaks <- c(0.2,0.4,0.6,0.8)

# Set contour breaks
breaks <- predict(foodweb_model, newdata=data.frame(sc.Diam = c(0,0), sc.log.Clutch = c(0,0), sc.sqrt.Pref = c(0,0), Foodweb = c("Original","Extinction")), type="response", re.form=~0)
names(breaks) <- c("Original","Extinction")

labels_2D <- list("Original"="Original", "Extinction"="Extinction")
labeller_2D <- function(variable, value) {
  return(labels_2D[value])
}

# Size x Clutch
RF_Diam.Clutch$absolute_fitness <- RF_Diam.Clutch$absolute_fitness %>%
  mutate(Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction")))

AF_plot_Diam.Clutch <- RF_Diam.Clutch$absolute_fitness %>%
  filter(Foodweb == "Original") %>%
  ggplot(., aes(x=sc.Diam, y=sc.log.Clutch)) +
  geom_raster(aes(fill=average)) +
  stat_contour(aes(z = average), breaks = breaks["Original"], linetype = "dotted", color = "black") +
  geom_raster(data = filter(RF_Diam.Clutch$absolute_fitness, Foodweb == "Extinction"), aes(fill=average)) +
  stat_contour(data = filter(RF_Diam.Clutch$absolute_fitness, Foodweb == "Extinction"), aes(z = average), breaks = breaks["Extinction"], linetype = "dotted", color = "black") +
  facet_wrap(~Foodweb) + # , labeller = labeller_2D
  xlab("Gall diameter (SDs)") +
  ylab("Clutch size (SDs)") +
  scale_x_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_fill_viridis(name = "Mean larval\nsurvival", trans="log", breaks = my_breaks, labels = my_breaks, limits = c(AF_range$min, AF_range$max)) +
  theme(strip.background = element_blank(),
        strip.text.x=element_text(margin=margin(b=4)))

# Size x Preference
RF_Diam.Pref$absolute_fitness <- RF_Diam.Pref$absolute_fitness %>% 
  mutate(Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction")))

AF_plot_Diam.Pref <- RF_Diam.Pref$absolute_fitness %>% 
  filter(Foodweb == "Original") %>%
  ggplot(., aes(x=sc.Diam, y=sc.sqrt.Pref)) +
  geom_raster(aes(fill=average)) +
  stat_contour(aes(z = average), breaks = breaks["Original"], linetype = "dotted", color = "black") +
  geom_raster(data = filter(RF_Diam.Pref$absolute_fitness, Foodweb == "Extinction"), aes(fill=average)) +
  stat_contour(data = filter(RF_Diam.Pref$absolute_fitness, Foodweb == "Extinction"), aes(z = average), breaks = breaks["Extinction"], linetype = "dotted", color = "black") +
  facet_wrap(~Foodweb) + # , labeller = labeller_2D
  xlab("Chamber diameter (SDs)") +
  ylab("Oviposition preference (SDs)") +
  scale_x_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_fill_viridis(name = "Mean larval\nsurvival", trans="log", breaks = my_breaks, labels = my_breaks, limits = c(AF_range$min, AF_range$max)) +
  theme(strip.background = element_blank(),
        strip.text.x=element_text(margin=margin(b=4)))

# Clutch x Preference
RF_Clutch.Pref$absolute_fitness <- RF_Clutch.Pref$absolute_fitness %>%
  mutate(Foodweb = factor(ifelse(Foodweb == "Original", "Original", "Extinction"), levels = c("Original", "Extinction")))

AF_plot_Clutch.Pref <- RF_Clutch.Pref$absolute_fitness %>%
  filter(Foodweb == "Original") %>%
  ggplot(., aes(x=sc.sqrt.Pref, y=sc.log.Clutch)) +
  geom_raster(aes(fill=average)) +
  stat_contour(aes(z = average), breaks = breaks["Original"], linetype = "dotted", color = "black") +
  geom_raster(data = filter(RF_Clutch.Pref$absolute_fitness, Foodweb == "Extinction"), aes(fill=average)) +
  stat_contour(data = filter(RF_Clutch.Pref$absolute_fitness, Foodweb == "Extinction"), aes(z = average), breaks = breaks["Extinction"], linetype = "dotted", color = "black") +
  facet_wrap(~Foodweb) + # , labeller = labeller_2D
  xlab("Oviposition preference (SDs)") +
  ylab("Clutch size (SDs)") +
  scale_x_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_y_continuous(labels = c(-1,-0.5,0,0.5,1)) +
  scale_fill_viridis(name = "Mean larval\nsurvival", trans="log", breaks = my_breaks, labels = my_breaks, limits = c(AF_range$min, AF_range$max)) +
  theme(strip.background = element_blank(),
        strip.text.x=element_text(margin=margin(b=4)))

```

```{r Figure-Multivariate-Landscapes, fig.cap="Two dimensional view of adaptive landscapes of gall midge phenotypes in the original food web and with the extinction of larval parasitoids. Dotted lines denote fitness for the mean phenotype. Arrows represent the vector of directional selection gradients acting on each phenotype."}

## Fit plots together to reproduce Figure 3.

# Get legend
AF_landscape_legend <- get_legend(AF_plot_Diam.Pref + theme(legend.text = element_text(size=10)))

# Arrow data
gradients_Diam.Clutch <- data.frame(Foodweb = c("Original","Extinction"),
                                    sc.Diam = c(0,0), 
                                    sc.log.Clutch = c(0,0), 
                                    xend = c(filter(tidy_foodweb_grads, type=="Original", term=="sc.Diam")$mean,
                                             filter(tidy_foodweb_grads, type=="Extinction", term=="sc.Diam")$mean), 
                                    yend = c(filter(tidy_foodweb_grads, type=="Original", term=="sc.log.Clutch")$mean, 
                                             filter(tidy_foodweb_grads, type=="Extinction", term=="sc.log.Clutch")$mean))

gradients_Clutch.Pref <- data.frame(Foodweb = c("Original","Extinction"),
                                    sc.sqrt.Pref = c(0,0), 
                                    sc.log.Clutch = c(0,0), 
                                    xend = c(filter(tidy_foodweb_grads, type=="Original", term=="sc.sqrt.Pref")$mean, 
                                             filter(tidy_foodweb_grads, type=="Extinction", term=="sc.sqrt.Pref")$mean), 
                                    yend = c(filter(tidy_foodweb_grads, type=="Original", term=="sc.log.Clutch")$mean, 
                                             filter(tidy_foodweb_grads, type=="Extinction", term=="sc.log.Clutch")$mean))

gradients_Diam.Pref <- data.frame(Foodweb = c("Original","Extinction"),
                                    sc.Diam = c(0,0), 
                                    sc.sqrt.Pref = c(0,0), 
                                    xend = c(filter(tidy_foodweb_grads, type=="Original", term=="sc.Diam")$mean,
                                             filter(tidy_foodweb_grads, type=="Extinction", term=="sc.Diam")$mean), 
                                    yend = c(filter(tidy_foodweb_grads, type=="Original", term=="sc.sqrt.Pref")$mean, 
                                             filter(tidy_foodweb_grads, type=="Extinction", term=="sc.sqrt.Pref")$mean))

# Format plots
AF_landscape_plots <- plot_grid(
  AF_plot_Diam.Clutch + theme(legend.position = "none", 
                              axis.title.x = element_blank(),
                             axis.text.x = element_blank()) +
    geom_segment(data = gradients_Diam.Clutch, 
                 aes(x = sc.Diam, y = sc.log.Clutch, xend = xend, yend = yend), 
                 arrow = arrow(length = unit(0.03, "npc"))),
  AF_plot_Clutch.Pref + theme(legend.position = "none", 
                              axis.title.y = element_blank(),
                              axis.text.y = element_blank()) + xlab("Preference (SDs)") +
    geom_segment(data = gradients_Clutch.Pref, 
                 aes(x = sc.sqrt.Pref, y = sc.log.Clutch, xend = xend, yend = yend), 
                 arrow = arrow(length = unit(0.03, "npc"))), 
  AF_plot_Diam.Pref + theme(legend.position = "none") + ylab("Preference (SDs)") +
    geom_segment(data = gradients_Diam.Pref, 
                 aes(x = sc.Diam, y = sc.sqrt.Pref, xend = xend, yend = yend), 
                 arrow = arrow(length = unit(0.03, "npc"))), 
  nrow=2, align="hv", labels = "AUTO")

AF_landscape_2d <- ggdraw(AF_landscape_plots) + draw_grob(AF_landscape_legend, x = 0.7, y=-0.2)
AF_landscape_2d
save_plot(filename = "MV_landscapes.pdf", plot = AF_landscape_2d, base_width=8, base_height = 6)
```

## Slope and curvature of the adaptive landscape of gall traits

Below is the code used to get estimates for the slope ($\beta$) and curvature ($\gamma-\beta\beta^\text{T}$) of the adaptive landscape in each food-web treatment. If 95% confidence intervals of selection gradients overlap zero, then they were set to zero; otherwise, we retain the estimate.

For the original food web, we create the matrix of $\beta$s,

```{r Betas for original food web, echo=TRUE}
# Get direction selection gradients from original food web
original_betas_df <- tidy_foodweb_grads %>%
  filter(type=="Original", gradient_type=="Directional") %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)

# Relabel each term 
original_betas_df$term <- factor(
  original_betas_df$term, 
  levels=c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref"), 
  labels=c("Diam","Clutch","Pref"), order=TRUE)

rownames(original_betas_df) <- original_betas_df$term

# Create empty matrix
original_betas <- matrix(nrow = 3, 
                         ncol = 1, 
                         dimnames = list(c("Diam","Clutch","Pref"), c("")))

# Populate the empty matrix with ifelse() statements
# If the 95% CI does not overlap zero, 
# then the "low" multiplied by the "high" interval is always greater than zero.
# If that's true, then I retain the estimate.
# If it's false, then I set the estimate to zero.
original_betas["Diam",] <- ifelse(
  original_betas_df["Diam","low"]*original_betas_df["Diam","high"] > 0, 
  original_betas_df["Diam","gradient"], 0)

original_betas["Clutch",] <- ifelse(
  original_betas_df["Clutch","low"]*original_betas_df["Clutch","high"] > 0, 
  original_betas_df["Clutch","gradient"], 0)

original_betas["Pref",] <- ifelse(
  original_betas_df["Pref","low"]*original_betas_df["Pref","high"] > 0,
  original_betas_df["Pref","gradient"], 0)

original_betas <- round(original_betas, 2)
```

and the matrix of $\gamma$s.

```{r Gammas for original food web, echo=TRUE}
# Get nonlinear selection gradients from original food web
original_gammas_df <- tidy_foodweb_grads %>%
  filter(type=="Original", gradient_type %in% c("Quadratic","Correlational")) %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)

# Relabel each term
original_gammas_df$term <- factor(
  original_gammas_df$term, 
  levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)",
           "sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"), 
  labels=c("Diam^2","Clutch^2","Pref^2","Diam:Clutch","Diam:Pref","Clutch:Pref"), order=TRUE)

rownames(original_gammas_df) <- original_gammas_df$term

# Create empty matrix  
original_gammas <- matrix(nrow = 3, 
                          ncol = 3, 
                          dimnames = list(c("Diam","Clutch","Pref"), c("Diam","Clutch","Pref")))

# Populate the empty matrix with ifelse() statements 
# to determine if 95% CI overlaps with zero or not (see previous logic)
original_gammas["Diam","Diam"] <- ifelse(
  original_gammas_df["Diam^2","low"]*original_gammas_df["Diam^2","high"] > 0,
  original_gammas_df["Diam^2","gradient"], 0)

original_gammas["Clutch","Clutch"] <- ifelse(
  original_gammas_df["Clutch^2","low"]*original_gammas_df["Clutch^2","high"] > 0, 
  original_gammas_df["Clutch^2","gradient"], 0)

original_gammas["Pref","Pref"] <- ifelse(
  original_gammas_df["Pref^2","low"]*original_gammas_df["Pref^2","high"] > 0, 
  original_gammas_df["Pref^2","gradient"], 0)

original_gammas["Diam","Clutch"] <- ifelse(
  original_gammas_df["Diam:Clutch","low"]*original_gammas_df["Diam:Clutch","high"] > 0,
  original_gammas_df["Diam:Clutch","gradient"], 0)

original_gammas["Clutch","Diam"] <- original_gammas["Diam","Clutch"]

original_gammas["Diam","Pref"] <- ifelse(
  original_gammas_df["Diam:Pref","low"]*original_gammas_df["Diam:Pref","high"] > 0,
  original_gammas_df["Diam:Pref","gradient"], 0)

original_gammas["Pref","Diam"] <- original_gammas["Diam","Pref"]

original_gammas["Clutch","Pref"] <- ifelse(
  original_gammas_df["Clutch:Pref","low"]*original_gammas_df["Clutch:Pref","high"] > 0,
  original_gammas_df["Clutch:Pref","gradient"], 0)

original_gammas["Pref","Clutch"] <- original_gammas["Clutch","Pref"]

original_gammas <- round(original_gammas, 2)
```

Likewise for the extinction web, we create the matrix of $\beta$s,

```{r Betas for extinction food web, echo=TRUE}
# Get direction selection gradients from extinction food web
extinction_betas_df <- tidy_foodweb_grads %>%
  filter(type=="Extinction", gradient_type=="Directional") %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)

# Relabel each term 
extinction_betas_df$term <- factor(
  extinction_betas_df$term, 
  levels=c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref"), 
  labels=c("Diam","Clutch","Pref"), order=TRUE)

rownames(extinction_betas_df) <- extinction_betas_df$term

# Create empty matrix
extinction_betas <- matrix(nrow = 3, 
                           ncol = 1, 
                           dimnames = list(c("Diam","Clutch","Pref"), c("")))

# Populate the empty matrix with ifelse() statements 
# to determine if 95% CI overlaps with zero or not (see previous logic)
extinction_betas["Diam",] <- ifelse(
  extinction_betas_df["Diam","low"]*extinction_betas_df["Diam","high"] > 0,
  extinction_betas_df["Diam","gradient"], 0)

extinction_betas["Clutch",] <- ifelse(
  extinction_betas_df["Clutch","low"]*extinction_betas_df["Clutch","high"] > 0,
  extinction_betas_df["Clutch","gradient"], 0)

extinction_betas["Pref",] <- ifelse(
  extinction_betas_df["Pref","low"]*extinction_betas_df["Pref","high"] > 0,
  extinction_betas_df["Pref","gradient"], 0)

extinction_betas <- round(extinction_betas, 2)
```

and $\gamma$s.

```{r Gammas for extinction food web, echo=TRUE}
# Get nonlinear selection gradients from extinction food web
extinction_gammas_df <- tidy_foodweb_grads %>%
  filter(type=="Extinction", gradient_type %in% c("Quadratic","Correlational")) %>%
  select(term, gradient=mean, low=conf.low, high=conf.high)

# Relabel each term
extinction_gammas_df$term <- factor(
  extinction_gammas_df$term, 
  levels=c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)",
           "sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref"),
  labels=c("Diam^2","Clutch^2","Pref^2","Diam:Clutch","Diam:Pref","Clutch:Pref"), order=TRUE)

rownames(extinction_gammas_df) <- extinction_gammas_df$term

# Create empty matrix
extinction_gammas <- matrix(nrow = 3, 
                            ncol = 3, 
                            dimnames = list(c("Diam","Clutch","Pref"), c("Diam","Clutch","Pref")))

# Populate the empty matrix with ifelse() statements 
# to determine if 95% CI overlaps with zero or not (see previous logic)
extinction_gammas["Diam","Diam"] <- ifelse(
  extinction_gammas_df["Diam^2","low"]*extinction_gammas_df["Diam^2","high"] > 0,
  extinction_gammas_df["Diam^2","gradient"], 0)

extinction_gammas["Clutch","Clutch"] <- ifelse(
  extinction_gammas_df["Clutch^2","low"]*extinction_gammas_df["Clutch^2","high"] > 0,
  extinction_gammas_df["Clutch^2","gradient"], 0)

extinction_gammas["Pref","Pref"] <- ifelse(
  extinction_gammas_df["Pref^2","low"]*extinction_gammas_df["Pref^2","high"] > 0,
  extinction_gammas_df["Pref^2","gradient"], 0)

extinction_gammas["Diam","Clutch"] <- ifelse(
  extinction_gammas_df["Diam:Clutch","low"]*extinction_gammas_df["Diam:Clutch","high"] > 0,
  extinction_gammas_df["Diam:Clutch","gradient"], 0)

extinction_gammas["Clutch","Diam"] <- extinction_gammas["Diam","Clutch"]

extinction_gammas["Diam","Pref"] <- ifelse(
  extinction_gammas_df["Diam:Pref","low"]*extinction_gammas_df["Diam:Pref","high"] > 0,
  extinction_gammas_df["Diam:Pref","gradient"], 0)

extinction_gammas["Pref","Diam"] <- extinction_gammas["Diam","Pref"]

extinction_gammas["Clutch","Pref"] <- ifelse(
  extinction_gammas_df["Clutch:Pref","low"]*extinction_gammas_df["Clutch:Pref","high"] > 0,
  extinction_gammas_df["Clutch:Pref","gradient"], 0)

extinction_gammas["Pref","Clutch"] <- extinction_gammas["Clutch","Pref"]

extinction_gammas <- round(extinction_gammas, 2)
```

After all of this, we can easily calculate the curvature matrix ($\gamma-\beta\beta^\text{T}$) in each food-web treatment.

```{r Calculate curvature in each food-web treatment, echo=TRUE}
original_curvature <- original_gammas - original_betas %*% t(original_betas)

extinction_curvature <- extinction_gammas - extinction_betas %*% t(extinction_betas)
```

I then used these estimates to populate the curvature matrices presented in the **Results** of the main text, but also reproduce them here:

$$\textbf{C} = \begin{pmatrix} \text{C}_{\text{Diam:Diam}}&& \\ \text{C}_{\text{Clutch:Diam}} & \text{C}_{\text{Clutch:Clutch}} & \\ \text{C}_{\text{Pref:Diam}} & \text{C}_{\text{Pref:Clutch}} & \text{C}_{\text{Pref:Pref}} \end{pmatrix}$$

$$\textbf{C}_{\text{Original}} = \begin{pmatrix} 
`r round(original_curvature["Diam","Diam"],2)` &  &  \\  
`r round(original_curvature["Clutch","Diam"],2)` & `r round(original_curvature["Clutch","Clutch"],2)` &  \\  
`r round(original_curvature["Pref","Diam"],2)` & `r round(original_curvature["Pref","Clutch"],2)` & `r round(original_curvature["Pref","Pref"],2)` \end{pmatrix}$$

$$\textbf{C}_{\text{Extinction}} = \begin{pmatrix} 
`r round(extinction_curvature["Diam","Diam"],2)` &  &  \\  
`r round(extinction_curvature["Clutch","Diam"],2)` & `r round(extinction_curvature["Clutch","Clutch"],2)` &  \\  
`r round(extinction_curvature["Pref","Diam"],2)` & `r round(extinction_curvature["Pref","Clutch"],2)` & `r round(extinction_curvature["Pref","Pref"],2)` \end{pmatrix}$$

\

## Selection surfaces for the egg parasitoid

We fit the same models as before, except presence/absence of an egg parasitoid is the response variable. Importantly, we can only estimate selection by calculating the difference in probability of observing egg parasitoids between food-web treatments. We do this by calculating the difference in bootstrapped estimates between treatments after fitting our models.

```{r Egg vs. egg ptoid model, cache=TRUE, echo=TRUE}

# convert "gall_survival" to egg parasitoid survival. Note that both Iteomyia pupa and larva parasitoids result in 0. We do not change the name of the response variable to make clear that we are using the same statistical models as for Iteomyia.
eggegg_df <- gall_selection.df %>%
  mutate(gall_survival = ifelse(egg.ptoid == 1, 1, 0))

# fit trait-fitness relationship using the same model structure
eggegg_model <- update(foodweb_model, data=eggegg_df)

# subset data for quantifying biased selection on chamber diameter
biased_eggegg_df <- eggegg_df %>%
  group_by(Foodweb, Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  ungroup()

# quantify bias
biased_eggegg_model <- update(biased_foodweb_model, data=biased_eggegg_df)
```

We used parametric bootstrapping to calculate 95% confidence intervals of selection surfaces.

```{r Bootstrap egg vs. egg ptoid model, cache=TRUE, echo=TRUE}
boot_eggegg_model <- bootMer(
  x = eggegg_model, 
  FUN = fixef, 
  nsim = n_boots_analysis, 
  parallel="multicore", ncpus=32, seed=34)
```

As before, we removed all higher-order terms to estimate linear relationships,

```{r Linear egg vs. egg ptoid model, cache=TRUE, echo=TRUE}
linear_eggegg_model <- update(linear_foodweb_model, data=eggegg_df)
```

and refit using parametric bootstrapping.

```{r Bootstrap linear egg vs. egg model, cache=TRUE, echo=TRUE}
boot_linear_eggegg_model <- bootMer(
  x = linear_eggegg_model, 
  FUN = fixef, 
  nsim = n_boots_analysis, 
  parallel="multicore", ncpus=32, seed=34)
```

We then gather bootstrapped estimates from the full (nonlinear terms only) and reduced models,

```{r gather egg vs. egg ptoid alphas, echo=TRUE}
eggegg_bind <- bind_cols(
  as_tibble(boot_linear_eggegg_model$t),
  # only retain nonlinear coefficients from full model
  select(as_tibble(boot_eggegg_model$t), `FoodwebExtinction:I(sc.Diam^2)`:`FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref`) 
)
```

and adjusted for biased coefficients in chamber diameter.

```{r Adjust egg vs. egg ptoid alphas for chamber diameter, echo=TRUE}
eggegg_adj_Diam <- as_tibble(boot_eggegg_model$t) %>%
  mutate(`FoodwebOriginal:sc.Diam` = `FoodwebOriginal:sc.Diam` - fixef(biased_eggegg_model)["FoodwebOriginal:sc.Diam"],
         `FoodwebExtinction:sc.Diam` = `FoodwebExtinction:sc.Diam` - fixef(biased_eggegg_model)["FoodwebExtinction:sc.Diam"])
```

Finally, we calculate the differences between food-web treatments in order to estimate selection imposed by larval parasitoids.

```{r Egg vs. egg ptoid alpha differences, echo=TRUE}
eggegg_alphas <- eggegg_adj_Diam %>%
  # logit transform to make intercept interpretable as a probability of survival
  mutate(`FoodwebExtinction:(Intercept)` = inverse_logit(FoodwebExtinction),
         `FoodwebOriginal:(Intercept)` = inverse_logit(FoodwebOriginal)) %>% 
  select(-FoodwebExtinction, -FoodwebOriginal) %>% # translated into (Intercept) terms
  # calculate differences
  mutate(`FoodwebDiff:(Intercept)` = `FoodwebOriginal:(Intercept)` - `FoodwebExtinction:(Intercept)`,
         `FoodwebDiff:sc.Diam` = `FoodwebOriginal:sc.Diam` - `FoodwebExtinction:sc.Diam`,
         `FoodwebDiff:sc.log.Clutch` = `FoodwebOriginal:sc.log.Clutch` - `FoodwebExtinction:sc.log.Clutch`,
         `FoodwebDiff:sc.sqrt.Pref` = `FoodwebOriginal:sc.sqrt.Pref` - `FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebDiff:I(sc.Diam^2)` = `FoodwebOriginal:I(sc.Diam^2)` - `FoodwebExtinction:I(sc.Diam^2)`,
         `FoodwebDiff:I(sc.log.Clutch^2)` = `FoodwebOriginal:I(sc.log.Clutch^2)` - `FoodwebExtinction:I(sc.log.Clutch^2)`,
         `FoodwebDiff:I(sc.sqrt.Pref^2)` = `FoodwebOriginal:I(sc.sqrt.Pref^2)` - `FoodwebExtinction:I(sc.sqrt.Pref^2)`,
         `FoodwebDiff:sc.Diam:sc.log.Clutch` = `FoodwebOriginal:sc.Diam:sc.log.Clutch` - `FoodwebExtinction:sc.Diam:sc.log.Clutch`,
         `FoodwebDiff:sc.Diam:sc.sqrt.Pref` = `FoodwebOriginal:sc.Diam:sc.sqrt.Pref` - `FoodwebExtinction:sc.Diam:sc.sqrt.Pref`,
         `FoodwebDiff:sc.log.Clutch:sc.sqrt.Pref` = `FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref`) %>%
  summarise_all(funs(mean, conf.high, conf.low))
```

Below, we display the estimates and 95% confidence intervals in selection surfaces for egg parasitoids as a table,

```{r Tidy egg vs. egg ptoid alpha coefficients}

# Tidy up estimates
tidy_eggegg_alphas <- eggegg_alphas %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  mutate(type = factor(type, levels = c("Original", "Extinction", "Diff"))) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         coefficient_type = ifelse(term=="(Intercept)","Mean fitness",
                                   ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                          ifelse(grepl(":", .$term)==TRUE, "Correlational","Linear"))))
```

```{r Egg ptoid coefficients in each food-web treatment, include=FALSE}
tidy_eggegg_alphas %>%
  filter(type != "Diff") %>%
  #select(term, type, mean, conf.low, conf.high) %>% # coefficient_type, P_cutoff
  transmute(Term = term, Treatment = type, Estimate = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
knitr::kable(., caption = "Table of selection surfaces for egg parasitoids in original and extinction food webs.")
```

```{r Table of differences in egg parasitoid coefficients between food-web treatments}
tidy_eggegg_alphas %>%
  filter(type == "Diff") %>%
  #select(term, mean, conf.low, conf.high) %>% # coefficient_type, P_cutoff
  transmute(Term = term, `Original - Extinction` = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
  knitr::kable(., caption = "Table of standardized selection surfaces acting on egg parasitoids (*Platygaster* sp.).")
```

and in figure form.

```{r Plot egg vs. egg ptoid coefficients, fig.cap="Plot of selection surfaces acting on egg parasitoids."}

# Plot helpers
dodge_width <- position_dodge(width=0.5)
y_limits_eggegg_coefs <- c(min(filter(tidy_eggegg_alphas, type=="Diff")$conf.low), 
                            max(filter(tidy_eggegg_alphas, type=="Diff")$conf.high))
P_asterisk_eggegg_coefs <- 1.7
P_size <- 8
legend_eggegg_labels <- c("Original","Extinction")
point.size <- 3

# Plot mean fitness
plot_eggegg_mean_fitness <- tidy_eggegg_alphas %>%
  filter(type=="Diff", term=="(Intercept)") %>%
  ggplot(., aes(x=type)) +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_eggegg_alphas, type=="Diff", term=="(Intercept)"), aes(y=0.9, x=1.5, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Probability of observing\negg parasitoid", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="") +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels) +
  scale_color_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels)

# Plot linear trait-fitness relationaships
plot_eggegg_linear_coefs <- tidy_eggegg_alphas %>%
  filter(type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_eggegg_alphas, type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")), aes(y=P_asterisk_eggegg_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Original - Extinction", limits=y_limits_eggegg_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam]","alpha[Clutch]","alpha[Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels) +
  scale_color_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels)

# Plot quadratic selection surfaces
plot_eggegg_quadratic_coefs <- tidy_eggegg_alphas %>%
  filter(type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_eggegg_alphas, type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")), aes(y=P_asterisk_eggegg_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Original - Extinction", limits=y_limits_eggegg_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam:Diam]","alpha[Clutch:Clutch]","alpha[Pref:Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels) +
  scale_color_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels)

# Plot correlational trait-fitness relationaships
plot_eggegg_correlational_coefs <- tidy_eggegg_alphas %>%
  filter(type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_eggegg_alphas, type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")), aes(y=P_asterisk_eggegg_coefs, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Original - Extinction", limits=y_limits_eggegg_coefs) +
  scale_x_discrete(name="", labels=parse(text=c("alpha[Diam:Clutch]","alpha[Diam:Pref]","alpha[Clutch:Pref]"))) +
  facet_wrap(~coefficient_type) +
  scale_fill_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels) +
  scale_color_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels)

# Format plots for manuscript
plot_eggegg_coefs <- plot_grid(
  #plot_eggegg_mean_fitness + theme(legend.position = "none"),
  plot_eggegg_linear_coefs + theme(legend.position = "none"),
  plot_eggegg_quadratic_coefs + theme(legend.position = "none"),
  plot_eggegg_correlational_coefs + theme(legend.position = "none"), 
  nrow=2, align="hv", labels = "")

figure_eggegg_coefs <- plot_eggegg_coefs 
figure_eggegg_coefs
```

\

## Selection gradients for the egg parasitoid

```{r Egg vs. egg ptoid selection gradients, echo=TRUE}

# Estimate mean fitness and mean "brackets" for each food-web treatment (see Janzen and Stern 1998 equation 4 for details about "brackets")
eggegg_original_predict <- predict(eggegg_model, newdata=filter(eggegg_df, Foodweb=="Original"), type="response")
eggegg_original_mean_brackets <- mean(eggegg_original_predict * (1 - eggegg_original_predict))
eggegg_original_mean_fitness <- mean(eggegg_original_predict)

eggegg_extinction_predict <- predict(eggegg_model, newdata=filter(eggegg_df, Foodweb=="Extinction"), type="response")
eggegg_extinction_mean_brackets <- mean(eggegg_extinction_predict * (1 - eggegg_extinction_predict))
eggegg_extinction_mean_fitness <- mean(eggegg_extinction_predict)

eggegg_original_gradient <- function(x) eggegg_original_mean_brackets * x / eggegg_original_mean_fitness
eggegg_extinction_gradient <- function(x) eggegg_extinction_mean_brackets * x / eggegg_extinction_mean_fitness

eggegg_original_raw_gradients <- select(eggegg_adj_Diam, starts_with("FoodwebOriginal:")) %>%
  transmute_all(funs(eggegg_original_gradient))

eggegg_extinction_raw_gradients <- select(eggegg_adj_Diam, starts_with("FoodwebExtinction:")) %>%
  transmute_all(funs(eggegg_extinction_gradient))
```

Calculate differences between food-web treatments, then estimate means and confidence intervals

```{r Calculate diffs in egg vs. egg ptoid gradients, echo=TRUE}
eggegg_grads <- bind_cols(eggegg_original_raw_gradients, eggegg_extinction_raw_gradients) %>%
  mutate(`FoodwebDiff:sc.Diam` = `FoodwebOriginal:sc.Diam` - `FoodwebExtinction:sc.Diam`,
         `FoodwebDiff:sc.log.Clutch` = `FoodwebOriginal:sc.log.Clutch` - `FoodwebExtinction:sc.log.Clutch`,
         `FoodwebDiff:sc.sqrt.Pref` = `FoodwebOriginal:sc.sqrt.Pref` - `FoodwebExtinction:sc.sqrt.Pref`,
         `FoodwebDiff:I(sc.Diam^2)` = `FoodwebOriginal:I(sc.Diam^2)` - `FoodwebExtinction:I(sc.Diam^2)`,
         `FoodwebDiff:I(sc.log.Clutch^2)` = `FoodwebOriginal:I(sc.log.Clutch^2)` - `FoodwebExtinction:I(sc.log.Clutch^2)`,
         `FoodwebDiff:I(sc.sqrt.Pref^2)` = `FoodwebOriginal:I(sc.sqrt.Pref^2)` - `FoodwebExtinction:I(sc.sqrt.Pref^2)`,
         `FoodwebDiff:sc.Diam:sc.log.Clutch` = `FoodwebOriginal:sc.Diam:sc.log.Clutch` - `FoodwebExtinction:sc.Diam:sc.log.Clutch`,
         `FoodwebDiff:sc.Diam:sc.sqrt.Pref` = `FoodwebOriginal:sc.Diam:sc.sqrt.Pref` - `FoodwebExtinction:sc.Diam:sc.sqrt.Pref`,
         `FoodwebDiff:sc.log.Clutch:sc.sqrt.Pref` = `FoodwebOriginal:sc.log.Clutch:sc.sqrt.Pref` - `FoodwebExtinction:sc.log.Clutch:sc.sqrt.Pref`) %>%
  summarise_all(funs(mean, conf.high, conf.low))
```

Below, we display estimates of standardized selection gradients (and 95% confidence intervals) imposed by larval parasitoids on egg parasitoids as a table,

```{r Tidy egg vs. egg ptoid selection gradients}

# Tidy up estimates
tidy_eggegg_grads <- eggegg_grads %>%
  t() %>%
  as.data.frame() %>%
  transmute(rows = rownames(.), value=V1) %>%
  separate(col = rows, into = c("term","estimate"), sep="_") %>%
  separate(col = term, into = c("type","term_1","term_2"), sep=":", extra="drop") %>%
  mutate(term = ifelse(is.na(term_2), term_1, paste(term_1,term_2,sep=":"))) %>%
  separate(col = type, into = c("extra","type"), sep = 7) %>%
  select(type, term, estimate, value) %>%
  spread(estimate, value) %>%
  mutate(P_cutoff = ifelse(conf.high*conf.low < 0, "","*"),
         gradient_type = ifelse(grepl("\\^2", .$term)==TRUE, "Quadratic",
                                ifelse(grepl(":", .$term)==TRUE, "Correlational","Directional")),
         mean = ifelse(gradient_type=="Quadratic", 2*mean, mean),
         conf.low = ifelse(gradient_type=="Quadratic", 2*conf.low, conf.low),
         conf.high = ifelse(gradient_type=="Quadratic", 2*conf.high, conf.high))
```

```{r Egg vs. ptoid selection gradients, include=FALSE}
tidy_eggegg_grads %>%
  filter(type != "Diff") %>%
  #select(term, type, mean, conf.low, conf.high) %>% # gradient_type, P_cutoff
  transmute(Term = term, Treatment = type, Estimate = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
knitr::kable(., caption = "Table of standardized selection gradients acting on egg parasitoids in original and extinction food webs.")
```

```{r Table of differences in selection gradients acting on egg parasitoids between food-web treatment}
tidy_eggegg_grads %>%
  filter(type == "Diff") %>%
  #select(term, mean, conf.low, conf.high) %>% # gradient_type, P_cutoff
  transmute(Term = term, `Original - Extinction` = round(mean, 2), `2.5%` = round(conf.low, 2), `97.5%` = round(conf.high, 2)) %>%
  knitr::kable(., caption = "Table of standardized selection gradients acting on egg parasitoids (*Platygaster* sp.).")
```

and in figure form.

```{r Plot egg vs. egg ptoid gradients, fig.cap="Plot of standardized selection gradients acting on egg parasitoids."}

# Plot helpers
dodge_width <- position_dodge(width=0.5)
y_limits_eggegg_grads <- c(min(filter(tidy_eggegg_grads, type=="Diff")$conf.low), 
                            max(filter(tidy_eggegg_grads, type=="Diff")$conf.high))
P_asterisk_eggegg_grads <- 0.45
P_size <- 8
legend_eggegg_labels <- c("Original","Extinction")
legend_eggegg_title <- "Food web"
point.size <- 3

# Plot directional selection gradients
plot_eggegg_directional_grads <- tidy_eggegg_grads %>%
  filter(type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_eggegg_grads, type=="Diff", term %in% c("sc.Diam","sc.log.Clutch","sc.sqrt.Pref")), aes(y=P_asterisk_eggegg_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_eggegg_grads) +
  scale_x_discrete(name="", labels=parse(text=c("beta[Diam]","beta[Clutch]","beta[Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels, name=legend_eggegg_title) +
  scale_color_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels, name=legend_eggegg_title)

# Plot quadratic selection gradients
plot_eggegg_quadratic_grads <- tidy_eggegg_grads %>%
  filter(type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_eggegg_grads, type=="Diff", term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.sqrt.Pref^2)")), aes(y=P_asterisk_eggegg_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_eggegg_grads) +
  scale_x_discrete(name="", labels=parse(text=c("gamma[Diam:Diam]","gamma[Clutch:Clutch]","gamma[Pref:Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels, name=legend_eggegg_title) +
  scale_color_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels, name=legend_eggegg_title)

# Plot correlational selection gradients
plot_eggegg_correlational_grads <- tidy_eggegg_grads %>%
  filter(type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")) %>%
  ggplot(., aes(x=term, color=type)) +
  geom_hline(yintercept=0, linetype="dotted") +
  geom_linerange(aes(ymax=conf.high, ymin=conf.low, color=type), position=dodge_width) +
  geom_point(aes(y=mean, fill=type), position=dodge_width, shape=21, color="black", size=point.size) +
  #geom_text(data=filter(tidy_eggegg_grads, type=="Diff", term %in% c("sc.Diam:sc.log.Clutch","sc.Diam:sc.sqrt.Pref","sc.log.Clutch:sc.sqrt.Pref")), aes(y=P_asterisk_eggegg_grads, label=P_cutoff), size=P_size, color="black") +
  scale_y_continuous(name="Gradient", limits=y_limits_eggegg_grads) +
  scale_x_discrete(name="", labels=parse(text=c("gamma[Diam:Clutch]","gamma[Diam:Pref]","gamma[Clutch:Pref]"))) +
  facet_wrap(~gradient_type) +
  scale_fill_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels, name=legend_eggegg_title) +
  scale_color_manual(values=cbPalette[c(2,5)], labels=legend_eggegg_labels, name=legend_eggegg_title)

# Get legend
legend_eggegg_grads <- get_legend(plot_eggegg_directional_grads)

# Format plots for manuscript
plot_eggegg_grads <- plot_grid(
  plot_eggegg_directional_grads + theme(legend.position = "none", axis.title = element_blank()),
  plot_eggegg_quadratic_grads + theme(legend.position = "none", axis.title = element_blank()),
  plot_eggegg_correlational_grads + theme(legend.position = "none", axis.title = element_blank()), 
  nrow=2, align="hv", labels = "")
#plot_legend_eggegg_grads <- ggdraw(plot_eggegg_grads) + draw_grob(legend_eggegg_grads, x = 0.7, y=-0.2)
y_title_eggegg_grads <- ggdraw() + draw_label("Original - Extinction", angle = 90)

figure_eggegg_grads <- plot_grid(y_title_eggegg_grads, plot_eggegg_grads, ncol=2, rel_widths = c(0.05,1)) 
figure_eggegg_grads
```

## Reproduce Figure 4 in main text

Using a similar approach as for gall midge traits, we plot the difference in the probability of observing egg parasitoids (original - extinction) as a function of gall midge (now egg parasitoid) traits. Gall midge preference was the only phenotype of egg parasitoids that appeared to be under selection, so we only calculate this relationship,

```{r Egg parasitoid bootstraps, cache=TRUE, echo=TRUE}
## Oviposition preference
eggegg_RF_Pref <- bootstrap_fitness(
  logistic_model = eggegg_model, 
  newdata = newdata_Pref,
  bootstraps=n_boots_plots) 
```

and plot it.

```{r Egg-Parasitoid-Selection, fig.cap="Selection imposed by larval parasitoids on egg parasitoids (*Platygaster* sp.)."}

eggptoid_plot_df <- eggegg_RF_Pref$absolute_fitness %>% 
  select(-sc.Diam, -sc.log.Clutch) %>%
  gather(ID, absolute_fitness, -sc.sqrt.Pref, -Foodweb) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>%
  spread(Foodweb, absolute_fitness) %>%
  mutate(fitness_difference = Original-Extinction)

eggptoid_selection_plot <- eggptoid_plot_df %>%
  ggplot(., aes(x=sc.sqrt.Pref, y=fitness_difference, group=ID, alpha=ID_group, size=ID_group)) +
  geom_line(color=cbPalette[5]) +
  scale_alpha_manual(values=c(1,0.2), guide=FALSE) +
  scale_size_manual(values=c(1.5,0.25), guide=FALSE) +
  xlab("Gall midge preference (SDs)") + 
  ylab(expression(paste(Delta," egg parasitoid survival (original - extinction)"))) +
  scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels = c(-1,-0.5,0,0.5,1)) +
  geom_hline(yintercept=0, linetype="dotted") +
  theme_cowplot(font_size = 12)

eggptoid_selection_plot
save_plot(filename = "selection_on_Platygaster.pdf", plot = eggptoid_selection_plot, base_height = 4) # base_height = 6 is to large for PDF of manuscript
```