### 2020-03-11 manuscript/manuscript_clean.tex
incorporate final minor edits to accepted version.

\documentclass[11pt,]{article}
\usepackage[]{mathpazo}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            colorlinks=true,
            linkcolor=black,
            citecolor=black,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{amnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

%%% Modified from Latex Template for Am. Nat., many other details aren't necessary because they are specified in YAML header of Rmarkdown document.
\usepackage{fullpage}
\linespread{1.7}
\usepackage{lineno}

% code below is important for holding figure positions in main text. Make sure to set fig.pos="H" in code chunk for figure.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\vspace*{0.1cm}

\begin{center} \LARGE Loss of consumers constrains phenotypic evolution in the resulting food web \end{center}

\bigskip

\begin{center} \large Matthew A. Barbour$^{1,2,\ast}$, Christopher J. Greyson-Gaito$^{2,3}$, Arezoo Sootodeh$^{2}$, Brendan Locke$^{4}$, Jordi Bascompte$^{1}$ \normalsize \end{center}

\bigskip

\noindent 1. University of Zurich, Department of Evolutionary Biology
and Environmental Studies, Zurich, 8057 ZH, Switzerland;

\noindent 2. University of British Columbia, Department of Zoology,
Vancouver, BC V6T 1Z4, Canada;

\noindent 3. University of Guelph, Department of Integrative Biology,
Guelph, ON N1G 2W1, Canada;

\noindent 4. Humboldt State University, Department of Biological
Sciences, Arcata, CA 95521, USA.

\(^\ast\) Corresponding author; e-mail:
\href{mailto:matthew.barbour@ieu.uzh.ch}{\nolinkurl{matthew.barbour@ieu.uzh.ch}}

\bigskip

\emph{Short Running Title}: Consumer loss constrains phenotypic
evolution

\bigskip

\emph{Keywords}: fitness landscape, adaptive landscape, ecological
networks, eco-evolutionary dynamics, natural selection, host-parasitoid,
community context, extinction.

\bigskip

\emph{Abstract Word Count}: 256

\bigskip

\emph{Total Word Count}: 6119

\bigskip

\emph{Manuscript type}: Letter.

\bigskip

\linenumbers{} \modulolinenumbers[3]

\newpage

\section{Abstract}\label{abstract}

The loss of biodiversity is altering the structure of ecological
networks; however, we are currently in a poor position to predict how
these altered communities will affect the evolutionary potential of
remaining populations. Theory on fitness landscapes provides a framework
for predicting how selection alters the evolutionary trajectory and
adaptive potential of populations, but often treats the network of
interacting populations as a ``black box''. Here, we integrate
ecological networks and fitness landscapes to examine how changes in
food-web structure shape phenotypic evolution. We conducted a field
experiment that removed a guild of larval parasitoids that imposed
direct and indirect selection pressures on an insect herbivore. We then
measured herbivore survival as a function of three key phenotypic traits
to estimate directional, quadratic, and correlational selection
gradients in each treatment. We used these selection gradients to
characterize the slope and curvature of the fitness landscape to
understand the direct and indirect effects of consumer loss on
phenotypic evolution. We found that the number of traits under
directional selection increased with the removal of larval parasitoids,
indicating evolution was more constrained toward a specific combination
of traits. Similarly, we found that the removal of larval parasitoids
altered the curvature of the fitness landscape in such a way that tended
to decrease the evolvability of the traits we measured in the next
generation. Our results suggest that the loss of trophic interactions
can impose greater constraints on phenotypic evolution. This indicates
that the simplification of ecological communities may constrain the
adaptive potential of remaining populations to future environmental
change.

\newpage

\section{Impact Summary}\label{impact-summary}

The loss of biodiversity is rewiring the web of life; however, it is
uncertain how this will affect the ability of remaining populations to
evolve and adapt to future environments. To gain insight into these
effects, we conducted a field experiment that either maintained a
natural community of predators or removed all but one of the predators
that was able to impose selection on a common prey. We found that the
loss of predators acted to constrain prey evolution toward a particular
combination of traits. Moreover, we found that the loss of predators
could make it more difficult for prey to adapt to uncertain future
environments. Taken together, our results suggest that the
simplification of the web of life may constrain the adaptive potential
of remaining populations.

\newpage

\section{Introduction}\label{introduction}

The fitness landscape provides a powerful framework for understanding
how natural selection has shaped the evolution of biodiversity ---from
genes to phenotypes to species
\citep{Wright1931, Simpson1944, Arnold2001}. More than a metaphor, the
fitness landscape links quantitative genetic and phenotypic variation in
multiple traits to evolution by natural selection
\citep{Lande1979, Arnold1984applications, Arnold1984theory}. Ecological
interactions often play a key role in shaping natural selection, as
evidenced by the role of antagonistic and mutualistic interactions in
driving evolutionary change
\citep{Schluter2000, Abrams2000, Bronstein2006}. Although there is clear
evidence that species interactions can shape the fitness landscape, we
still have a poor understanding of how the fitness landscape is shaped
by a community of interacting species
\citep{McPeek2017, terHorst2018, Hui2018}. Resolution on how change in
ecological communities shape phenotypic evolution is urgently needed
though, given the rapid losses of biodiversity we are observing in the
Anthropocene \citep{Scheffers2016}.

Ecological networks, such as food webs describing who-eats-whom, provide
an explicit representation of the direct and indirect effects that
emerge in a community of interacting species
\citep{Bascompte2014, McCann2012}. Here, we integrate ecological
networks and fitness landscapes to understand how selection imposed by
ecological communities alter the evolutionary trajectory and adaptive
potential of interacting populations \citep{Hui2018}. The effects of
natural selection on multiple phenotypic traits can be inferred by
quantifying the slope and curvature of the fitness landscape
\citep{Arnold1992}. The slope is determined by the strength of
directional selection acting on each trait, which influences the
trajectory of evolution \citep{Lande1979, Arnold1992}. This fact is made
clear by the `Lande equation', \(\Delta\bar z = \text{G}\beta\), where
\(\Delta \bar z\) is a column vector of the average change in each trait
between generations, \text{G} is the additive genetic (co)variance
matrix for these traits (i.e.~G-matrix), and \(\beta\) is a column
vector of directional selection gradients acting on each trait
(i.e.~slope). Grant and Grant \citeyearpar{Grant1995} used the slope of
the fitness landscape for multiple traits related to body and beak size
in Darwin's finches to accurately predict the effects of drought on the
evolutionary trajectory of these traits. The curvature of the fitness
landscape is governed by the strength of stabilizing, disruptive, and
correlational selection acting on each trait, which can alter the
evolutionary potential of a population through its effect on the
G-matrix \citep{Hansen2008}. For example, stabilizing selection acts to
erode genetic variance in a trait, which can impose a constraint on the
ability of this trait to respond to future selection \citep{Hansen2008}.
In contrast, disruptive selection toward extreme trait values acts to
increase genetic variance in a trait, thus increasing the capacity for
future adaptation \citep{Bolnick2008}. Correlational selection alters
the genetic covariance between traits, which may facilitate or hinder
future adaptation depending on the pattern of selection on those traits
and the structure of the G-matrix \citep{Brodie1992}. If we want to
predict how ecological communities shape phenotypic evolution, we must
understand how ecological networks shape the fitness landscape of
interacting populations.

The loss of biodiversity is altering the structure of ecological
networks \citep{Landi2018ComplexityStability}, which may influence the
slope and curvature of the fitness landscape in a number of ways. For
example, consider how changes in consumer diversity in a food web may
alter the slope of the fitness landscape for a shared resource. If
different consumers impose directional selection on different traits of
the resource, then a more diverse consumer community would increase the
number of traits under selection, which may constrain the trajectory of
evolution toward a specific phenotype. Alternatively, if consumers
impose selection on different values of a trait, then their selective
effects would cancel each other out in more diverse communities, which
would allow for a greater diversity of phenotypes to persist. Now
consider the effects of consumer diversity on the curvature of the
fitness landscape. If consumers impose selection on different ends of a
resource's trait distribution, then a more diverse consumer community
may impose stabilizing selection, which would decrease genetic variance
in this trait. In contrast, additional consumers may impose disruptive
selection if their cumulative effect decreases the relative fitness of a
resource's average trait value, which would increase the genetic
variance in this trait.

To examine these different possibilities, we conducted a field
experiment that removed a consumer guild that parasitizes an abundant
insect herbivore (\emph{Iteomyia salicisverruca}, Family Cecidomyiidae;
fig. \ref{fig:Conceptual}). The larvae of this herbivore induce
tooth-shaped galls when they feed on the developing leaves of willow
trees \citep[\emph{Salix} sp.,][]{Russo2006}. These galls protect larva
from attack by generalist predators (e.g.~ants, spiders), but they
suffer high mortality from egg and larval parasitoids
\citep{Barbour2016}. We manipulated food-web structure by either
removing larval parasitoids (removal food web) or allowing both egg and
larval parasitoids to impose selection on gall midge traits (original
food web, fig. \ref{fig:Conceptual}). Larval parasitoids also impose
indirect effects on gall midge fitness through intraguild predation on
the egg parasitoid (fig. \ref{fig:Conceptual}). We applied modern
statistical methods to quantify how changes in food-web structure
altered the slope and curvature of the gall midge's fitness landscape.
Taken together, our study gives insight to how the loss of biodiversity
may alter both the evolutionary trajectory and adaptive potential of
interacting populations.

\begin{figure}
\centering
\includegraphics{../analyses/complex_simple_foodwebs_v4_5.jpg}
\caption{\label{fig:Conceptual}Experimental manipulation of food-web
structure associated with a leaf-galling midge (b, \emph{Iteomyia
salicisverruca}) feeding on the willow \emph{Salix hookeriana} (a).
Black arrows denote the flow of energy in this network of trophic
interactions. In the original food web, we allowed the full suite of egg
and larval parasitoids to impose selection. In the removal food web, we
used mesh bags to exclude the guild of larval parasitoids, only allowing
the egg parasitoid (c, \emph{Platygaster} sp.) to impose selection. Note
that larval parasitoids also impose indirect effects on gall midge
fitness through intraguild predation on the egg parasitoid. Larval
parasitoids include the following species: \emph{Mesopolobus} sp. (d,
Family: Pteromalidae); \emph{Tetrastichus} sp. (e, Family: Eulophidae);
and \emph{Torymus} sp. (f, Family: Torymidae).}
\end{figure}

\section{Methods}\label{methods}

\subsection{Study Site}\label{study-site}

We conducted our study within a four-year old common garden experiment
of coastal willow (\emph{Salix hookeriana}) located at Humboldt Bay
National Wildlife Refuge (HBNWR) (40 40'53``N, 124 12'4''W) near Loleta,
California, USA. This common garden consists of 26 different willow
genotypes that were collected from a single population of willows
growing around Humboldt Bay. Stem cuttings of each genotype (25
replicates per genotype) were planted in a completely randomized design
in two hectares of a former cattle pasture at HBNWR. Willows at our
study site begin flowering in February and reach their peak growth in
early August. During this study, willows had reached 5 - 9m in height.
Further details on the genotyping and planting of the common garden are
available in \citet{Barbour2015}.

\subsection{Manipulating Food-web
Structure}\label{manipulating-food-web-structure}

We setup our food-web manipulation across 128 plants soon after galls
began developing on willows in early June of 2013. These 128 plants came
from 8 different plant genotypes that spanned the range of trait
variation observed in this willow population \citep{Barbour2015}. For
the original food web (eight replicates per genotype), we used flagging
tape to mark 14 galled leaves per plant (\textasciitilde{}30 larvae),
allowing the full suite of egg and larval parasitoids to impose
selection. Marking galls with flagging tape ensured that we compared
galls with similar phenology in both treatments when we collected galls
later in the season. For the removal food web, we enclosed 14 galled
leaves with 10x15cm organza bags (ULINE, Pleasant Prairie, WI, USA) to
exclude three parasitoid species that attack during larval development.
This treatment did not exclude the egg parasitoid \emph{Platygaster}
sp., which attacks prior to gall initiation \citep[larva initiate gall
development in Cecidomyiid midges:][]{Gagne1989}. It was not possible to
reciprocally manipulate parasitoid attack (i.e.~exclude egg parasitoid,
but allow larval parasitoids), because it was not possible to identify
midge oviposition sites prior to gall formation. In late August, we
collected marked and bagged galls from each plant, placed them into 30
mL vials and kept them in the lab for 4 months at room temperature. We
then opened galls under a dissecting scope and determined whether larvae
survived to pupation (our measure of fitness) or died due to parasitism.
We further distinguished whether mortality was caused by an egg or
larval parasitoid. Early larval death was another important source of
mortality (17\%), but we excluded these cases from our analysis. We did
this because early larval death can stunt gall growth, which would
confound estimates of selection due to parasitism on one of the
phenotypes we measured (chamber diameter; details given in next
section). For the food-web treatment that excluded larval parasitoids,
we further restricted our data by removing any incidental instances of
parasitism by a larval parasitoid. This represented less than 3\% of the
observations in this food-web treatment and allowed us to focus our
inferences of selection on those imposed by the egg parasitoid. Our
final dataset contains survival estimates for 1285 larvae from 613 galls
and 111 plants.

\subsection{Measuring Phenotypic
Traits}\label{measuring-phenotypic-traits}

We collected data on three different traits that we expected to
influence larval survival based on previous work in this system
\citep{Barbour2016} and other work with gall midges
\citep{Weis1983, Heath2018}. First, we measured gall diameter as the
size of each gall chamber to the nearest 0.01 mm at its maximum diameter
(perpendicular to the direction of plant tissue growth). Previous work
in this system has shown that larger galls are associated with higher
survival \citep{Barbour2016}. Second, we measured clutch size by
counting the number of chambers in each gall
\citep{Weis1983, Heath2018}. All larvae collected from the same
multi-chambered gall were scored with the same clutch size. Third, we
measured oviposition preference for individual plants as the density of
larvae observed on a plant in an independent survey. We did this by
randomly sampling five branches per tree and counting the number of
individual gall chambers (number of larvae). We then converted these
counts to a measure of larval density per 100 shoots by counting the
number of shoots on the last branch we sampled. All larvae collected
from the same plant were scored with the same oviposition preference.
Measuring larval densities on plants in the field is a common method for
measuring oviposition preference \citep{Gripenberg2010}; however,
caution must be taken in inferring `preference' as larval densities can
be influenced by processes other than preference \citep{Singer1986}.
Fortunately, two features of our study system suggest that larval
densities may be a good proxy for oviposition preference. First, since
our data comes from a randomized placement of plant genotypes in a
common garden, females midges have equal exposure to many possible plant
genotypes when choosing an oviposition site. Second, egg predation is a
minor source of mortality for galling insects in general
\citep{Hawkins1997}; therefore, we do not expect any prior egg predation
to bias our estimates of observed larval densities.

\subsection{Quantifying the Fitness
Landscape}\label{quantifying-the-fitness-landscape}

Inferring the fitness landscape assumes that trait distributions are
multivariate normal \citep{Lande1983}. To approximate this assumption,
we log-transformed clutch size and square-root transformed oviposition
preference. Chamber diameter already had a normal distribution so we did
not transform it. We then scaled all phenotypic traits (mean=0 and SD=1)
across treatments prior to our analyses (detailed below) to ensure that
our estimates of selection were comparable across traits and with other
studies.

Our analysis consisted of four parts. First, we used generalized linear
mixed models (GLMM) to quantify selection surfaces \textemdash linear
and nonlinear relationships between absolute fitness (\(W\)) and
standardized phenotypic traits (\(i\)) of individuals \textemdash in
each food-web treatment. Second, we scaled selection surfaces relative
to mean fitness (\(\bar W\)) in order to calculate standardized
selection gradients. Third, we used our estimates of directional
selection gradients to characterize the slope of the fitness landscape,
which we used to quantify the effects of food-web structure on the
trajectory of evolution. Finally, we estimated the curvature of the
fitness landscape and used a simulation to explore its effects on the
adaptive potential of the gall midge population in the next generation.

\textbf{Selection surfaces}: Since larval survival was our measure of
absolute fitness, we used a GLMM that assumed a binomial error
distribution (and logit-link function). To approximate the selection
surface, we modelled larval survival as a function of food-web treatment
as well as linear (\(\alpha_i\)), quadratic (\(\alpha_{ii}\)), and
statistical interactions (\(\alpha_{ij}\)) between each trait. Note that
to obtain valid estimates of linear selection surfaces, we removed
nonlinear terms prior to estimating linear relationships
\citep{Lande1983}. Other approaches have been advocated for
approximating selection surfaces \citep{Schluter1988}; however, our
approach enables us to calculate selection gradients, and thus is more
appropriate for approximating the fitness landscape \citep{Arnold2003}.
To account for the nonindependence of clutch size (measured at gall
level) and oviposition preference (measured at plant level) as well as
any independent effects of willow genotype on larval survival, we
modelled gall identity nested within plant identity nested within
genotype identity as random effects. Although statistical models with
random effects are not common in analyses of natural selection, we think
that modelling random effects can mitigate biased estimates of selection
due to environmental covariances between traits and fitness
\citep{Rausher1992}. Since our end goal was to characterize the
relationship between mean trait values and mean fitness (fitness
landscape), we assumed the mean value of our random effects
(i.e.~setting them to zero) when calculating selection surfaces. We then
used parametric bootstrapping (1,000 replicates) to estimate the effect
of food-web treatment on larval survival as well as selection surfaces
in each food-web treatment. To determine whether trait-fitness
relationships differed between food-web treatments, we calculated the
difference in bootstrapped replicates between treatments.

\textbf{Selection gradients}: Selection gradients cannot be estimated
directly from GLMMs of selection surfaces because the response is in
terms of absolute fitness and the coefficients are on a nonlinear scale.
For example, the coefficients in the previously described binomial model
measure the change in the log-odds of survival associated with 1 SD
change in a trait with all other traits held fixed at their mean.
Therefore, we used the method developed by \citet{Janzen1998} to
translate selection surfaces from the above model into the scale of
relative fitness in order to calculate directional (\(\beta_i\)),
quadratic (\(\gamma_{ii}\)), and correlational (\(\gamma_{ij}\))
selection gradients. Briefly, this method calculates the average
gradient of selection surfaces by multiplying the average of
\(W(z)[1-W(z)]\) by each regression coefficient (e.g. \(\alpha_i\),
\(\alpha_{ii}\), or \(\alpha_{ij}\)). We then divided this average
gradient by the mean fitness (\(\bar W\)) to put it on the scale of
relative fitness (\(w=W/\bar W\)), and thus interpretable as a selection
gradient. We estimated selection gradients separately for each food-web
treatment. We also determined whether selection gradients differed
between food-web treatments by calculating the difference in
bootstrapped replicates between treatments. Note that we doubled all
quadratic terms prior to calculating selection gradients to put them on
the same scale as estimates of directional and correlational selection
\citep{Stinchcombe2008}.

\textbf{Evolutionary trajectory}: The effect of selection on the
trajectory of evolution is determined by the slope of the fitness
landscape, which in our study corresponds to:

\[\textbf{Slope} = \begin{pmatrix} \beta_{\text{Diam}} \\ \beta_{\text{Clutch}} \\ \beta_{\text{Pref}} \end{pmatrix} \]
where each \(\beta_i\) corresponds to the directional selection gradient
acting on each trait. By comparing selection gradients in each food-web
treatment (i.e.~95\% CI does not overlap zero), we can infer the effect
of food-web structure on the trajectory of phenotypic evolution.

\textbf{Evolutionary potential}: The indirect effects of selection on
the G-matrix (\(\Delta\text{G}=\text{G}\text{C}\text{G}\)) are governed
by the curvature of the fitness landscape
(\(\text{C}=\gamma - \beta \beta^\text{T}\)), which in our study
corresponds to:

\[\textbf{Curvature} = \begin{pmatrix} \gamma_{\text{Diam:Diam}}&& \\ \gamma_{\text{Clutch:Diam}}&\gamma_{\text{Clutch:Clutch}}& \\ \gamma_{\text{Pref:Diam}} & \gamma_{\text{Pref:Clutch}} &\gamma_{\text{Pref:Pref}} \end{pmatrix} - \begin{pmatrix} \beta_{\text{Diam}}\beta_{\text{Diam}}&& \\ \beta_{\text{Clutch}}\beta_{\text{Diam}}&\beta_{\text{Clutch}}\beta_{\text{Clutch}}& \\ \beta_{\text{Pref}}\beta_{\text{Diam}} & \beta_{\text{Pref}}\beta_{\text{Clutch}} &\beta_{\text{Pref}}\beta_{\text{Pref}} \end{pmatrix}\]

where each \(\gamma_{ii}\) (diagonal) corresponds to the quadratic
selection gradient acting on a trait, and each \(\gamma_{ij}\)
(off-diagonal) corresponds to the correlational selection gradient
acting on a particular trait combination. Note that we ommitted the
upper triangle of each matrix for clarity since it is simply the
reflection of the lower triangle. Subtracting these two matrices results
in the curvature matrix of the fitness landscape:

\[\textbf{Curvature} = \begin{pmatrix} \text{C}_{\text{Diam:Diam}}&& \\ \text{C}_{\text{Clutch:Diam}} & \text{C}_{\text{Clutch:Clutch}} & \\ \text{C}_{\text{Pref:Diam}} & \text{C}_{\text{Pref:Clutch}} & \text{C}_{\text{Pref:Pref}} \end{pmatrix}\]
where each \(\text{C}_{ii}\) (diagonal) represents the effect of
selection on the additive genetic variance in a trait, and each
\(\text{C}_{ij}\) (off-diagonal) represents the effect of selection on
the additive genetic covariance between a particular trait combination.
We used bootstrapped values of each selection gradient to estimate the
curvature of each component of the matrix and its associated 95\% CI. We
also used this information to determine whether the curvature of each
component differed between our food-web treatments.

Knowledge of the curvature matrix alone gives an incomplete picture of
its indirect effect on evolutionary potential. This is because the
evolutionary potential of a population is ultimately determined by the
structure of its G-matrix, and therefore also depends on its structure
before selection. Although we do not know the underlying G-matrix for
the traits we measured in this experiment, we can still gain insight to
how our food-web treatment would alter genetic constraints more
generally. Specifically, we calculated how our best estimate of the
curvature matrix (mean values) in each treatment changed the structure
of \ensuremath{10^{4}} random G-matrices for the next generation. We
restricted the range of additive genetic variance
(\emph{V\textsubscript{G}}) for each element in these G-matrices to
between 0 and 0.5 to reflect typical ranges in narrow-sense heritability
values (\emph{h}\textsuperscript{2}; note that
\emph{h}\textsuperscript{2} = \emph{V\textsubscript{G}} when the
phenotypic variance is scaled to 1). Note that this analysis assumes
that the effects of recombination and mutation on the G-matrix are much
smaller than the effects of selection, which appears to be a reasonable
assumption over short time scales \citep{Arnold2008}.

The G-matrix itself is a complex structure, but has a clear theoretical
link to the evolutionary potential, or evolvability of phenotypic traits
\citep{Hansen2008}. Evolvability measures the ability for a trait to
evolve toward a given direction of selection \citep{Hansen2008}. In the
absence of knowledge about the direction of selection that a population
will actually experience in the next generation, we can compute its
average evolvability over random directional selection gradients
\citep{Hansen2008, Melo2015}. By computing the average evolvability
(here, we used 1000 random \(\beta\)s) for each of \ensuremath{10^{4}}
random G-matrices, we can then assess how changes in the curvature
matrix alter the evolutionary potential of the associated traits. We
report the distribution of these effect sizes, rather than conduct a
statistical test, because the number of replicates in a simulation can
be arbitrarily high, thus making any effect size statistically
significant \citep{White2014}.

\subsection{Adjusting for biased measurements of
selection}\label{adjusting-for-biased-measurements-of-selection}

Rather than imposing selection, parasitoids may influence the expression
of herbivore traits which could bias measurements of selection. In our
system, it was plausible that parasitoids may influence chamber diameter
by altering larval feeding behavior or killing larvae before they
complete their development. To estimate this potential bias, we subset
our data to only include galls where there was variation in larval
survival within the same gall (i.e.~1 \textgreater{} survival
\textgreater{} 0). If we assume that larvae within each gall should have
similar chamber diameters because they come from the same clutch and
experience the same local environment (an assumption our data supports:
gall identity explains 54\% of the variance in chamber diameter), then
the relationship between chamber diameter and larval survival in this
data subset represents the effect of parasitism on trait expression
(i.e.~bias). We used a GLMM with the same structure as described
previously except that we modeled only a linear relationship between
chamber diameter and larval survival (\(\alpha_{\text{Diam}}\)). We
detected a positive bias in both food-web treatments (original
\(\alpha_{\text{Diam}}\)= 0.36 {[}0.05, 0.67{]}; removal
\(\alpha_{\text{Diam}}\)= 0.42 {[}0.01, 0.82{]}), indicating that
unadjusted relationships would overestimate the strength of selection on
chamber diameter. To account for this bias, we subtracted our mean
estimates of bias from our estimates with the full dataset prior to
calculating chamber diameter's selection surface and directional
selection gradient.

\subsection{Measuring selection on egg
parasitoids}\label{measuring-selection-on-egg-parasitoids}

Once parasitized, the gall phenotype becomes the phenotype of the egg
parasitoid. This phenotype may influence the egg parasitoid's survival
in the face of larval parasitoids, and thus experiences selection. Our
food-web manipulation allows us to measure selection imposed by larval
parasitoids on the phenotype of egg parasitoids. Using the same models
as described above, we substituted egg parasitism as our response
variable to quantify selection surfaces and selection gradients acting
on the egg parasitoid. Note that we cannot test the effect of food-web
structure on the egg parasitoid's fitness landscape ---we can only
estimate selection imposed by larval parasitoids. This comparison is
still useful though in determining the extent to which the loss of
consumers may have indirect evolutionary effects by altering selection
on multiple interacting populations.

All analyses and visualizations were conducted in R \citep{R2018}.
Unless otherwise noted, we report mean estimates of selection surfaces
and selection gradients with 95\% confidence intervals in brackets. Note
that for visualizing the fitness landscape we restrict trait axes to
\(\pm 1\) SD of the mean trait value. This emphasizes the fact that we
can only reliably estimate the shape of the fitness landscape near the
mean phenotype of the population \citep{Arnold2001}. We also plot mean
larval survival on a natural log scale to accurately reflect the
mathematical definition of the fitness landscape \citep{Arnold2003}. All
data and code to reproduce the reported results are publicly available
on GitHub (\url{https://github.com/mabarbour/complexity_selection}) and
have been archived on Zenodo
(\url{https://zenodo.org/badge/latestdoi/108833263}).

\section{Results}\label{results}

\subsection{Consumer removal constrains the evolutionary trajectory of
gall
midges}\label{consumer-removal-constrains-the-evolutionary-trajectory-of-gall-midges}

\indent We found that the removal of larval parasitoids increased the
number of gall midge traits under directional selection (3 of 3)
relative to the original food web (1 of 3)(table \ref{Table:Gradients}).
For example, we observed directional selection for smaller clutch sizes
when we removed larval parasitoids, but there was no evidence of
selection acting on this trait in the original food web (fig.
\ref{fig:UV_Landscape}A). This absence of selection appeared to be a
result of conflicting selection pressures imposed by each guild of
parasitoids. Specifically, when we subset our data to focus on
differences between parasitoid guilds, we found that larval parasitoids
actually impose directional selection for larger clutch sizes (larval
parasitoids \(\beta_{\text{Clutch}}\)= 0.13 {[}0.04, 0.24{]}). This
conflicting selection is likely due to trait differences between guilds,
as larger clutches may be easier targets for egg parasitoids to find,
but the more complex gall structure may be more difficult for larval
parasitoids to oviposit through.

\bigskip

\begin{table}[h]
\caption{Standardized selection gradients acting on gall midges in the original food web and with the removal of larval parasitoids.}
\label{Table:Gradients}
\centering
\begin{tabular}{lccc}
\\ 
\hline
\textbf{Selection gradient} & \textbf{Original} & \textbf{Removal} & \textbf{Contrast} (\small{Original - Removal}) \\
\hline
$\beta_{\text{Diam}}$ & 
\textbf{
0.34 [
0.22,
0.48] } & 
\textbf{
0.21 [
0.12,
0.31] } & 
\textbf{
0.14 [
0,
0.27] }\\

$\beta_{\text{Clutch}}$ & 
0.06 [
-0.05,
0.17] & 
\textbf{
-0.09 [
-0.17,
-0.01] } & 
\textbf{
0.15 [
0.03,
0.29] }\\

$\beta_{\text{Pref}}$ &
-0.13 [
-0.29,
0.05] & 
\textbf{
-0.16 [
-0.26,
-0.06] } &

0.03 [
-0.15,
0.21] \\

$\gamma_{\text{Diam:Diam}}$ &
0.13 [
-0.06,
0.33] &

0.1 [
-0.02,
0.23] &

0.03 [
-0.2,
0.27] \\

$\gamma_{\text{Clutch:Clutch}}$ & 
-0.05 [
-0.27,
0.18] &

-0.11 [
-0.28,
0.03] &

0.06 [
-0.2,
0.32] \\

$\gamma_{\text{Pref:Pref}}$ & 
\textbf{
0.34 [
0.07,
0.63] }&

0.02 [
-0.15,
0.18] & 
\textbf{
0.32 [
0.01,
0.64] }\\

$\gamma_{\text{Diam:Clutch}}$ & 
-0.04 [
-0.16,
0.08] &

-0.07 [
-0.15,
0.02] &

0.02 [
-0.12,
0.17] \\

$\gamma_{\text{Diam:Pref}}$ & 
-0.13 [
-0.29,
0.02] &

-0.02 [
-0.1,
0.07] &

-0.12 [
-0.3,
0.05] \\

$\gamma_{\text{Clutch:Pref}}$ & 
0.03 [
-0.1,
0.18] &

0 [
-0.07,
0.07] &

0.03 [
-0.12,
0.18] \\ 
\hline
\end{tabular}
\bigskip{}
\\
{\footnotesize Note: Values in brackets represent 95\% confidence intervals. Bold values indicate that the 95\% CI does not overlap zero. $\beta_{\text{Diam}}$ has been adjusted for bias.}
\end{table}

We also observed clear evidence of directional selection for midges to
avoid ovipositing on plants with high densities of conspecifics when we
removed larval parasitoids (fig. \ref{fig:UV_Landscape}B); however, the
overall magnitude of selection did not differ between treatments (table
\ref{Table:Gradients}). Still, there was no clear evidence of
directional selection on oviposition in the original food web (table
\ref{Table:Gradients}). Chamber diameter experienced positive
directional selection in both food-web treatments (fig.
\ref{fig:UV_Landscape}C). Although the magnitude of selection on chamber
diameter was relatively higher in the original food web (table
\ref{Table:Gradients}), this was not due to any difference in the
relationship between chamber diameter and survival (selection surfaces:
contrast \(\alpha_{\text{Diam}}\)= 0.04 {[}-0.5, 0.55{]}), but was
simply a consequence of the (expected) lower survival of gall midges in
the original food web (contrast \(\bar W\)= 0.27 {[}0.11, 0.42{]}). We
expect this difference to be transient though, since egg parasitoids
would increase in abundance once they are released from intraguild
predation, which would make the strength of selection on gall diameter
comparable to the original food web (if removal \(\bar W\)=original
\(\bar W\), then contrast \(\beta_{\text{Diam}}\)= -0.06 {[}-0.2,
0.1{]}). It is worth noting that positive selection on chamber diameter
in both treatments was unexpected. For example, the fact that larger
galls provide more of a refuge from larval parasitoids makes sense since
they attack after the gall is formed; however, egg parasitoids attack
prior to gall formation, which suggests that chamber diameter is either
directly related to survival or strongly correlated with an unmeasured
trait that is under selection (e.g.~immune response).

\bigskip

\begin{figure}
\centering
\includegraphics{../analyses/UV_landscapes.pdf}
\caption{\label{fig:UV_Landscape}Adaptive landscape of gall midge
phenotypes in the original food web and with the removal of larval
parasitoids. Each panel corresponds to a different phenotypic trait:
clutch size (A); oviposition preference (B); and chamber diameter (C).
Bold lines represent selection experienced in the original (orange) and
removal (blue) food webs. Thin lines represent bootstrapped replicates
to show the uncertainty in selection. For clarity, we only display 100
bootstraps even though inferences are based on 1,000 replicates. Note
that mean larval survival is plotted on a natural log scale to reflect
the mathematical definition of the \textbf{fitness} landscape.}
\end{figure}

\bigskip

To visualize the multivariate effects of natural selection, we plotted
the fitness landscape for each trait combination in each treatment (fig.
\ref{fig:MV_Landscape}). The arrows in fig. \ref{fig:MV_Landscape}
represent mean estimates of directional selection gradients, while the
dotted lines represent predicted survival of the mean phenotype in each
food-web treatment. Notice that arrows point more toward a corner of the
fitness landscape for each combination of traits with the removal of
larval parasitoids compared to the original food web. This indicates
that the removal of consumers more strongly favored a specific
combination of traits, rather than allowing for multiple trait
combinations to have comparable fitness.

\begin{figure}
\centering
\includegraphics{../analyses/MV_landscapes.pdf}
\caption{\label{fig:MV_Landscape}Two dimensional view of
\textbf{fitness} landscapes of gall midge phenotypes in the original
food web and with the removal of larval parasitoids. Each panel
corresponds to a different combination of phenotypic traits: clutch size
and chamber diameter (A); clutch size and oviposition preference (B);
oviposition preference and chamber diameter (C). Arrows represent mean
estimates of directional selection gradients, while
\textbf{dotted lines} represent predicted larval survival of the mean
phenotype in each food-web treatment. Notice that arrows point more
toward a corner of the \textbf{fitness} landscape for each combination
of traits with the removal of larval parasitoids compared to the
original food web. This indicates that the removal of consumers more
strongly favored a specific combination of traits. Note that mean larval
survival is plotted on a natural log scale to reflect the mathematical
definition of the \textbf{fitness} landscape.}
\end{figure}

\bigskip

\subsection{Consumer removal constrains the evolutionary potential of
gall
midges}\label{consumer-removal-constrains-the-evolutionary-potential-of-gall-midges}

\indent The curvature of the fitness landscape indirectly affects
evolutionary potential and is influenced by directional, quadratic, and
correlational selection gradients. There was no clear evidence of
correlational selection for any combination of traits in either food-web
treatment (table \ref{Table:Gradients}). Similarly, there was no clear
evidence of quadratic selection on chamber diameter or clutch size in
either food-web treatment (table \ref{Table:Gradients}). In contrast,
our food-web treatment did alter quadratic selection acting on
oviposition preference (table \ref{Table:Gradients}). In particular, we
observed disruptive selection in the original food web, with selection
favoring females that either avoided high densities (\textless{}0.5 SD
above mean density), or if gall densities were high enough
(\textgreater{}0.5 SD above mean), then selection favored females that
were attracted to high densities (fig. \ref{fig:UV_Landscape}B). This
nonlinear relationship was partly due to a trend for disruptive
selection imposed by larval parasitoids (\(\gamma_{\text{Pref:Pref}}\)=
0.18 {[}-0.02, 0.42{]}), but was also magnified by the lower average
survival in the original food web. This was likely a result of larval
parasitoids imposing greater mortality on egg parasitoids at high gall
midge densities (see \textbf{Selection on egg parasitoids} section),
which would act to increase the relative fitness of gall midges at high
densities.

Using our estimates of directional (\(\beta_i\)), quadratic
(\(\gamma_{ii}\)), and correlation selection (\(\gamma_{ij}\)), we
calculated the curvature (\(\text{C}=\gamma - \beta \beta^\text{T}\)) of
the fitness landscape in each food-web treatment.

\[\textbf{C} = \begin{pmatrix} \text{C}_{\text{Diam:Diam}}&& \\ \text{C}_{\text{Clutch:Diam}} & \text{C}_{\text{Clutch:Clutch}} & \\ \text{C}_{\text{Pref:Diam}} & \text{C}_{\text{Pref:Clutch}} & \text{C}_{\text{Pref:Pref}} \end{pmatrix}\]

Of the different components of the curvature matrix, we found that only
the curvature of oviposition preference differed between food-web
treatments. Specifically, selection in the removal food web acted to
decrease the additive genetic variance in preference relative to the
original food web.

\[\textbf{C}_{\text{Original}} = \begin{pmatrix} 
-0.05\: [-0.15,0.05] &  &  \\  
-0.06\: [-0.18,0.06] & -0.03\: [-0.14,0.09] &  \\  
-0.09\: [-0.25,0.06] & 0.04\: [-0.1,0.18] & \textbf{0.15\: [0.02,0.3]} \end{pmatrix}\]

\[\textbf{C}_{\text{Removal}} = \begin{pmatrix} 
0.01\: [-0.05,0.07] &  &  \\  
-0.05\: [-0.13,0.03] & -0.06\: [-0.15,0.01] &  \\  
0.02\: [-0.07,0.1] & -0.01\: [-0.09,0.06] & -0.02\: [-0.1,0.06] \end{pmatrix}\]

\[\textbf{C}_{\text{Removal-Original}} = \begin{pmatrix} 
0.06\: [-0.06,0.18] &  &  \\  
0.02\: [-0.13,0.16] & -0.03\: [-0.16,0.1] &  \\  
0.11\: [-0.06,0.29] & -0.05\: [-0.21,0.1] & \textbf{-0.17\: [-0.33,-0.01]} \end{pmatrix}\]

Interestingly, when we translate the effect of the curvature matrix onto
evolutionary potential in the next generation, we see that the removal
of consumers had a general tendency to decrease evolvability (fig.
\ref{fig:Evolvability}). Specifically, the removal food web decreased
the average evolvability of 71\% of the random G-matrices in our
simulation. If anything, we expect that this underestimates the true
effect of our removal treatment. For example, if we assume egg
parasitoids would eventually impose similar impact on mean fitness once
they are released from intraguild predation (i.e.~removal
\(\bar W\)=original \(\bar W\)), then the removal food web decreases the
average evolvability in 78\% of the G-matrix scenarios.

\bigskip

\begin{figure}
\centering
\includegraphics{../analyses/delta_evolvability.pdf}
\caption{\label{fig:Evolvability}Change in average evolvability for
10,000 random G-matrices using our best (mean) estimate of the curvature
matrix for each food-web treatment. We found that the curvature of the
removal food web decreased evolvability in 71\% of the G-matrices
(i.e.~the change in evolvability was negative for 71\% of the
simulations), suggesting that the removal of consumers tended to
decrease evolutionary potential of traits in our study.}
\end{figure}

\bigskip

\subsection{Selection on egg
parasitoids}\label{selection-on-egg-parasitoids}

The removal of larval parasitoids not only had direct effects on gall
midge fitness, but also imposed indirect effects that would be felt in
the next generation. For example, the removal of larval parasitoids
altered the relationship between gall midge preference and the
probability of observing egg parasitoids
(\(\gamma_{\text{Pref:Pref}}\)=-0.46 {[} -1.07, -0.02{]}, table S1),
such that the impact of larval parasitoids increased nonlinearly with
higher gall midge densities (fig. \ref{fig:EggPtoid_Selection}). This
may indicate that larval parasitoids switch to target galls that have
been parasitized by an egg parasitoid once egg parasitoid densities are
high enough. This prey switching behavior would contribute to the
disruptive selection we observed on gall midge preference by increasing
the relative fitness of gall midges at high densities (fig.
\ref{fig:UV_Landscape}B).

\bigskip

\begin{figure}
\centering
\includegraphics{../analyses/selection_on_Platygaster.pdf}
\caption{\label{fig:EggPtoid_Selection}Selection imposed by larval
parasitoids on egg parasitoids (\emph{Platygaster} sp.). The bold line
represents the average difference in the probability of observing the
egg parasitoid (original minus removal of larval parastioids) as a
function of gall midge oviposition preference. Thin lines represent
bootstrapped replicates to show the uncertainty in selection. For
clarity, we only display 100 bootstraps even though inferences are based
on 1,000 replicates. The decrease in the probability of observing egg
parasitoids at high gall-midge densities indicate that larval
parasitoids impose nonlinear selection on egg parasitoids.}
\end{figure}

\bigskip

\section{Discussion}\label{discussion}

We found that the removal of larval parasitoids constrained phenotypic
evolution in gall midges in two key ways. First, we observed directional
selection on more traits in the absence of larval parasitoids,
suggesting greater constraints on the trajectory of phenotypic
evolution. Second, removing larval parasitoids altered the curvature of
the fitness landscape in such a way that tended to decrease the
evolvability of associated traits. Assuming these traits have other
ecological functions, then this decrease in evolvability could constrain
the gall midge's adaptive potential in the face of novel selection
pressures. Our experiment also revealed evidence of indirect selection
pressures, suggesting that the loss of consumers may have complex
effects on the trajectories of phenotypic evolution. Taken together, our
study provides experimental evidence from the field that the loss of
consumers may constrain the adaptive potential of remaining populations.

The generality of our results likely depends on the relative abundance
and functional differences between consumers in a community. For
example, if consumers do not differ from each other, then we do not
expect the loss of consumers to modify selective constraints. Also, many
consumers may be at too low of abundances to impose selection on their
resources. Rank abundance curves \citep{Preston1948} and the
disproportionate number of weak interactions in diverse communities
\citep{Paine1992} support this notion. This logic suggests that we may
not have observed the effects we did if we had only removed one larval
parasitoid, because each species had relatively low abundance
\citep{Barbour2016} and they likely share similar ecological roles. When
consumers are functionally different and abundant though, the effect of
consumer loss will depend on whether different consumers impose
conflicting selection pressures or select for distinct traits. When
consumers impose conflicting selection on traits, as in our study and
others \citep{Weis1985, Abrahamson1997, Start2016, Start2018}, then
consumer diversity acts to neutralize selection, which can maintain
larger evolvability. On the other hand, different consumers may impose
selection on different traits; therefore, a more diverse consumer
community may favor a particular combination of traits and increase
selective constraints. Examples of this include strong genetic
covariances in plant resistance to different insect herbivores
\citep{Maddox1990, Wise2007, Wise2013}, although there are also examples
where these covariances are weak \citep{Roche1997, Barbour2015}, or vary
from year-to-year \citep{Johnson2007}. We suggest that gaining
predictive insight to the evolutionary consequences of food-web
disassembly requires an understanding of the mechanisms governing the
assembly of trophic interactions \citep{Bascompte2009}.

We also found evidence for a general decrease in trait evolvability when
we excluded larval parasitoids due to changes in the curvature of the
fitness landscape. This result was driven by disruptive selection on
gall midge oviposition preference in the original food web, which was
likely due to both increases in intraguild predation on egg parasitoids
(i.e.~prey switching) and the lower mean survival of gall midges. This
pattern of selection acts to increase genetic variation in oviposition
preference, which in turn enhances the ability of the gall midge
population to respond to uncertain selection pressures in the next
generation (i.e.~evolvability). This pattern of selection also indicates
the possibility of evolutionary bistability, where different initial
conditions (e.g.~mean oviposition preference) select for different
phenotypes of the same species in similar environments
\citep{Landi2018LifeHistory}. Thus, this pattern of selection may
contribute to genetic and phenotypic diversity both within and among
populations. Current theory, however, often assumes genetic variances
and covariances remain constant over time and space rather than
dynamically changing with the network context
\citep{McPeek2017, Guimaraes2017, Medeiros2018}. Our empirical results
highlight the need to explore the evolutionary consequences of not only
direct effects of selection, but indirect effects on genetic constraints
that emerge in a network of interacting species.

An important caveat of our study is that we did not do a factorial
manipulation of both parasitoid guilds, making it difficult to conclude
whether our results would change if we manipulated the presence/absence
of the dominant egg parasitoid. If we assume that higher-order
interactions \citep{Levine2017} are weak between parasitoid guilds, then
we can gain insight to how the loss of the egg parasitoid would alter
selection by isolating the contribution of larval parasitoids to
selection in our original food-web treatment. When we do this, we see
the same qualitative effects as we do when we removed larval
parasitoids. For example, we see clear evidence of all three traits
being under directional selection (i.e.~greater constraints on the
trajectory of evolution, table S2) as well as a decrease, albeit
smaller, in trait evolvability under different G-matrix scenarios (57\%,
fig. S1). This suggests that our results could be robust to this caveat,
which was simply not possible to manipulate given the biology of our
system (see \textbf{Manipulating Food-web Structure} section for
explanation).

Our results suggest that the loss of consumers may not only directly
affect connected species, but also result in indirect evolutionary
effects. In our study, this indirect effect arises from egg parasitoids
being released from intraguild predation when we excluded larval
parasitoids. This release occurs more on trees with high larval
densities, which could intensify future selection on gall midge
oviposition preference. This suggests that the loss of larval
parasitoids would likely impose even greater constraints on the
trajectory of evolution than our current estimates indicate.
Interestingly, this increase in the strength of selection on gall midges
may reduce their densities, which would weaken selection imposed on
willows by the gall midge through an evolutionary trophic cascade. A
growing number of experiments over the past two decades have
demonstrated the presence and potential importance of indirect
evolutionary effects that emerge in ecological communities
\citep{Pilson1996, Juenger1998, Stinchcombe2001, Lankau2007, Walsh2008, Walsh2010, terHorst2010, Sahli2011, Lau2012, terHorst2015, Schiestl2018, Start2019}.
If indirect evolutionary effects are common
\citep{Miller1996, Walsh2013, Guimaraes2017}, then predicting
evolutionary trajectories resulting from the loss of consumers will
require evolutionary studies to explicitly account for the ecological
networks that species are embedded in.

Our study gives insight to how the loss of consumers alters natural
selection, and in turn the evolutionary trajectory and adaptive
potential of remaining populations. In particular, it hints at a
potential insidious effect of local extinctions that compromises the
robustness of remaining populations to future environmental change. Our
work also highlights some key challenges for predicting phenotypic
evolution in rapidly changing communities. For example, many theoretical
models of eco-evolutionary dynamics focus on phenotypic change in a
single trait \citep[but see][]{Brown2007, Leimar2009}, yet our results
highlight that the number of traits under selection may change with the
network context. Importantly, we found that different species/guilds
imposed different selection pressures. Knowing these hidden selection
pressures is critical for prediction, because the trajectory of
evolution will depend on the nature of change in the ecological
community. We expect that a continued integration of fitness landscapes
and ecological networks will enhance our ability to predict the
evolutionary consequences of changes in ecological communities.

\section{Acknowledgements}\label{acknowledgements}

We thank the staff of Humboldt Bay National Wildlife Refuge (U.S. Fish
and Wildlife Service) for facilitating experimental logistics. For
assistance with fieldwork, we thank Ruthie Espanol and Andrew MacDonald.
This manuscript was improved by incorporating feedback from two
anonymous reviewers. For funding support, we thank the University of
British Columbia (James Robert Thompson Fellowship and Four-Year
Fellowship to M.A.~Barbour), NSERC (Discovery grant to Greg Crutsinger),
and the Swiss National Science Foundation (grant 31003A\_160671 to J.
Bascompte). The authors declare no conflict of interest.

\section{Author Contributions}\label{author-contributions}

M.A.B. conceived the idea behind the study and designed the field
experiment. M.A.B. and B.L. setup and conducted the experiment. M.A.B.,
A.S., and C.J.G. collected the data. M.A.B. analyzed the data. M.A.B.
wrote the manuscript with primary input from J.B. and additional
feedback from C.J.G.

\section{Data Accessibility}\label{data-accessibility}

All data and code to reproduce the reported results are publicly
available on GitHub
(\url{https://github.com/mabarbour/complexity_selection}) and have been
archived on Zenodo (\url{https://zenodo.org/badge/latestdoi/108833263}).

\bibliography{references}

\end{document}